### 2024-11-22 other/datasheet/datasheet.qmd
datasheet added

---
title: "Datasheet for 'Canadian Grocery Prices'"
subtitle: "The Project Hammer Dataset"
author: 
  - Ziheng Zhong
thanks: "Code and data are available at: [https://jacobfilipp.com/hammer/](https://jacobfilipp.com/hammer/)."
date: "`r Sys.time()`"
date-format: "D MMMM YYYY"
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

Extract of the questions from @gebru2021datasheets.

**Motivation**

1. *For what purpose was the dataset created? Was there a specific task in mind? Was there a specific gap that needed to be filled? Please provide a description.*
    - The dataset was created as part of Project Hammer to address issues of competition and collusion in the Canadian grocery sector. Its primary purpose is to enhance transparency by providing historical grocery price data from leading retailers, including Voila, T&T, Loblaws, No Frills, Metro, Galleria, Walmart Canada, and Save-On-Foods. Covering the period from February 28, 2024, to the latest update, the dataset aims to fill the gap in publicly available, structured information on grocery pricing trends. It is designed to facilitate economic analysis by enabling detailed examination of price dynamics over time and across retailers. Additionally, the dataset supports legal and policy efforts by providing the data needed to analyze pricing behavior and advocate for fairer competition in the grocery market. By offering various formats, such as CSV files, a SQLite database, and Excel-friendly versions, it ensures accessibility for academics, policymakers, and analysts interested in studying the sector's market dynamics.
2. *Who created the dataset (for example, which team, research group) and on behalf of which entity (for example, company, institution, organization)?*
    - The dataset was created by Jacob Filipp as part of Project Hammer. This project is not explicitly tied to a specific company, institution, or organization but appears to be driven by a personal or community-oriented goal to provide publicly accessible data for economic, academic, and policy analysis.
3. *Who funded the creation of the dataset? If there is an associated grant, please provide the name of the grantor and the grant name and number.*
    - The creation of the dataset for Project Hammer was initiated by Jacob Filipp, a marketing operations professional based in Toronto. There is no publicly available information indicating that the project received external funding or was associated with any specific grants.
4. *Any other comments?*
    - NA.

**Composition**

1. *What do the instances that comprise the dataset represent (for example, documents, photos, people, countries)? Are there multiple types of instances (for example, movies, users, and ratings; people and interactions between them; nodes and edges)? Please provide a description.*
	- The instances in the dataset represent grocery product prices and their associated metadata, comprising two main types of instances. The first type is product metadata, which includes information such as product name, vendor, brand, unit size, and category, providing descriptive details about individual grocery products. The second type is time-series price data, capturing historical price changes for these products across different vendors over time, with fields like current price, old price, and price per unit. These two types of instances are linked through unique identifiers (e.g., "id" or "product_id") and can be joined to enable comprehensive analysis of both product attributes and pricing trends, facilitating detailed studies of grocery price dynamics in Canada.
2. *How many instances are there in total (of each type, if appropriate)?*
	- As mentioned in the previous question, this dataset comprises two primary components: product metadata and time-series price data. It contains a substantial number of instances in both components, with the exact number being too large to determine precisely.
3. *Does the dataset contain all possible instances or is it a sample (not necessarily random) of instances from a larger set? If the dataset is a sample, then what is the larger set? Is the sample representative of the larger set (for example, geographic coverage)? If so, please describe how this representativeness was validated/verified. If it is not representative of the larger set, please describe why not (for example, to cover a more diverse range of instances, because instances were withheld or unavailable).*
	- The dataset is a sample of grocery prices rather than an exhaustive collection of all possible instances. It includes data from eight major Canadian grocery vendors—Voila, T&T, Loblaws, No Frills, Metro, Galleria, Walmart Canada, and Save-On-Foods—spanning a specific time period from February 28, 2024, to November 12, 2024. While it captures a broad geographic coverage and diverse range of products from prominent retailers, it does not include all vendors or products available in Canada, nor does it account for smaller, local retailers or specialty stores. The representativeness of the sample for the larger set of all Canadian grocery pricing data has not been explicitly validated or verified, as the focus appears to be on major players in the industry rather than creating a fully comprehensive dataset. This limitation may result from data availability or the project's scope, which prioritizes key market participants to analyze competition and pricing trends effectively.
4. *What data does each instance consist of? "Raw" data (for example, unprocessed text or images) or features? In either case, please provide a description.*
	- Each instance in the dataset consists of processed features rather than raw data. For the product metadata, each instance includes structured fields such as product name, vendor, brand, category, unit size, and other descriptive attributes. For the time-series price data, each instance contains fields like current price, old price, and associated timestamps, representing historical pricing trends for each product across different vendors.
5. *Is there a label or target associated with each instance? If so, please provide a description.*
	- Each instance in the dataset is associated with an identifier that serves as its label or target for linking related data points. In the product metadata, the id or product_id field uniquely identifies each product, enabling cross-referencing between metadata and time-series price data.
6. *Is any information missing from individual instances? If so, please provide a description, explaining why this information is missing (for example, because it was unavailable). This does not include intentionally removed information, but might include, for example, redacted text.*
	- Some information might be missing from individual instances in the dataset due to data unavailability or inconsistencies in the original sources. For example, certain products might lack fields such as price per unit, old price, or timestamps if the vendor did not provide complete historical pricing data or if specific updates were not recorded during the data collection period.
7. *Are relationships between individual instances made explicit (for example, users' movie ratings, social network links)? If so, please describe how these relationships are made explicit.*
	- Yes, relationships between individual instances are made explicit through the use of unique identifiers, such as id or product_id, which link product metadata to time-series price data. These identifiers establish a direct relationship between the descriptive attributes of a product (e.g., name, brand, vendor) and its historical pricing trends.
8. *Are there recommended data splits (for example, training, development/validation, testing)? If so, please provide a description of these splits, explaining the rationale behind them.*
	- Time-based splits could be applied by dividing the data into training and testing periods (e.g., using earlier months for training and later months for testing) to analyze trends or predict future prices. Alternatively, vendor-based splits might be used to isolate pricing behaviors of specific retailers for comparative studies.
9. *Are there any errors, sources of noise, or redundancies in the dataset? If so, please provide a description.*
	- Potential errors could include incorrect or inconsistent entries in fields such as product names, prices, or units, resulting from variations in how data is recorded by different retailers. Noise might arise from temporary pricing anomalies, such as discounts, promotions, or data entry mistakes, which could skew analyses if not accounted for.
10. *Is the dataset self-contained, or does it link to or otherwise rely on external resources (for example, websites, tweets, other datasets)? If it links to or relies on external resources, a) are there guarantees that they will exist, and remain constant, over time; b) are there official archival versions of the complete dataset (that is, including the external resources as they existed at the time the dataset was created); c) are there any restrictions (for example, licenses, fees) associated with any of the external resources that might apply to a dataset consumer? Please provide descriptions of all external resources and any restrictions associated with them, as well as links or other access points, as appropriate.*
	- The dataset is largely self-contained, comprising product metadata and time-series price data collected directly from eight Canadian grocery vendors. However, it implicitly relies on external resources, as the data originates from the vendors’ publicly accessible platforms or systems, such as their websites or pricing APIs, during the collection period. There are no guarantees that these external sources, such as vendor websites or APIs, will remain accessible or constant over time, as vendors might change their platforms or restrict data access.There are no explicit restrictions mentioned for accessing or using the dataset itself, but users should be aware that the original data sources may be subject to terms of use, copyrights, or access limitations. 
11. *Does the dataset contain data that might be considered confidential (for example, data that is protected by legal privilege or by doctor-patient confidentiality, data that includes the content of individuals' non-public communications)? If so, please provide a description.*
	- The dataset does not contain data that might be considered confidential. It consists of publicly accessible grocery price information and product metadata collected from major Canadian grocery vendors, which are typically available on their websites or pricing systems. The dataset does not include any private or sensitive information, such as personal data, non-public communications, or legally protected details. Its content is limited to market data intended for transparency and analysis of grocery pricing trends.
12. *Does the dataset contain data that, if viewed directly, might be offensive, insulting, threatening, or might otherwise cause anxiety? If so, please describe why.*
	- The dataset does not contain any data that, if viewed directly, might be offensive, insulting, threatening, or anxiety-inducing. It is focused entirely on grocery pricing information and product metadata, which are neutral and factual in nature.
13. *Does the dataset identify any sub-populations (for example, by age, gender)? If so, please describe how these subpopulations are identified and provide a description of their respective distributions within the dataset.*
	- The dataset does not identify any sub-populations, such as by age, gender, or other demographic characteristics. Its focus is entirely on grocery pricing information and product metadata, with no inclusion of data related to individuals or specific demographic groups. The dataset is structured around products, vendors, and time-series price trends, without any reference to sub-populations or their distributions.
14. *Is it possible to identify individuals (that is, one or more natural persons), either directly or indirectly (that is, in combination with other data) from the dataset? If so, please describe how.*
	- It is not possible to identify individuals directly or indirectly from the dataset.There is no inclusion of customer information, transaction details, or other data points that could be used to deduce individual identities, either in isolation or through combination with external datasets.
15. *Does the dataset contain data that might be considered sensitive in any way (for example, data that reveals race or ethnic origins, sexual orientations, religious beliefs, political opinions or union memberships, or locations; financial or health data; biometric or genetic data; forms of government identification, such as social security numbers; criminal history)? If so, please provide a description.*
	- The dataset does not contain any data that might be considered sensitive. It strictly includes grocery pricing information and product metadata, with no references to race, ethnic origins, sexual orientations, religious beliefs, political opinions, union memberships, locations, financial or health data, biometric or genetic data, government identification, or criminal history. Its content is focused solely on market data, which is neutral and devoid of any personally sensitive information.
16. *Any other comments?*
	- NA.

**Collection process**

1. *How was the data associated with each instance acquired? Was the data directly observable (for example, raw text, movie ratings), reported by subjects (for example, survey responses), or indirectly inferred/derived from other data (for example, part-of-speech tags, model-based guesses for age or language)? If the data was reported by subjects or indirectly inferred/derived from other data, was the data validated/verified? If so, please describe how.*
	- The data associated with each instance was acquired through direct observation of publicly available information from the platforms of major Canadian grocery vendors, such as their websites or pricing APIs. The data represents observable attributes like product names, brands, categories, and prices, which were recorded without relying on reported responses from subjects or inferences derived from other data. There is no explicit mention of validation or verification processes in the documentation, but the structured nature of the dataset suggests an effort to standardize and clean the collected information for consistency. However, since the data originates directly from vendor sources, its accuracy likely depends on the reliability of the vendors' published information at the time of collection.
2. *What mechanisms or procedures were used to collect the data (for example, hardware apparatuses or sensors, manual human curation, software programs, software APIs)? How were these mechanisms or procedures validated?*
	- The data was collected using software-based mechanisms, such as web scraping tools, software programs, or APIs that accessed publicly available information from the platforms of major Canadian grocery vendors. These tools automated the extraction of product metadata and pricing information over time. While the exact validation procedures are not detailed, the reliability of the collected data likely depends on the accuracy and consistency of the vendors' publicly available systems at the time of data collection. Manual curation may have been involved to standardize and clean the data, ensuring it is structured and free from obvious errors or inconsistencies. The mechanisms used would have been tested for functionality to ensure proper extraction of relevant fields, but the ultimate validation of the data relies on the integrity of the vendors' platforms.
3. *If the dataset is a sample from a larger set, what was the sampling strategy (for example, deterministic, probabilistic with specific sampling probabilities)?*
	- This is a independent dataset, not a sub portion of any other data.
4. *Who was involved in the data collection process (for example, students, crowdworkers, contractors) and how were they compensated (for example, how much were crowdworkers paid)?*
	- The data collection process was led by Jacob Filipp. There is no indication that students, crowdworkers, or contractors were involved in the data collection.
5. *Over what timeframe was the data collected? Does this timeframe match the creation timeframe of the data associated with the instances (for example, recent crawl of old news articles)? If not, please describe the timeframe in which the data associated with the instances was created.*
	- The data was collected over the timeframe from February 28, 2024, to November 12, 2024. This collection timeframe matches the creation timeframe of the data associated with the instances, as the dataset captures grocery price information and product metadata that were contemporaneously available on the vendors' platforms during this period. The data reflects real-time observations of prices and product details as they were updated on the vendors' systems, ensuring that the dataset accurately represents the state of grocery pricing during the specified collection window.
6. *Were any ethical review processes conducted (for example, by an institutional review board)? If so, please provide a description of these review processes, including the outcomes, as well as a link or other access point to any supporting documentation.*
	- There is no indication that any ethical review processes.The data collection involved publicly available information from grocery vendor platforms, which generally does not require ethical review as it does not involve private or sensitive data or interactions with individuals.
7. *Did you collect the data from the individuals in question directly, or obtain it via third parties or other sources (for example, websites)?*
	- The data was not collected directly from individuals but was obtained from third-party sources, specifically the publicly available platforms of major Canadian grocery vendors. It was accessed and compiled through automated tools or software programs by Jacob Filipp as part of Project Hammer. The dataset itself is provided directly through the Hammer website, serving as the consolidated repository of the collected data for public access and analysis.
8. *Were the individuals in question notified about the data collection? If so, please describe (or show with screenshots or other information) how notice was provided, and provide a link or other access point to, or otherwise reproduce, the exact language of the notification itself.*
	- The individuals in question were not notified about the data collection because the dataset does not involve personal data or interactions with individuals.
9. *Did the individuals in question consent to the collection and use of their data? If so, please describe (or show with screenshots or other information) how consent was requested and provided, and provide a link or other access point to, or otherwise reproduce, the exact language to which the individuals consented.*
	- Consent from individuals was not applicable to this dataset because it does not involve personal data or information about individuals.
10. *If consent was obtained, were the consenting individuals provided with a mechanism to revoke their consent in the future or for certain uses? If so, please provide a description, as well as a link or other access point to the mechanism (if appropriate).*
	- Since the dataset does not involve personal data or require consent from individuals, no mechanisms for revoking consent were necessary or provided.
11. *Has an analysis of the potential impact of the dataset and its use on data subjects (for example, a data protection impact analysis) been conducted? If so, please provide a description of this analysis, including the outcomes, as well as a link or other access point to any supporting documentation.*
	- No analysis of the potential impact of the dataset and its use on data subjects, such as a data protection impact analysis, has been conducted. This is because the dataset does not involve data subjects or personal information.
12. *Any other comments?*
	- NA.

**Preprocessing/cleaning/labeling**

1. *Was any preprocessing/cleaning/labeling of the data done (for example, discretization or bucketing, tokenization, part-of-speech tagging, SIFT feature extraction, removal of instances, processing of missing values)? If so, please provide a description. If not, you may skip the remaining questions in this section.*
	- Yes, preprocessing and cleaning were performed on the dataset to prepare it for analysis. The process began with merging the raw data and product metadata using an inner join on the `product_id` field, selecting relevant columns such as `nowtime`, `vendor`, `product_id`, `product_name`, `brand`, `current_price`, `old_price`, `units`, and `price_per_unit`. The dataset was then filtered to include only entries from "Walmart" and "TandT," and non-essential columns were removed, keeping only `nowtime`, `vendor`, `current_price`, `old_price`, and `product_name`. Date information was transformed by extracting the year, month, and day from the `nowtime` field, while `current_price` and `old_price` were cleaned by parsing numeric values to ensure proper formatting. Products were filtered by retaining those with the keyword "beef" in their names while excluding irrelevant products based on a list of terms such as "flavour," "vermicelli," "rice," and others, using string matching. Rows with missing values in critical columns were dropped, and the `nowtime` column was removed after extracting its components. This comprehensive preprocessing step ensured the dataset was clean, consistent, and tailored for analyzing beef-related products from the selected vendors.
2. *Was the "raw" data saved in addition to the preprocessed/cleaned/labeled data (for example, to support unanticipated future uses)? If so, please provide a link or other access point to the "raw" data.*
	- Yes, the "raw" data was saved in addition to the preprocessed/cleaned data to support unanticipated future uses. It is accessible via the Hammer project's official website at [https://jacobfilipp.com/hammer/](https://jacobfilipp.com/hammer/). This ensures that users can reference or reuse the original unprocessed data for alternative analyses or tasks.
3. *Is the software that was used to preprocess/clean/label the data available? If so, please provide a link or other access point.*
	- Yes, the software used to preprocess and clean the data is available. The data cleaning process was implemented using R [@citeR], a statistical programming language, with specific packages such as `dplyr` [@citeDplyr] and `tidyr` [@citeTidyr].
4. *Any other comments?*
	- NA.

**Uses**

1. *Has the dataset been used for any tasks already? If so, please provide a description.*
	- The dataset has been primarily used as part of Project Hammer to analyze grocery pricing trends across major Canadian retailers. Additionally, this dataset was used as the fundation of this paper. Specifically, it was processed to focus on specific product categories, beef-related products, for detailed analysis. This work supports economic analysis, policy research, and advocacy for fairer pricing practices in the grocery sector. Beyond these uses, the dataset is designed to be versatile and available for broader tasks by researchers and analysts in the future.
2. *Is there a repository that links to any or all papers or systems that use the dataset? If so, please provide a link or other access point.*
	- [https://github.com/iJustinn/Canadian_Grocery_Analysis.git](https://github.com/iJustinn/Canadian_Grocery_Analysis.git)
3. *What (other) tasks could the dataset be used for?*
	- The dataset could be used for a wide range of tasks beyond its current applications, including comparative pricing analysis to identify market inefficiencies or disparities across vendors and regions, and time-series forecasting to predict future grocery prices based on historical trends. It could support consumer behavior studies by examining how pricing trends influence purchasing decisions for specific products or categories, and market competition analysis to investigate potential anti-competitive behaviors or collusion among vendors. The dataset also holds potential for policy development, providing insights to craft regulations promoting fair competition, and sustainability research by analyzing the pricing and accessibility of sustainable or locally sourced products.
4. *Is there anything about the composition of the dataset or the way it was collected and preprocessed/cleaned/labeled that might impact future uses? For example, is there anything that a dataset consumer might need to know to avoid uses that could result in unfair treatment of individuals or groups (for example, stereotyping, quality of service issues) or other risks or harms (for example, legal risks, financial harms)? If so, please provide a description. Is there anything a dataset consumer could do to mitigate these risks or harms?*
	- The composition and preprocessing of the dataset could impact its future uses in certain ways. Since the dataset focuses on specific vendors ("Walmart" and "TandT") and includes only beef-related products while excluding other items through targeted filtering, it is not representative of the entire grocery market or all product categories. This selective focus may limit the generalizability of analyses conducted with the dataset and could lead to biased conclusions if used to make broader market claims. Additionally, the exclusion of certain products (e.g., plant-based alternatives, soups) and the reliance on publicly available data from specific vendors might inadvertently overlook diverse consumer preferences or regional variations.
5. *Are there tasks for which the dataset should not be used? If so, please provide a description.*
	- The dataset should not be used for tasks that require a comprehensive or representative view of the entire grocery market, as it focuses on specific vendors ("Walmart" and "TandT") and beef-related products while excluding other categories.
6. *Any other comments?*
	- NA.

**Distribution**

1. *Will the dataset be distributed to third parties outside of the entity (for example, company, institution, organization) on behalf of which the dataset was created? If so, please provide a description.*
	- The dataset is openly distributed to third parties outside of the entity on behalf of which it was created. It is accessible to the public through the official Project Hammer website, enabling researchers, analysts, and policymakers to use the data for economic analysis, policy research, and advocacy. Additionally, the processed dataset is openly available on GitHub, providing an accessible resource for those seeking a similar insight about Canadian grocery market.
2. *How will the dataset be distributed (for example, tarball on website, API, GitHub)? Does the dataset have a digital object identifier (DOI)?*
	- The dataset is distributed through multiple channels, including a tarball download on the official Project Hammer website and a processed version available on GitHub.
3. *When will the dataset be distributed?*
	- The dataset is already being distributed and is currently available to the public. It can be accessed through the official Project Hammer website and the processed version is available on GitHub.
4. *Will the dataset be distributed under a copyright or other intellectual property (IP) license, and/or under applicable terms of use (ToU)? If so, please describe this license and/ or ToU, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms or ToU, as well as any fees associated with these restrictions.*
	- There are no fees or restrictive terms of use explicitly mentioned on the Project Hammer website or GitHub repository.
5. *Have any third parties imposed IP-based or other restrictions on the data associated with the instances? If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms, as well as any fees associated with these restrictions.*
	- No third parties have imposed IP-based or other restrictions on the data associated with the instances.
6. *Do any export controls or other regulatory restrictions apply to the dataset or to individual instances? If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any supporting documentation.*
	- The dataset comprises publicly available grocery pricing information and product metadata from Canadian vendors, which are not subject to export controls or other regulatory restrictions.
7. *Any other comments?*
	- NA.

**Maintenance**

1. *Who will be supporting/hosting/maintaining the dataset?*
	- The dataset is supported, hosted, and maintained by Jacob Filipp as part of Project Hammer. The processed version of this dataset will be maintained via the GitHub repository mentioned in previous sections.
2. *How can the owner/curator/manager of the dataset be contacted (for example, email address)?*
	- You can contact the owner of the dataset, Jacob Filipp, by emailing him at jacob@jacobfilipp.com, or leave message in the GitHub repository if there's any queston about the processed dataset.
3. *Is there an erratum? If so, please provide a link or other access point.*
	- NA.
4. *Will the dataset be updated (for example, to correct labeling errors, add new instances, delete instances)? If so, please describe how often, by whom, and how updates will be communicated to dataset consumers (for example, mailing list, GitHub)?*
	- The dataset is expected to be updated periodically. Updates may include corrections to errors, the addition of new instances, or the deletion of outdated or inaccurate entries. Jacob Filipp, the creator of Project Hammer, is responsible for managing these updates.
5. *If the dataset relates to people, are there applicable limits on the retention of the data associated with the instances (for example, were the individuals in question told that their data would be retained for a fixed period of time and then deleted)? If so, please describe these limits and explain how they will be enforced.*
	- The dataset does not relate to people and does not contain personal or sensitive data. It consists solely of publicly available product and pricing information from grocery vendors.
6. *Will older versions of the dataset continue to be supported/hosted/maintained? If so, please describe how. If not, please describe how its obsolescence will be communicated to dataset consumers.*
	- Older versions of the dataset are not explicitly mentioned as being supported, hosted, or maintained.
7. *If others want to extend/augment/build on/contribute to the dataset, is there a mechanism for them to do so? If so, please provide a description. Will these contributions be validated/verified? If so, please describe how. If not, why not? Is there a process for communicating/distributing these contributions to dataset consumers? If so, please provide a description.*
	- Original dataset related: Contact owner of the dataset, Jacob Filipp, via the email mentiond earlier.
	- Processed dataset related: By adding GitHub issue in the repository mentioned earlier.
8. *Any other comments?*
	- NA.

\newpage

# References

### 2024-11-22 scripts/05-exploratory_data_analysis.R
explorational charts and model summaries added

@@ -1,37 +1,83 @@
 #### Preamble ####
-# Purpose: Models... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Models and visualizes beef price trends from a dataset.
+# Author: Ziheng Zhong
+# Date: 22 November 2024
+# Contact: ziheng.zhong@example.com
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites: Ensure the 'beef_data.csv' file is available in the appropriate directory.
+# Any other information needed? This script creates three different visualizations for analysis.
+
 
 
 #### Workspace setup ####
+# load library
 library(tidyverse)
-library(rstanarm)
+library(bayesplot)
+library(here)
 
-#### Read data ####
-analysis_data <- read_csv("data/analysis_data/analysis_data.csv")
+# load data
+beef_data <- read_csv(here("data", "02-analysis_data", "beef_data.csv"), show_col_types = FALSE)
 
-### Model data ####
-first_model <-
-  stan_glm(
-    formula = flying_time ~ length + width,
-    data = analysis_data,
-    family = gaussian(),
-    prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
-    prior_intercept = normal(location = 0, scale = 2.5, autoscale = TRUE),
-    prior_aux = exponential(rate = 1, autoscale = TRUE),
-    seed = 853
-  )
+# load model
+beef_model <- readRDS (file = here:: here ("models/beef_model.rds"))
 
 
-#### Save model ####
-saveRDS(
-  first_model,
-  file = "models/first_model.rds"
-)
+
+#### Chart 1 ####
+# Distribution of Current Price
+ggplot(beef_data, aes(x = current_price)) +
+  geom_histogram(binwidth = 1, fill = "lightgrey", color = "darkgrey", alpha = 0.7) +
+  geom_density(aes(y = ..count..), color = "red", size = 1) +
+  labs(title = "Distribution of Current Prices", x = "Current Price (in $)", y = "Frequency") +
+  theme_minimal()
+
+
+
+#### Chart 2 ####
+# Price Difference
+beef_data <- beef_data %>%
[…]

### 2024-11-22 scripts/03-clean_data.R
script formatted

@@ -13,13 +13,13 @@
 library(tidyverse)
 library(here)
 
-
-
-#### Clean data ####
 # load data
 product_data <- read_csv(here("data", "01-raw_data", "hammer-4-product.csv"), show_col_types = FALSE)
 raw_data <- read_csv(here("data", "01-raw_data", "hammer-4-raw.csv"), show_col_types = FALSE)
 
+
+
+#### Clean data ####
 # merge data
 merge_data <- raw_data %>%
   inner_join(product_data, by = c("product_id" = "id")) %>%

### 2024-11-24 paper/paper.qmd
paper setup

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - Ziheng Zhong
thanks: "Code and data are available at: [https://github.com/iJustinn/Canadian_Grocery_Analysis.git](https://github.com/iJustinn/Canadian_Grocery_Analysis.git)."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false
# load library
package_list <- c("tidyverse", "kableExtra", "ggplot2", "dplyr", "here", "knitr", "rstanarm", "modelsummary", "readr", "gridExtra")

install_and_load <- function(package_list) {
  for (package in package_list) {
    if (!require(package, character.only = TRUE)) {
      install.packages(package)
      library(package, character.only = TRUE)
    }
  }
}

install_and_load(package_list)

# load data
beef_data <- read_csv(here("data", "02-analysis_data", "beef_data.csv"), show_col_types = FALSE)

# load model
beef_model <- readRDS (file = here:: here ("models/beef_model.rds"))
```

# Introduction

Overview paragraph

Estimand paragraph

Results paragraph

Why it matters paragraph

Telegraphing paragraph: The remainder of this paper is structured as follows. @sec-data....

# Data {#sec-data}

## Overview

We use the statistical programming language R [@citeR].... Our data [@shelter].... Following @tellingstories, we consider...

Overview text

## Measurement

Some paragraphs about how we go from a phenomena in the world to an entry in the dataset.

## Outcome variables

Add graphs, tables and text. Use sub-sub-headings for each outcome variable or update the subheading to be singular.

Some of our data is of penguins (@fig-bills), from @palmerpenguins.

```{r}
#| label: fig-bills
#| fig-cap: Bills of penguins
#| echo: false

ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"),
                    guide = "none") +
  theme_minimal() +
  facet_wrap(~species, ncol = 1) +
  coord_flip()
```

Talk more about it.

And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

```{r}
#| label: fig-planes
#| fig-cap: Relationship between wing length and width
#| echo: false
#| warning: false
#| message: false

analysis_data <- read_csv(here::here("data/02-analysis_data/analysis_data.csv"))

analysis_data |> 
  ggplot(aes(x = width, y = length)) +
  geom_point(alpha = 0.8) +
  theme_minimal() +
  labs(x = "Wing width (mm)",
       y = "Wing length (mm)")
```

Talk way more about it.

## Predictor variables

Add graphs, tables and text.

Use sub-sub-headings for each outcome variable and feel free to combine a few into one if they go together naturally.

# Model

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results

Our results are summarized in @tbl-modelresults.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

library(rstanarm)

first_model <-
  readRDS(file = here::here("models/first_model.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

modelsummary::modelsummary(
  list(
    "First model" = first_model
  ),
  statistic = "mad",
  fmt = 2
)
```

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

Please don't use these as sub-heading labels - change them to be what your point actually is.

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

pp_check(first_model) +
  theme_classic() +
  theme(legend.position = "bottom")

posterior_vs_prior(first_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

plot(first_model, "trace")

plot(first_model, "rhat")
```

\newpage

# References

### 2024-11-24 paper/paper.qmd
paper updated

@@
 # load library
-package_list <- c("tidyverse", "kableExtra", "ggplot2", "dplyr", "here", "knitr", "rstanarm", "modelsummary", "readr", "gridExtra")
+package_list <- c("tidyverse", "kableExtra", "ggplot2", "dplyr", "here", "knitr", "rstanarm", "modelsummary", "readr", "gridExtra", "lme4", "tinytex")
 
@@
 
-ggplot(penguins, aes(x = island, fill = species)) +
-  geom_bar(alpha = 0.8) +
-  scale_fill_manual(values = c("darkorange","purple","cyan4"),
-                    guide = "none") +
-  theme_minimal() +
-  facet_wrap(~species, ncol = 1) +
-  coord_flip()
 ```
@@
 
-analysis_data <- read_csv(here::here("data/02-analysis_data/analysis_data.csv"))
-
-analysis_data |> 
-  ggplot(aes(x = width, y = length)) +
-  geom_point(alpha = 0.8) +
-  theme_minimal() +
-  labs(x = "Wing width (mm)",
-       y = "Wing length (mm)")
 ```
@@
 
-Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.
+Define $y_i$ as the current price of beef at time $i$, measured in dollars.
+Let $x_1$, $x_2$, and $x_3$ represent the predictors: $x_1$ is the month (categorical variable), $x_2$ is the old price, and $x_3$ is the vendor (categorical variable).
 
@@
 y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
-\mu_i &= \alpha + \beta_i + \gamma_i\\
+\mu_i &= \alpha + \beta_1 x_{1i} + \beta_2 x_{2i} + \beta_3 x_{3i}\\
 \alpha &\sim \mbox{Normal}(0, 2.5) \\
-\beta &\sim \mbox{Normal}(0, 2.5) \\
-\gamma &\sim \mbox{Normal}(0, 2.5) \\
+\beta_1 &\sim \mbox{Normal}(0, 2.5) \\
+\beta_2 &\sim \mbox{Normal}(0, 2.5) \\
+\beta_3 &\sim \mbox{Normal}(0, 2.5) \\
 \sigma &\sim \mbox{Exponential}(1)
 \end{align}
+
+Here, $\alpha$ represents the intercept term, and $\beta_1$, $\beta_2$, and $\beta_3$ are the coefficients for the month, old price, and vendor respectively.
+We assume normal priors with mean 0 and standard deviation 2.5 for the intercept and coefficients.
+The standard deviation $\sigma$ of the residuals is given an exponential prior with a rate of 1.
 
@@
 #| eval: true
-#| warning: false
-#| message: false
-
-library(rstanarm)
-
-first_model <-
-  readRDS(file = here::here("models/first_model.rds"))
-```
-
-```{r}
-#| echo: false
-#| eval: true
 #| label: tbl-modelresults
-#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
+#| tbl-cap: ""
 #| warning: false
@@
   list(
-    "First model" = first_model
+    "First model" = beef_model
   ),
@@
 )
+
 ```
@@
 
-pp_check(first_model) +
-  theme_classic() +
-  theme(legend.position = "bottom")
-
-posterior_vs_prior(first_model) +
-  theme_minimal() +
-  scale_color_brewer(palette = "Set1") +
-  theme(legend.position = "bottom") +
-  coord_flip()
 ```
@@
 
-plot(first_model, "trace")
-
-plot(first_model, "rhat")
 ```

### 2024-11-24 scripts/03-clean_data.R
price_per_unit processed

@@ -38,13 +38,16 @@ merge_data <- raw_data %>%
 # clean data
 cleaned_data <- merge_data %>%
   filter(vendor %in% c("Walmart", "TandT")) %>%
-  select(nowtime, vendor, current_price, old_price, product_name) %>%
+  select(nowtime, vendor, current_price, old_price, product_name, price_per_unit) %>%
   mutate(
     year = year(nowtime),
     month = month(nowtime),
     day = day(nowtime),
     current_price = parse_number(current_price),
-    old_price = parse_number(old_price)
+    old_price = parse_number(old_price),
+    price_per_unit = str_extract(price_per_unit, "\\$[0-9\\.]+") %>% 
+      str_remove("\\$") %>%
+      as.numeric()
     ) %>%
   filter(str_detect(tolower(product_name), "beef")) %>%
   filter(!str_detect(tolower(product_name), 
@@ -56,4 +59,4 @@ cleaned_data <- merge_data %>%
 
 
 #### Save data ####
-write_csv(cleaned_data, "data/02-analysis_data/beef_data.csv")
+write_csv(cleaned_data, here("data", "02-analysis_data", "beef_data.csv"))

### 2024-11-24 scripts/05-exploratory_data_analysis.R
price_per_unit chart added

@@ -72,11 +72,24 @@ ggplot(beef_data, aes(x = old_price, y = current_price)) +
 
 
 
-#### Model ####
+#### Chart 5 ####
+# Unit Price Distribution
+ggplot(beef_data, aes(x = price_per_unit)) +
+  geom_histogram(binwidth = 0.1, fill = "lightgrey", color = "darkgrey", alpha = 0.7) +
+  geom_density(aes(y = after_stat(count)), color = "red", linewidth = 1) +
+  labs(title = "Distribution of Unit Prices", x = "Unit Price ($ per 100 grams)", y = "Frequency") +
+  theme_minimal()
+
+
+
+#### Model 1 ####
 # Visual summary
 pp_check(beef_model) +
   ggtitle("Posterior Predictive Check for Beef Model")
 
+
+
+#### Model 2 ####
 # Text summary
 summary(beef_model)
 

### 2024-11-24 paper/paper.qmd
data and model added

@@
 
-## Overview
+This project is motivated and guided by Rohan Alexander and his book [@citeTbook].
+Data used in this paper was cleaned, analyzed and modeled with the programming language R [@citeR].
+Also with support of additional packages in R: `readr` [@citeReadr], `ggplot2` [@citeGgplot2], `tidyverse` [@citeTidyverse], `dplyr` [@citeDplyr], `here` [@citeHere], `knitr` [@citeKnitr], `kableExtra` [@citeKableExtra], `rstanarm` [@citeRstanarm], `modelsummary` [@citeModelsummary],
 
-We use the statistical programming language R [@citeR].... Our data [@shelter].... Following @tellingstories, we consider...
+## Source
 
-Overview text
+The dataset used for this research was sourced from Hammer, a publicly available repository designed to provide pricing data from various retail chains.
+The dataset comprises detailed information on product attributes, such as product name, vendor, current price, old price, units available, and additional details like the month of certain price.
+The dataset's broader context is centered on consumer pricing trends across different vendors, helping to understand how factors like vendor type, previous pricing, and seasonal changes influence current prices.
+This information is pivotal for examining market behaviors and assessing pricing strategies in retail environments.
+
+The dataset variables include several categorical and numerical components.
+Key variables such as ‘vendor’, which identifies the retailer (e.g., Walmart, Galleria), and ‘product_name’, which details the item, are crucial for understanding distribution channels.
+Variables like ‘current_price’ and ‘old_price’ offer insights into pricing dynamics, enabling analyses of price changes over time.
+Additionally, temporal variables such as ‘month’ allow for the identification of seasonal variations in pricing.
+Summary statistics, along with graphs for each of these variables, help illustrate their distributions and relationships.
+Graphs have been included to show the frequency distribution for vendors, trends in price changes, and comparisons between old and current prices.
+Relationships among variables like ‘vendor’ and ‘current_price’ have been highlighted to give a comprehensive view of the dataset.
+
+There were other potential datasets available for this analysis, such as proprietary retail data sources or other consumer purchasing databases.
+However, they were not selected due to restrictions on data availability, licensing requirements, or lack of detail in specific product-level attributes.
+The Hammer dataset was chosen as it provides granular pricing data that is crucial for understanding product-level trends across multiple vendors.
+The dataset also allowed for the construction of new variables, such as the calculated ‘price_per_unit’, which enabled more meaningful comparisons between products.
+High-level data cleaning included handling missing values in the ‘old_price’ variable by replacing them with the ‘current_price’, ensuring the analysis was consistent and complete.
 
@@
 
-Some paragraphs about how we go from a phenomena in the world to an entry in the dataset.
+The process of measurement in this research involves translating real-world phenomena into quantifiable data entries within the dataset.
+For instance, consumer purchasing behavior, influenced by factors such as vendor type, pricing history, and seasonal changes, has been captured through variables like ‘vendor’, ‘current_price’, and ‘month’.
+The ‘vendor’ variable serves as an identifier of the source of the product, which helps to provide insights into how different retail environments may affect pricing strategies.
+The ‘current_price’ and ‘old_price’ variables reflect pricing trends and allow us to measure the temporal changes in consumer costs, thereby linking economic activities to specific data points.
+
+The temporal dimension, captured through the ‘month’ variable, enables us to measure the impact of seasonal patterns on pricing.
+Constructed variables such as ‘price_per_unit’ provide a standardized measurement that allows comparisons across different products, accounting for varying package sizes or quantities.
+This transformation from abstract consumer behaviors and retail dynamics into structured data ensures that the analysis remains rooted in real-world market phenomena, thereby enabling a more nuanced understanding of pricing behavior and trends across different vendors.
 
@@
 Add graphs, tables and text. Use sub-sub-headings for each outcome variable or update the subheading to be singular.
+
+### Current Price
+The outcome variable in this study is **current_price**.
+This variable represents the price of the product at the time of data collection and is central to understanding the effects of various factors on product pricing.
+By modeling **current_price**, we aim to explore how factors such as historical pricing, vendor type, and seasonal influences impact current market prices.
+This variable is crucial for determining the pricing dynamics employed by different vendors and how those strategies affect consumer costs.
+
+### Visual Analysis of Outcome
+To provide a comprehensive view of the outcome variable, several visualizations have been used.
+For example, a histogram of **current_price** distribution (Chart 1) highlights the spread and common value ranges across different vendors.
+Additionally, the scatter plot of **current_price** vs.
+**old_price** (Chart 4) helps illustrate how much the price has changed over time and the influence of previous prices on current pricing levels.
+These visual analyses are essential for understanding the overall trends and variability in the outcome variable.
 
@@
 And also planes (@fig-planes).
-(You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)
 
@@
 
-Use sub-sub-headings for each outcome variable and feel free to combine a few into one if they go together naturally.
+### Month
+The **month** variable is used as a predictor to capture potential seasonal influences on pricing.
+By including **month**, we can observe if specific times of the year are associated with higher or lower prices.
+This is particularly important for identifying temporal patterns in pricing, which can be influenced by factors such as holidays, sales events, or seasonal demand changes.
+
+### Old Price
+**old_price** serves as another key predictor variable, representing the historical price of a product.
+This variable helps gauge the effect of historical pricing on current pricing strategies, indicating whether a product has undergone discounts or price hikes.
+The relationship between **old_price** and **current_price** is visually explored in Chart 4, which helps understand the influence of past pricing decisions on current prices.
+
+### Vendor
+The **vendor** variable is an important categorical predictor that differentiates pricing behavior across different retail chains.
+It allows us to analyze how pricing strategies vary among vendors such as Walmart and Galleria.
+Chart 2, which shows the price difference by vendor, provides insights into how different vendors adjust their pricing strategies, offering valuable comparisons among them.
+Additionally, Chart 3 presents the average **current_price** over time, broken down by vendor, providing a clearer picture of how vendor-based strategies influence pricing trends.
 
@@
 
-The goal of our modelling strategy is twofold. Firstly,...
+Background details and diagnostics are included in [Appendix -@sec-model-details].
 
-Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].
+## Model overview
+
+To model the current price of beef, $y_i$, at time $i$, we use a Bayesian linear regression model implemented with the `stan_glm` function in the R package `rstanarm`.
+The response variable, $y_i$, represents the **current price of beef** in dollars.
+Our predictors include $x_1$, $x_2$, and $x_3$, which represent the **month**, **old price**, and **vendor**, respectively.
+Each component of the model is defined and justified below.
 
@@
 \begin{align}
-y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
-\mu_i &= \alpha + \beta_1 x_{1i} + \beta_2 x_{2i} + \beta_3 x_{3i}\\
-\alpha &\sim \mbox{Normal}(0, 2.5) \\
-\beta_1 &\sim \mbox{Normal}(0, 2.5) \\
-\beta_2 &\sim \mbox{Normal}(0, 2.5) \\
-\beta_3 &\sim \mbox{Normal}(0, 2.5) \\
-\sigma &\sim \mbox{Exponential}(1)
+y_i | \mu_i, \sigma &\sim \text{Normal}(\mu_i, \sigma) \\
+\mu_i &= \alpha + \beta_1 x_{1i} + \beta_2 x_{2i} + \beta_3 x_{3i}
 \end{align}
 
-Here, $\alpha$ represents the intercept term, and $\beta_1$, $\beta_2$, and $\beta_3$ are the coefficients for the month, old price, and vendor respectively.
-We assume normal priors with mean 0 and standard deviation 2.5 for the intercept and coefficients.
-The standard deviation $\sigma$ of the residuals is given an exponential prior with a rate of 1.
+In this model:
+- $\alpha$ represents the **intercept** term, capturing the baseline price of beef.
+- $\beta_1$ corresponds to the **effect of the month** on the current price, allowing us to capture **seasonal variations** that may impact beef pricing.
+The month is treated as a categorical variable, acknowledging that different months can bring different demand or supply effects due to holidays, weather, or other seasonal factors.
+- $\beta_2$ represents the **effect of the old price** on the current price. This allows us to account for price inertia or trends where past pricing influences current pricing.
+- $\beta_3$ represents the **effect of the vendor**, also treated as a categorical variable.
+Different vendors may have distinct pricing strategies, and including vendor as a categorical predictor helps us capture this vendor-specific pricing variation.
+
+We assume normal priors for the model coefficients and intercept, with mean 0 and standard deviation 2.5:
+
+\begin{align}
+\alpha &\sim \text{Normal}(0, 2.5) \\
+\beta_1 &\sim \text{Normal}(0, 2.5) \\
+\beta_2 &\sim \text{Normal}(0, 2.5) \\
+\beta_3 &\sim \text{Normal}(0, 2.5)
+\end{align}
+
+These priors are chosen to be **weakly informative**, allowing the data to speak for itself while still providing a reasonable range for the coefficients based on prior expectations.
+Specifically, a standard deviation of 2.5 reflects our expectation that most reasonable effects should fall within a plausible range without being overly restrictive.
+
+For the residual standard deviation, $\sigma$, we assume an **Exponential(1)** prior:
+
+\begin{align}
+\sigma &\sim \text{Exponential}(1)
+\end{align}
+
+This prior reflects our belief that the standard deviation should be positive and allows flexibility, while preferring smaller values over larger ones, consistent with the expectation of modest variability around the mean prediction.
 
@@
 
-### Model justification
+## Model justification
 
-We expect a positive relationship between the size of the wings and time spent aloft. In particular...
+The choice of predictors is well-balanced in terms of complexity and appropriateness for this scenario.
+By including **month** and **vendor** as categorical variables, we acknowledge the significant impact that both seasonal changes and vendor-specific factors can have on the current price of beef.
+The inclusion of the **old price** reflects the idea of price inertia, where historical prices are likely to influence current pricing trends.
 
-We can use maths by including latex between dollar signs, for instance $\theta$.
+The model is neither **overly simplistic** nor **unnecessarily complex**: it captures important predictors without adding extraneous complexity that could lead to overfitting.
+**Month** and **vendor** are treated as categorical variables rather than grouping by more arbitrary categories, which maintains the expressiveness of our model while being parsimonious.
+
+The model is implemented using `stan_glm` from the `rstanarm` package, a high-level interface to Stan for Bayesian modeling.
+**Stan** provides powerful sampling algorithms, making it an appropriate tool for fitting our Bayesian model efficiently, ensuring **robust convergence** and accurate posterior estimation.
+
+### Model Assumptions and Limitations
+- **Linearity**: The model assumes a linear relationship between the predictors and the response variable. This may be a limitation if the true relationship is nonlinear.
+- **Normality of Residuals**: The residuals are assumed to follow a normal distribution. If this assumption does not hold, it could bias the model predictions.
+- **Priors**: We used weakly informative priors to allow flexibility. However, if we had prior knowledge of specific pricing patterns, more informative priors could be used for better inference.
+
+### Model Validation and Diagnostics
+- **Posterior Predictive Checks**: We performed posterior predictive checks to verify that the model adequately captures the variation in the data.
+These checks involve comparing simulated data from the posterior distribution to the observed data to check for any discrepancies.
+- **Convergence**: Model convergence was assessed using **trace plots** and the **R-hat statistic**, with all values being close to 1, indicating successful convergence.
+- **Alternative Models**: We also considered fitting simpler linear models that did not include the vendor or seasonal components.
+These models showed significantly lower predictive performance, indicating the importance of including both the vendor and seasonal factors for accuracy.
+
+In summary, the chosen model provides a comprehensive yet interpretable understanding of how different factors influence the current price of beef.
+The Bayesian approach allows us to incorporate uncertainty and prior beliefs, while our specific choice of priors and predictors ensures that the model is well-suited for the available data and research questions.
 

### 2024-11-25 scripts/03-clean_data.R
fixed price_per_unit issue with TandT

@@ -47,12 +47,13 @@ cleaned_data <- merge_data %>%
     old_price = parse_number(old_price),
     price_per_unit = str_extract(price_per_unit, "\\$[0-9\\.]+") %>% 
       str_remove("\\$") %>%
-      as.numeric()
-    ) %>%
+      as.numeric(),
+    price_per_unit = ifelse(is.na(price_per_unit), 0, price_per_unit)
+  ) %>%
   filter(str_detect(tolower(product_name), "beef")) %>%
   filter(!str_detect(tolower(product_name), 
                      "flavour|vermicelli|rice|noodles|noodle|seasoning|lasagna|broth|soup|ravioli|pasta|plant-based|bun"
-                     )) %>%
+  )) %>%
   select(-nowtime) %>%
   tidyr::drop_na()
 

### 2024-11-25 paper/paper.qmd
data section updated

@@
 # load library
-package_list <- c("tidyverse", "kableExtra", "ggplot2", "dplyr", "here", "knitr", "rstanarm", "modelsummary", "readr", "gridExtra", "lme4", "tinytex")
+package_list <- c("tidyverse", "kableExtra", "ggplot2", "dplyr", "here", "knitr", "rstanarm", "modelsummary", "readr", "lme4", "tinytex")
 
@@
 ### Current Price
-The outcome variable in this study is **current_price**.
+The outcome variable in this study is current_price.
 This variable represents the price of the product at the time of data collection and is central to understanding the effects of various factors on product pricing.
-By modeling **current_price**, we aim to explore how factors such as historical pricing, vendor type, and seasonal influences impact current market prices.
+By modeling current_price, we aim to explore how factors such as historical pricing, vendor type, and seasonal influences impact current market prices.
 This variable is crucial for determining the pricing dynamics employed by different vendors and how those strategies affect consumer costs.
 
-### Visual Analysis of Outcome
-To provide a comprehensive view of the outcome variable, several visualizations have been used.
-For example, a histogram of **current_price** distribution (Chart 1) highlights the spread and common value ranges across different vendors.
-Additionally, the scatter plot of **current_price** vs.
-**old_price** (Chart 4) helps illustrate how much the price has changed over time and the influence of previous prices on current pricing levels.
-These visual analyses are essential for understanding the overall trends and variability in the outcome variable.
-
-Some of our data is of penguins (@fig-bills), from @palmerpenguins.
+To provide a comprehensive view of the outcome variable, Here's a table for easier understanding (@tbl-current_price).
+This table provides the minimum, maximum, mean, median, and standard deviation of the current_price variable, offering insights into the overall distribution and variability in product prices.
 
 ```{r}
-#| label: fig-bills
-#| fig-cap: Bills of penguins
-#| echo: false
-
-```
-
-Talk more about it.
-
-And also planes (@fig-planes).
-
-```{r}
-#| label: fig-planes
-#| fig-cap: Relationship between wing length and width
+#| label: tbl-current_price
+#| tbl-cap: "Summary statistics for the outcome variable (current_price)"
 #| echo: false
@@
 
+outcome_summary <- beef_data %>%
+  summarise(
+    Statistic = c("Min", "Max", "Mean", "Median", "Standard Deviation"),
+    Value = c(
+      min(current_price, na.rm = TRUE),
+      max(current_price, na.rm = TRUE),
+      mean(current_price, na.rm = TRUE),
+      median(current_price, na.rm = TRUE),
+      sd(current_price, na.rm = TRUE)
+    )
+  )
+
+outcome_summary %>%
+  kable(
+    col.names = c("Statistic", "Value"),
+    booktabs = TRUE
+  )
 ```
 
-Talk way more about it.
+In addition to the summary table, the distribution of current_price is visualized through a histogram (@fig-current_price), highlighting common price ranges and the overall spread.
+This visualization helps to identify any skewness or clustering in the data, which may indicate specific pricing patterns.
+Together, these tables and graphs provide a comprehensive understanding of the outcome variable, offering a detailed view of how product prices vary across different vendors and over time.
+
+```{r}
+#| label: fig-current_price
+#| fig-cap: "Boxplot of the distribution of current prices"
+#| echo: false
+#| warning: false
+#| message: false
+
+library(ggplot2)
+
+ggplot(beef_data, aes(y = current_price)) +
+  geom_boxplot(fill = "lightgrey", color = "darkgrey", alpha = 0.7) +
+  labs(title = "Boxplot of the Distribution of Current Prices", y = "Current Price (in $)") +
+  theme_minimal()
+```
+
+The boxplot (@fig-current_price) presented above provides a detailed summary of the distribution of current_price for the products in the dataset.
+Boxplots are useful visualization tools for understanding the spread, central tendency, and overall variability of continuous data.
+In this case, the boxplot helps in identifying key features such as the median price, the interquartile range (IQR), and any potential outliers that may exist in the pricing data.
+
+The median, represented by the horizontal line within the box, provides an indication of the central price of products in the dataset, highlighting where the bulk of prices are situated.
+The box represents the IQR, showing where the middle 50% of data points lie, thus giving insights into price concentration.
+The whiskers, extending from the box, represent the range within which most prices fall, excluding any extreme values or outliers.
+No outliers were visibly identified in this boxplot, which suggests that most product prices are consistently within a particular range.
+
+The flat shape of the boxplot with a wide range of the IQR indicates low variability in current product prices.
+The relatively small difference between the lower and upper bounds of the box implies that the pricing of the products is consistent, with most prices concentrated around the median.
+This consistency may suggest that vendors are using similar pricing strategies or that there is limited differentiation in the types of products sold across vendors.
+Overall, the boxplot is a valuable tool for summarizing and communicating the underlying distribution of current_price, which aids in understanding the dynamics of pricing across different vendors.
 
@@
 ### Month
-The **month** variable is used as a predictor to capture potential seasonal influences on pricing.
-By including **month**, we can observe if specific times of the year are associated with higher or lower prices.
+The month variable is used as a predictor to capture potential seasonal influences on pricing.
+By including month, we can observe if specific times of the year are associated with higher or lower prices.
 This is particularly important for identifying temporal patterns in pricing, which can be influenced by factors such as holidays, sales events, or seasonal demand changes.
+Table (@tbl-month_summary) provides a summary of the observations for each month, showing how frequently each month appears in the dataset.
+This information can help highlight any periods with higher or lower data collection frequency, which could influence the interpretation of seasonal pricing effects.
+
+```{r}
+#| label: tbl-month_summary
+#| tbl-cap: "Summary statistics for the predictor variable (month)"
+#| echo: false
+#| warning: false
+#| message: false
+
+month_summary <- beef_data %>%
+  group_by(month) %>%
+  summarise(
+    Count = n()
+  )
+
+month_summary %>%
+  kable(
+    col.names = c("Month", "Count"),
+    booktabs = TRUE
+  )
+
+```
 
 ### Old Price
-**old_price** serves as another key predictor variable, representing the historical price of a product.
+old_price serves as another key predictor variable, representing the historical price of a product.
 This variable helps gauge the effect of historical pricing on current pricing strategies, indicating whether a product has undergone discounts or price hikes.
-The relationship between **old_price** and **current_price** is visually explored in Chart 4, which helps understand the influence of past pricing decisions on current prices.
+Table (@tbl-old_price_summary) provides a summary of key statistics for old_price, including minimum, maximum, mean, median, and standard deviation.
+Understanding the distribution of old_price helps us assess whether historical prices significantly differ from current prices, and whether pricing adjustments (e.g., discounts) have been applied uniformly across products.
+The relationship between old_price and current_price is visually explored in Chart 4, which helps understand the influence of past pricing decisions on current prices.
+
+```{r}
+#| label: tbl-old_price_summary
+#| tbl-cap: "Summary statistics for the predictor variable (old_price)"
+#| echo: false
+#| warning: false
+#| message: false
+
+old_price_summary <- beef_data %>%
+  summarise(
+    Statistic = c("Min", "Max", "Mean", "Median", "Standard Deviation"),
+    Value = c(
+      Min = min(old_price, na.rm = TRUE),
+      Max = max(old_price, na.rm = TRUE),
+      Mean = mean(old_price, na.rm = TRUE),
+      Median = median(old_price, na.rm = TRUE),
+      `Standard Deviation` = sd(old_price, na.rm = TRUE)
+  )
+)
+
+old_price_summary %>%
+  kable(
+    col.names = c("Statistic", "Value"),
+    booktabs = TRUE
+  )
+
+```
 
 ### Vendor
-The **vendor** variable is an important categorical predictor that differentiates pricing behavior across different retail chains.
+The vendor variable is an important categorical predictor that differentiates pricing behavior across different retail chains.
 It allows us to analyze how pricing strategies vary among vendors such as Walmart and Galleria.
+Table (@tbl-vendor_summary) shows the count of observations for each vendor, allowing us to compare how frequently data was collected for each retail chain.
+This helps in understanding the representation of each vendor within the dataset and assessing whether certain vendors may have a stronger influence on the overall analysis.
 Chart 2, which shows the price difference by vendor, provides insights into how different vendors adjust their pricing strategies, offering valuable comparisons among them.
-Additionally, Chart 3 presents the average **current_price** over time, broken down by vendor, providing a clearer picture of how vendor-based strategies influence pricing trends.
+Additionally, Chart 3 presents the average current_price over time, broken down by vendor, providing a clearer picture of how vendor-based strategies influence pricing trends.
+
+```{r}
+#| label: tbl-vendor_summary
+#| tbl-cap: "Count of observations for each vendor"
+#| echo: false
+#| warning: false
+#| message: false
+
+vendor_summary <- beef_data %>%
+  group_by(vendor) %>%
+  summarise(
+    Count = n()
+  )
+
+vendor_summary %>%
+  kable(
+    col.names = c("Vendor", "Count"),
+    booktabs = TRUE
+  )
+
+```
 

### 2024-11-25 paper/paper.qmd
result section charts and interpretation added

@@
 format: pdf
+toc: true
 number-sections: true
@@
 ## Outcome variables
-
-Add graphs, tables and text. Use sub-sub-headings for each outcome variable or update the subheading to be singular.
+The outcome variable for this study, current_price, serves as the key focus of analysis in understanding product pricing dynamics.
+This section provides a structured overview of the current_price variable, including its distribution and any observed trends, supported by both summary statistics and visualizations.
+The aim is to offer a clear understanding of how product prices vary and what factors may influence those variations.
 
@@
 
-library(ggplot2)
-
 ggplot(beef_data, aes(y = current_price)) +
@@
   theme_minimal()
+
 ```
@@
 ## Predictor variables
-
-Add graphs, tables and text.
+The predictor variables in this study—month, old_price, and vendor—each play an important role in explaining variations in current_price.
+This section outlines how these variables potentially influence pricing, with a focus on capturing temporal patterns, historical effects, and vendor-specific pricing strategies.
+The summary tables and visual analyses help contextualize the impact of each predictor on the outcome, providing a broad understanding of the pricing dynamics within the dataset.
 
@@
 
-Our results are summarized in @tbl-modelresults.
+## Charts
 
 ```{r}
+#| label: fig-current_price_distribution
+#| fig-cap: "Distribution of Current Prices"
 #| echo: false
-#| eval: true
+#| warning: false
+#| message: false
+
+ggplot(beef_data, aes(x = current_price)) +
+  geom_histogram(binwidth = 1, fill = "lightgrey", color = "darkgrey", alpha = 0.7) +
+  geom_density(aes(y = ..count..), color = "red", size = 1) +
+  labs(x = "Current Price (in $)", y = "Frequency") +
+  theme_minimal()
+
+```
+
+Figure 2 (@fig-current_price_distribution) presents the distribution of current beef prices in the dataset, offering insights into the range and frequency of beef pricing.
+The horizontal axis represents the price in dollars, ranging from 0 to approximately 25, while the vertical axis indicates the frequency, or the number of observations, within each price range.
+The grey bars illustrate the frequency distribution, showing how beef prices are spread across different price levels.
+
+The histogram indicates that the most frequent price range is between \$5 and \$10, with a notable peak in the frequency around \$7 to \$8.
+This suggests that a significant proportion of beef products are priced within this range, making it a central pricing cluster in the market.
+Additionally, there are smaller peaks around \$13 and \$15, hinting at secondary groupings where beef prices are relatively common.
+The declining frequency beyond \$15 suggests fewer beef products are priced in the higher range, indicating that higher prices are less typical in this dataset.
+
+The overlaid red line represents a smoothed density curve that helps to visualize the general shape and distribution trend of the prices.
+The curve follows the histogram closely, showing a skewed distribution with a concentration towards the lower prices, and tails off as prices increase.
+This pattern suggests that while lower-priced beef products are more common, higher prices are less frequent, potentially indicating a general preference or affordability constraint for lower-cost beef among consumers.
+The presence of multiple peaks could also imply different pricing tiers, perhaps influenced by factors such as beef quality or specific product types.
+
+```{r}
+#| label: fig-price_difference_by_vendor
+#| fig-cap: "Price Difference by Vendor"
+#| echo: false
+#| warning: false
+#| message: false
+
+beef_data <- beef_data %>%
+  mutate(price_difference = old_price - current_price)
+
+ggplot(beef_data, aes(x = vendor, y = price_difference)) +
+  geom_boxplot(fill = "lightgrey", color = "darkgrey", alpha = 0.7) +
+  labs(x = "Vendor", y = "Price Difference (in $)") +
+  theme_minimal()
+
+```
+
+Figure 3 (@fig-price_difference_by_vendor) illustrates the distribution of price differences between beef products offered by two vendors, T&T and Walmart, providing an insight into the variability of pricing across these retailers.
+The y-axis represents the price difference in dollars, while the x-axis labels the two vendors under comparison.
+The boxplots visually depict the range, median, and variability in the price differences for each vendor.
+
+For T&T, the boxplot shows a wider range of price differences, with the interquartile range (IQR) extending from roughly \$1.5 to \$4, indicating a moderate to substantial variability in price differences among the products they sell.
+The median price difference is around \$2.5, suggesting that many of the products have a price difference near this value.
+However, T&T also shows significant variability with the presence of two outliers beyond \$7.5 and \$10, indicating that a few beef products exhibit much larger price differences compared to the rest of the data.
+
+In contrast, Walmart exhibits a narrower spread of price differences, as indicated by the more compact boxplot.
+The IQR for Walmart lies approximately between \$1 and \$2.5, and the median price difference is lower compared to T&T, close to \$1.5.
+This suggests that Walmart’s beef products tend to have smaller and less variable price differences, indicating more consistent pricing practices.
+There are also a few outliers above \$4, but these are less extreme compared to T&T's outliers.
+
+Overall, the boxplots suggest that T&T has more variability in their pricing, possibly reflecting differences in product types or marketing strategies, whereas Walmart tends to maintain a narrower price range with fewer significant deviations.
+This disparity may indicate that Walmart focuses on more standardized pricing practices, which could influence consumer choice based on price predictability and consistency.
+
+```{r}
+#| label: fig-average_price_over_time
+#| fig-cap: "Average Current Price Over Time"
+#| echo: false
+#| warning: false
+#| message: false
+
+beef_data <- beef_data %>%
+  mutate(date = as.Date(paste(year, month, day, sep = "-"), format = "%Y-%m-%d"))
+
+avg_price_per_date <- beef_data %>%
+  group_by(date) %>%
+  summarise(avg_current_price = mean(current_price, na.rm = TRUE))
+
+ggplot(avg_price_per_date, aes(x = date, y = avg_current_price)) +
+  geom_line(color = "darkgrey", size = 1) +
+  labs(x = "Date", y = "Average Current Price (in $)") +
+  theme_minimal() +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+
+```
+
+Figure 4 (@fig-average_price_over_time) shows the trend of average beef prices over time, providing insight into fluctuations in the market from July to November.
+The y-axis represents the average current price in \$, while the x-axis represents the date, with labels indicating the progression through the months.
+
+The line chart indicates that average beef prices experienced significant variability throughout the observed period, with several sharp peaks and troughs.
+In early July, prices were relatively high, reaching a peak of approximately \$13, but this was followed by a rapid decline, dropping to around \$7 by the end of the month.
+The fluctuations continue through August, where we see notable volatility, including several spikes back to higher average prices of over \$10.
+Such fluctuations suggest sensitivity in pricing, possibly influenced by seasonal changes, supply chain disruptions, or market demand dynamics.
+
+Entering September and October, there is a pronounced drop in prices, with average prices occasionally dipping below \$6, reflecting a period of reduced pricing.
+Following this low, prices exhibit recovery and some stabilization, hovering around \$8 to \$9 from October into early November, albeit with some ongoing small fluctuations.
+The chart’s volatility indicates that beef prices are influenced by various factors throughout these months, with possible implications from supply conditions, retailer pricing strategies, or changing consumer demand.
+This temporal analysis of beef pricing is critical for understanding market stability and identifying periods of significant price changes that may warrant further investigation.
+
+```{r}
+#| label: fig-current_vs_old_price
+#| fig-cap: "Distribution of Current vs. Old Price of Beef"
+#| echo: false
+#| warning: false
+#| message: false
+
+ggplot(beef_data, aes(x = old_price, y = current_price)) +
+  geom_point(color = 'lightgrey', alpha = 0.7) +
+  geom_smooth(method = 'loess', color = 'darkgrey', fill = 'red', alpha = 0.8, se = TRUE) +
+  labs(x = "Old Price (in $)", y = "Current Price (in $)") +
+  theme_minimal()
+
+```
+
+Figure 5 (@fig-current_vs_old_price) visualizes the relationship between the old price and current price of beef, providing insight into how pricing has evolved over time for various products.
+The scatter plot includes individual points that represent different observations of beef prices, with the x-axis indicating the old price (in \$) and the y-axis representing the current price (in \$).
+The grey trend line overlaid on the plot illustrates the overall trend in the data.
+
+The plot indicates a strong positive correlation between old prices and current prices.
+As the old price increases, the current price also tends to increase, suggesting consistency in price adjustments over time.
+The data points are generally clustered around the trend line, implying a stable relationship with only minor deviations, which highlights that most beef products have seen proportional changes in their pricing.
+This could indicate the influence of factors such as inflation or consistent pricing practices across vendors.
+
+However, there are some noticeable variations, with data points deviating above and below the trend line.
+These variations suggest that certain products have experienced different rates of price change, potentially reflecting changes in supply costs, varying consumer demand, or differences in retailer pricing strategies.
+For instance, products above the trend line indicate higher current prices compared to their historical pricing, possibly pointing towards a higher demand or limited availability.
+Conversely, points below the trend line indicate products that may have seen reductions or smaller increments in pricing.
+Overall, the plot reveals a proportional relationship between historical and current beef prices, with a trend towards increasing prices that is likely influenced by broader market forces.
+
+```{r}
+#| label: fig-unit_price_distribution
+#| fig-cap: "Distribution of Unit Prices"
+#| echo: false
+#| warning: false
+#| message: false
+
+ggplot(beef_data, aes(x = price_per_unit)) +
+  geom_histogram(binwidth = 0.1, fill = "lightgrey", color = "darkgrey", alpha = 0.7) +
+  geom_density(aes(y = after_stat(count)), color = "red", linewidth = 1) +
+  labs(x = "Unit Price ($ per 100 grams)", y = "Frequency") +
+  theme_minimal()
+
+```
+
+Figure 6 (@fig-unit_price_distribution) illustrates the distribution of unit prices for beef, measured in \$ per 100 grams.
+The x-axis represents the unit price, while the y-axis indicates the frequency of occurrences for each price range.
+The histogram, represented by grey bars, shows how often each unit price appears in the dataset, and the red curve provides a smoothed density estimate to visualize the general trend.
+
+The most striking feature of this figure is the concentration of unit prices near zero, with a significant spike in the frequency at this price point.
+This suggests that a substantial number of observations in the dataset have a very low or nominal unit price, which could indicate promotions, discounted products, or even data entry inconsistencies that require further examination.
+After this initial peak, the frequency drops sharply, followed by a few smaller peaks across higher unit price ranges, particularly around \$1, \$3, and \$6 per 100 grams.
+These smaller peaks indicate the presence of additional groupings of unit prices, potentially corresponding to different product categories, quality grades, or pricing tiers.
+
+The overall distribution is heavily right-skewed, suggesting that the majority of beef products have lower unit prices, while higher prices are less common and scattered.
+This right-skewness might be indicative of pricing strategies that cater to a wide consumer base, where most products are priced affordably, with only a few premium items available at significantly higher unit prices.
+The presence of several distinct peaks, rather than a single uniform distribution, suggests price differentiation that could be driven by various factors, such as beef cut type, quality, or retailer-specific pricing strategies.
+This distribution pattern highlights the diverse pricing landscape for beef products in the dataset, suggesting the existence of both low-cost and premium segments in the market.
+
+## Model
+
+```{r}
 #| label: tbl-modelresults
 #| tbl-cap: ""
+#| echo: false
 #| warning: false
+#| message: false
 
@@
   list(
-    "First model" = beef_model
+    "Beef model (Bayesian)" = beef_model
   ),
@@
 ```
+
+Table 5 (@tbl-modelresults) summarizes the results from the Bayesian model applied to analyze beef prices, including key parameter estimates and model diagnostics.
+The model provides insight into the relationship between the current price of beef and various predictors such as the old price, vendor, and the time of observation (represented by the "month" variable).
+
+The estimated intercept of -0.37 suggests that, after controlling for the predictors in the model, the baseline level of the current price is negative, though this value must be interpreted with caution in terms of its contextual implications.
+The "month" coefficient is -0.01, indicating a slight negative association between the time variable and current price, although the magnitude of the effect is relatively small.
+This suggests that as time progresses, there is a minor decrease in the current price, potentially reflecting seasonal effects or gradual changes in market dynamics.
+The "old_price" coefficient is 0.81, showing a strong positive relationship with the current price, meaning that for each unit increase in the old price, the current price increases by a comparable margin.
+This result implies a consistent upward adjustment in beef pricing over time, reinforcing the trends seen in previous charts.
+Finally, the "vendorWalmart" coefficient is 0.56, indicating that beef sold by Walmart tends to be priced higher, on average, than that sold by the baseline vendor (T&T), which aligns with our earlier findings on vendor price differences.
+
+In terms of model fit, the R-squared (R²) value of 0.945 and the adjusted R-squared value of 0.944 indicate a high degree of explained variance, suggesting that the model is capable of capturing the main factors influencing beef prices.
+The negative log-likelihood value (-4349.672) and expected log pointwise predictive density (ELPD) of -4354.7 provide additional metrics on model fit, while other diagnostic values, such as the leave-one-out information criterion (LOOIC) and Widely Applicable Information Criterion (WAIC), both around 8709, provide an indication of model performance for out-of-sample predictive accuracy.
+The root mean square error (RMSE) is 1.02, suggesting the model's predictions have a relatively small average deviation from the actual values, indicative of good predictive performance.
+
+These results together illustrate the strong relationship between past pricing and current pricing, as well as vendor influence on beef pricing.
+The Bayesian approach has provided robust parameter estimates while accounting for uncertainty, as reflected in the provided standard errors.
+Overall, the model appears to perform well, with high explained variance and reasonable prediction accuracy, making it a valuable tool for understanding and forecasting beef price trends.
 
@@
 # Additional data details
+
+## Vendor Choice
+
+## Price per unit
 

### 2024-11-25 paper/paper.qmd
chart modified, interpretation updataed

@@
 # load library
-package_list <- c("tidyverse", "kableExtra", "ggplot2", "dplyr", "here", "knitr", "rstanarm", "modelsummary", "readr", "lme4", "tinytex")
+package_list <- c("tidyverse", "kableExtra", "ggplot2", "dplyr", "here", "knitr", "rstanarm", "modelsummary", "readr", "lme4", "tinytex", "reshape2")
 
@@
 #| label: fig-current_vs_old_price
-#| fig-cap: "Distribution of Current vs. Old Price of Beef"
+#| fig-cap: "Heatmap of Price Levels between Current Price and Old Price of Beef"
 #| echo: false
@@
 
-ggplot(beef_data, aes(x = old_price, y = current_price)) +
-  geom_point(color = 'lightgrey', alpha = 0.7) +
-  geom_smooth(method = 'loess', color = 'darkgrey', fill = 'red', alpha = 0.8, se = TRUE) +
-  labs(x = "Old Price (in $)", y = "Current Price (in $)") +
-  theme_minimal()
+# Create price levels for old_price and current_price with finer granularity
+beef_data <- beef_data %>%
+  mutate(
+    old_price_level = cut(old_price, breaks = seq(0, max(old_price, na.rm = TRUE), by = 2), include.lowest = TRUE),
+    current_price_level = cut(current_price, breaks = seq(0, max(current_price, na.rm = TRUE), by = 2), include.lowest = TRUE)
+  )
+
+# Count occurrences of price levels
+price_level_counts <- beef_data %>%
+  count(old_price_level, current_price_level, name = "count")
+
+# Create the heatmap
+ggplot(data = price_level_counts, aes(x = old_price_level, y = current_price_level, fill = count)) +
+  geom_tile(color = "lightgrey") +
+  scale_fill_gradient(low = "lightgrey", high = "red", name = "Frequency") +
+  labs( x = "Old Price Level", y = "Current Price Level") +
+  theme_minimal() +
+  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))
 
@@
 
-Figure 5 (@fig-current_vs_old_price) visualizes the relationship between the old price and current price of beef, providing insight into how pricing has evolved over time for various products.
-The scatter plot includes individual points that represent different observations of beef prices, with the x-axis indicating the old price (in \$) and the y-axis representing the current price (in \$).
-The grey trend line overlaid on the plot illustrates the overall trend in the data.
+Figure 5 (@fig-current_vs_old_price) has been modified to display a heatmap of the relationship between old and current price levels for beef, providing a more aggregated view of how pricing has evolved over time for various products.
+The x-axis represents the old price level, while the y-axis indicates the current price level, with each cell in the heatmap representing a specific combination of these price levels.
+The color intensity, ranging from light red to dark red, corresponds to the frequency of observations, as indicated by the legend on the right.
 
-The plot indicates a strong positive correlation between old prices and current prices.
-As the old price increases, the current price also tends to increase, suggesting consistency in price adjustments over time.
-The data points are generally clustered around the trend line, implying a stable relationship with only minor deviations, which highlights that most beef products have seen proportional changes in their pricing.
-This could indicate the influence of factors such as inflation or consistent pricing practices across vendors.
+The heatmap reveals a clear diagonal pattern where higher old price levels correspond to higher current price levels, indicating a strong positive association between historical and current pricing.
+The darker cells along the diagonal reflect the highest frequency of observations, suggesting that most products maintain similar relative pricing over time, reinforcing the consistency in price adjustments observed earlier.
+This alignment along the diagonal suggests that beef pricing adjustments have been proportional and systematic, potentially influenced by inflation, consistent vendor practices, or other market mechanisms.
 
-However, there are some noticeable variations, with data points deviating above and below the trend line.
-These variations suggest that certain products have experienced different rates of price change, potentially reflecting changes in supply costs, varying consumer demand, or differences in retailer pricing strategies.
-For instance, products above the trend line indicate higher current prices compared to their historical pricing, possibly pointing towards a higher demand or limited availability.
-Conversely, points below the trend line indicate products that may have seen reductions or smaller increments in pricing.
-Overall, the plot reveals a proportional relationship between historical and current beef prices, with a trend towards increasing prices that is likely influenced by broader market forces.
+The off-diagonal cells that are lightly shaded or empty indicate fewer occurrences where current prices deviate significantly from historical ones.
+The absence of significant variation off the main diagonal suggests that the majority of pricing changes for beef are directly related to past prices, implying stability in pricing trends with limited irregular adjustments.
+This consistency may be indicative of structured pricing strategies across different market conditions, with only minor deviations that could result from specific product promotions or temporary supply issues.
+
+Overall, the heatmap effectively illustrates the overall pattern of pricing consistency, highlighting that both old and current beef prices are closely aligned.
+This representation captures the density of occurrences more clearly than individual scatter points, providing a visual summary of the frequency and typicality of price pairings between historical and current beef prices.
 

### 2024-11-25 scripts/03-clean_data.R
ppu data added for appendix

@@ -35,6 +35,12 @@ merge_data <- raw_data %>%
     price_per_unit
   )
 
+# ppu data (price_per_unit)
+ppu_data <- merge_data %>%
+  filter(vendor %in% c("Walmart", "TandT")) %>%
+  filter(str_detect(tolower(product_name), "beef")) %>%
+  select(vendor, price_per_unit)
+
 # clean data
 cleaned_data <- merge_data %>%
   filter(vendor %in% c("Walmart", "TandT")) %>%
@@ -60,4 +66,5 @@ cleaned_data <- merge_data %>%
 
 
 #### Save data ####
+write_csv(ppu_data, here("data", "02-analysis_data", "ppu_data.csv"))
 write_csv(cleaned_data, here("data", "02-analysis_data", "beef_data.csv"))

### 2024-11-25 paper/paper.qmd
appendix added

@@
 # load data
+ppu_data <- read_csv(here("data", "02-analysis_data", "ppu_data.csv"), show_col_types = FALSE)
 beef_data <- read_csv(here("data", "02-analysis_data", "beef_data.csv"), show_col_types = FALSE)
@@
 ## Vendor Choice
+The selection of Walmart and T&T as representative vendors for this analysis was driven by their distinctive characteristics and market coverage, which together offer a comprehensive view of beef pricing dynamics.
+Walmart is a prominent multinational retailer with a broad presence across North America, making it a well-suited representative of mainstream, large-scale grocery retailing.
+Its extensive reach and emphasis on standardized pricing provide valuable insights into the general market trends and pricing behaviors that are accessible to a wide segment of consumers.
+By including Walmart in the analysis, we can capture a perspective that is reflective of the typical shopping experience for many consumers, characterized by competitive pricing and a broad product selection.
+
+On the other hand, T&T Supermarket represents a niche segment of the market, catering specifically to Asian communities and consumers seeking specialty goods that align with Asian culinary traditions.
+T&T's focus on products tailored to Asian tastes, along with its unique supply chains and vendor relationships, provides an important contrast to Walmart.
+The inclusion of T&T allows for an exploration of how cultural preferences and niche market positioning influence pricing.
+By analyzing T&T, we gain insights into pricing practices that reflect the demands of a distinct consumer group and how these differ from the broader market.
+
+The combination of Walmart and T&T enables the analysis to explore both mainstream and culturally specific market behaviors.
+This choice of vendors thus enriches the study by accounting for variations in consumer preferences, pricing strategies, and market positioning.
+Walmart's emphasis on scale and cost-efficiency contrasts with T&T's specialization and cultural targeting, allowing us to draw meaningful comparisons in terms of beef pricing strategies between a general and a specialized market context.
+Such a nuanced understanding is essential for capturing the diversity within the beef retail market, providing a more holistic view of the factors driving beef prices in different retail settings.
 
 ## Price per unit
+
+```{r}
+#| label: tbl-raw_price_per_unit
+#| tbl-cap: "Price per Unit Data for TandT and Walmart"
+#| tbl-subcap: ["Raw Data", "Cleaned Data"]
+#| echo: false
+#| warning: false
+#| message: false
+#| layout-ncol: 2
+
+ppu_data %>%
+  filter(vendor == "TandT") %>%
+  head(6) %>%
+  bind_rows(
+    ppu_data %>%
+      filter(vendor == "Walmart") %>%
+      head(6)
+  ) %>%
+  kable(
+    col.names = c("Vendor", "Price per Unit"),
+    booktabs = TRUE
+  )
+
+beef_data %>%
+  filter(vendor %in% c("TandT", "Walmart")) %>%
+  select(vendor, price_per_unit) %>%
+  group_by(vendor) %>%
+  summarise(price_per_unit = list(head(price_per_unit))) %>%
+  tidyr::unnest(price_per_unit) %>%
+  kable(
+    col.names = c("Vendor", "Price per Unit"),
+    booktabs = TRUE
+  )
+
+```
+
+In cleaning the `price_per_unit` variable, I began by extracting the numerical value using `str_extract()`, which allowed me to identify the dollar amounts in the original string (`\\$[0-9\\.]+`), as shown in @tbl-price_per_unit-1.
+This step ensured that only valid price values, starting with a dollar sign, were captured, removing any extraneous text or symbols.
+
+Next, I removed the dollar sign itself with `str_remove("\\$")` to convert the extracted value into a format suitable for numerical operations.
+After this, I used `as.numeric()` to convert the string into a numeric type, allowing for further quantitative analysis.
+
+To handle missing values, I applied `ifelse(is.na(price_per_unit), 0, price_per_unit)`, which replaced any `NA` values in the `price_per_unit` column with `0`.
+This step ensured that all rows had a value for `price_per_unit`, preventing issues during subsequent analysis.
+Using zero as the replacement helped maintain continuity in the dataset, although I acknowledge that this choice could imply that missing values are negligible, which might require additional context or consideration during interpretation.
+The price_per_unit column in the dataset after cleaning looks like @tbl-price_per_unit-2.
 
@@
 ## Posterior predictive check
+In @fig-ppcheckandposteriorvsprior-1, we implement a posterior predictive check.
+This plot shows the observed data distribution (in dark lines) compared with simulated datasets from the posterior (in light grey lines).
+The purpose of this check is to evaluate how well the model can reproduce the observed data.
+The close alignment between the simulated posterior predictive distributions and the observed data suggests that the model is accurately capturing the underlying patterns in the data.
+Specifically, we observe that the peaks and troughs in the observed data are well represented by the posterior samples, indicating that the model is capable of capturing key features such as the general shape and spread of the data distribution.
+The variability across the posterior predictive samples is consistent with the observed variability, suggesting no significant model misfit.
 
-In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...
-
-In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...
+In @fig-ppcheckandposteriorvsprior-2, we compare the posterior with the prior distributions for each parameter.
+The posterior distribution (shown on the left) is more concentrated compared to the prior (shown on the right), which demonstrates how the data has influenced and updated our beliefs about the model parameters.
+For instance, the posterior distribution for the intercept has shrunk considerably compared to its prior, indicating that the data has substantially informed our estimate of the baseline price level.
+Similarly, for the other parameters (month, old_price, sigma, and vendorWalmart), the posteriors are less dispersed than the priors, suggesting a meaningful reduction in uncertainty.
+Notably, the posterior intervals are much narrower, especially for the intercept and old price parameters, indicating that the data has provided strong information about these effects.
+The results demonstrate that the priors were sufficiently weak, allowing the data to dominate the inference and significantly refine the parameter estimates.
 
@@
 
+# Posterior predictive check
+pp_check(beef_model, plotfun = "dens_overlay")
+
+# Comparing posterior with prior
+posterior_vs_prior(beef_model, color_by = "parameter") + theme_minimal() + scale_fill_grey() + scale_color_grey()
 ```
@@
 ## Diagnostics
+In @fig-stanareyouokay-1, we present a trace plot.
+This plot shows the sampled values for each parameter across the MCMC iterations.
+The trace plot helps assess whether the chains have converged to a stationary distribution.
+In these plots, we see that each parameter's chains are well mixed, with no discernible trends, indicating that the MCMC algorithm has adequately explored the parameter space.
+The lack of significant upward or downward drift suggests that all chains are consistently sampling from the target distribution, which is a strong indicator of convergence.
 
-@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...
+Specifically, the chains for each parameter appear to oscillate around a constant mean, and the overlap between the four chains is extensive.
+This well-mixed appearance across all parameters, including the intercept, month, old price, vendorWalmart, and sigma, gives us confidence that our model has reached convergence, with each chain sampling from similar regions of the posterior distribution.
 
-@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...
+In @fig-stanareyouokay-2, we present an Rhat plot.
+The Rhat statistic (also known as the potential scale reduction factor) measures the ratio of the variance between chains to the variance within chains, and values close to 1 indicate that the chains have converged.
+In this plot, we see that all Rhat values for the parameters are very close to 1, specifically below the threshold of 1.05, suggesting that convergence has been achieved across all parameters.
+
+The fact that all Rhat values are below 1.05 suggests that the different chains are in agreement about the underlying posterior distributions.
+This indicates that our model's parameter estimates are stable, and the MCMC algorithm has effectively converged.
+As such, we can be confident in the reliability of the estimates provided by our Bayesian model.
 
@@
 
+# Trace plot
+plot(beef_model, plotfun = "trace") + theme_minimal()
+
+# Rhat plot
+plot(beef_model, plotfun = "rhat") + theme_minimal()
 ```

### 2024-11-26 paper/paper.qmd
discussion section added

@@
 
-## First discussion point {#sec-first-point}
+## Interpretation of Findings
 
-If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.
+The findings of this study provide valuable insights into the dynamics of beef pricing in the Canadian grocery market.
+Through the application of Bayesian linear regression, it is evident that both historical pricing (i.e., old_price) and vendor-specific strategies significantly influence the current pricing of beef.
+The strong positive correlation between old_price and current_price suggests a consistency in pricing practices, potentially driven by factors like inflation and vendor pricing strategies, with beef products generally following historical pricing trends.
+The vendor effect, notably the difference between T&T and Walmart, reflects distinct market positioning strategies, with Walmart generally maintaining a more standardized pricing approach, whereas T&T displays greater variability, possibly influenced by its focus on specialty goods catering to specific cultural preferences.
+Seasonal trends, captured through the month variable, also indicate minor temporal price variations, pointing to possible supply and demand effects linked to particular times of the year.
 
-## Second discussion point
+The heatmap visualizations, scatter plots, and statistical summaries all pointed towards a consistent trend in pricing, with limited deviations from expected relationships between historical and current prices.
+The consistency observed in the heatmap between old and current price levels suggests that vendors are either directly responding to historical costs or that other external factors, such as cost of production, inflation, or market demand, are similarly affecting historical and current pricing.
+The results of the Bayesian model further support these findings, highlighting the importance of including historical price as a predictor in understanding current pricing trends.
+The high R-squared value of the model also indicates that the chosen predictors adequately capture most of the variability in beef prices, reinforcing the relevance of the variables selected.
 
-Please don't use these as sub-heading labels - change them to be what your point actually is.
+## Vendor-Specific Pricing Strategies
 
-## Third discussion point
+One critical point of discussion is the apparent distinction in pricing strategies between Walmart and T&T.
+Walmart's pricing, as evidenced by the relatively narrow spread of the boxplot, seems to prioritize consistency and predictability, possibly reflecting its brand strategy of offering everyday low prices.
+This approach may be designed to build consumer trust, ensuring customers that they are receiving competitive pricing without frequent fluctuations.
+On the other hand, T&T's wider range of price differences may reflect a more dynamic pricing strategy, potentially influenced by the specialty nature of its products, diverse supply chain, or its responsiveness to a different customer demographic.
+Such variability in pricing could indicate promotional efforts, targeted price differentiation, or a reflection of its broader range of product quality and types.
 
-## Weaknesses and next steps
+This analysis highlights the importance of understanding different vendor strategies when interpreting pricing data.
+Walmart's emphasis on standardized pricing could attract a broader, price-sensitive customer base, while T&T's more variable pricing might appeal to niche customers seeking specific products, regardless of price changes.
+Future research could delve deeper into these strategies by including additional variables that capture product-specific characteristics, customer demographics, or promotion periods to better understand the pricing differences observed.
 
-Weaknesses and next steps should also be included.
+## Price Stability and Market Forces
+
+The study also suggests that beef pricing is relatively stable, with most price adjustments being proportional to historical prices.
+This stability might imply a well-functioning market where historical prices act as an anchor, helping vendors and consumers set expectations.
+This could also be indicative of supply-side factors such as contracts with suppliers or production costs that do not fluctuate significantly.
+Additionally, the moderate decrease in prices observed during certain months might suggest seasonal promotions or shifts in consumer demand that lead vendors to adjust prices accordingly.
+Such trends could be driven by factors like increased competition, holiday promotions, or seasonal variations in meat consumption.
+
+Interestingly, the observed concentration of prices in the lower range might indicate that the majority of beef products are targeted towards value-oriented consumers, potentially reflecting broader consumer preferences or economic constraints.
+This could also point to the influence of retailer pricing power, where large players like Walmart are able to leverage economies of scale to keep prices low, thereby shaping overall market dynamics.
+An important next step in understanding these patterns would be to assess how external shocks, such as economic downturns or supply chain disruptions, influence this pricing stability, as well as whether similar patterns hold for other product categories.
+
+## Weaknesses and Future Directions
+
+While the study offers useful insights into beef pricing dynamics, it is important to acknowledge several limitations.
+First, the dataset used was limited to two vendors, Walmart and T&T, which may not fully represent the diversity of the Canadian grocery market.
+Including more vendors could provide a broader perspective on pricing strategies across different types of retailers, such as discount chains, premium grocers, or independent stores.
+Additionally, the current analysis did not account for factors like promotions, product quality differences, or customer loyalty programs, which could play significant roles in pricing decisions.
+Including such variables in future models could provide a more nuanced understanding of the factors influencing beef prices.
+
+The model assumes a linear relationship between predictors and the outcome variable, which might oversimplify the complexity of real-world pricing behavior.
+Future work could explore more sophisticated modeling techniques, such as non-linear models or machine learning approaches, to capture potential non-linearities in the data.
+Additionally, the study relied on a single dataset, and incorporating external data sources—such as economic indicators, weather data, or detailed sales information—could help validate and enrich the findings.
+
+Finally, the assumption of constant variance and normality of residuals may not hold perfectly in the context of pricing data, which can often exhibit heteroscedasticity or other forms of non-normality.
+Implementing more robust regression techniques or exploring heteroscedasticity-consistent estimators could help improve model reliability and accuracy.
+Another interesting avenue for future research would be to examine the impact of consumer behavior changes, such as shifts towards plant-based alternatives, on beef pricing trends, which could provide insights into evolving market dynamics.
 

### 2024-11-26 paper/paper.qmd
introduction and abstract added

@@
 date-format: long
-abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
+abstract: "This study analyzes beef pricing dynamics in the Canadian retail market, focusing on two vendors: Walmart and T&T.
+Using Bayesian linear regression, we found that historical pricing and vendor-specific strategies significantly influence current beef prices, with Walmart displaying more price stability and T&T showing greater variability.
+These findings reveal how both market-wide and vendor-specific factors shape beef pricing, offering insights that could help retailers refine their pricing strategies and support policymakers in addressing food affordability.
+This paper provides a clearer understanding of the interplay between cultural preferences and retail strategies in determining food prices."
 format: pdf
@@
 
-Overview paragraph
+The pricing of grocery products, particularly beef, plays a significant role in understanding consumer behavior and market dynamics within the Canadian retail sector.
+Beef, as a staple of many households, exhibits complex pricing behaviors influenced by a variety of factors, including vendor strategies, historical pricing, and seasonal trends.
+The broader context for this study lies in exploring how large-scale retailers and specialty stores adjust their beef prices over time and how these changes reflect broader economic and cultural trends.
+Given the increasing concerns around food affordability and accessibility, understanding the dynamics of beef pricing is not only crucial for consumers but also for stakeholders involved in food distribution and policy-making.
 
-Estimand paragraph
+This paper specifically focuses on two prominent vendors, Walmart and T&T, to examine how beef pricing varies across different types of retail environments.
+Walmart, as a major multinational retailer, represents mainstream, standardized pricing strategies, while T&T, catering primarily to Asian communities, represents a more specialized market segment.
+By analyzing data from these two vendors, this paper aims to address an important gap in understanding how cultural preferences and retail strategies influence product pricing.
+While previous studies have highlighted general trends in food pricing, they often overlook the nuanced differences between general retail chains and culturally specific stores.
+This paper fills that gap by providing an in-depth analysis of beef pricing patterns across these two distinct retail settings.
 
-Results paragraph
+Using a Bayesian linear regression approach, this study examines the effect of historical pricing, vendor-specific strategies, and seasonal factors on current beef prices.
+The analysis finds that old prices significantly predict current pricing, indicating a high degree of price consistency.
+Additionally, notable differences in pricing strategies between Walmart and T&T suggest that these vendors approach their markets with distinct pricing objectives, with Walmart favoring price stability and T&T showing more variability in pricing, possibly to cater to niche preferences.
+The findings are significant as they reveal the underlying mechanisms of price determination in a diverse retail environment, highlighting both market-wide factors and vendor-specific influences.
+Such insights could be useful for retailers looking to optimize their pricing strategies and for policymakers interested in addressing food affordability concerns.
 
-Why it matters paragraph
-
-Telegraphing paragraph: The remainder of this paper is structured as follows. @sec-data....
+The structure of the paper is organized as follows: @sec-data outlines the data collection and cleaning process, along with a description of the outcome and predictor variables used in the analysis.
+@sec-model introduces the forecasting models and discusses the rationale behind choosing these models for beef pricing prediction.
+@sec-result then presents the main findings, including a breakdown of vendor-level pricing effects and seasonal trends.
+Finally, @sec-discussion interprets the results, highlighting significant trends, comparing vendor strategies, and discussing potential limitations of the analysis.
 

### 2024-11-26 paper/paper.qmd
citations added, cross-references fixed

@@
 Data used in this paper was cleaned, analyzed and modeled with the programming language R [@citeR].
-Also with support of additional packages in R: `readr` [@citeReadr], `ggplot2` [@citeGgplot2], `tidyverse` [@citeTidyverse], `dplyr` [@citeDplyr], `here` [@citeHere], `knitr` [@citeKnitr], `kableExtra` [@citeKableExtra], `rstanarm` [@citeRstanarm], `modelsummary` [@citeModelsummary],
+Also with support of additional packages in R: `readr` [@citeReadr], `ggplot2` [@citeGgplot2], `tidyverse` [@citeTidyverse], `dplyr` [@citeDplyr], `here` [@citeHere], `knitr` [@citeKnitr], `kableExtra` [@citeKableExtra], `rstanarm` [@citeRstanarm], `modelsummary` [@citeModelsummary], `lme4` [@citeLme4], `tinytex` [@citeTinytex], `reshape2` [@citeReshape2].
 
@@
 
-# Model
+# Model {#sec-model}
 
@@
 
-We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
+We run the model in R [@citeR] using the `rstanarm` package of @citeRstanarm. We use the default priors from `rstanarm`.
 
@@
 
-# Results
+# Results {#sec-result}
 
@@
 
-# Discussion
+# Discussion {#sec-discussion}
 

### 2024-11-26 paper/paper.qmd
titles added

@@
 ---
-title: "My title"
-subtitle: "My subtitle if needed"
+title: "Beef Pricing Dynamics in the Canadian Grocery Market: A Comparative Analysis of Walmart and T&T"
+subtitle: "Evidence of Cultural and Strategic Influences on Beef Pricing"
 author: 

### 2024-11-28 scripts/03-clean_data.R
scripts edited for .parquet data

@@ -11,6 +11,7 @@
 
 #### Workspace setup ####
 library(tidyverse)
+library(arrow)
 library(here)
 
 # load data
@@ -66,5 +67,5 @@ cleaned_data <- merge_data %>%
 
 
 #### Save data ####
-write_csv(ppu_data, here("data", "02-analysis_data", "ppu_data.csv"))
-write_csv(cleaned_data, here("data", "02-analysis_data", "beef_data.csv"))
+write_parquet(ppu_data, here("data", "02-analysis_data", "ppu_data.parquet"))
+write_parquet(cleaned_data, here("data", "02-analysis_data", "beef_data.parquet"))

### 2024-11-28 scripts/05-exploratory_data_analysis.R

@@ -13,10 +13,11 @@
 # load library
 library(tidyverse)
 library(bayesplot)
+library(arrow)
 library(here)
 
 # load data
-beef_data <- read_csv(here("data", "02-analysis_data", "beef_data.csv"), show_col_types = FALSE)
+beef_data <- read_parquet(here("data", "02-analysis_data", "beef_data.parquet"))
 
 # load model
 beef_model <- readRDS (file = here:: here ("models/beef_model.rds"))

### 2024-11-28 scripts/06-model_data.R

@@ -13,10 +13,11 @@
 # load library
 library(tidyverse)
 library(rstanarm)
+library(arrow)
 library(here)
 
 # laod data
-beef_data <- read_csv(here("data", "02-analysis_data", "beef_data.csv"))
+beef_data <- read_parquet(here("data", "02-analysis_data", "beef_data.parquet"))
 
 
 

### 2024-11-28 paper/paper.qmd
parquet data modified, citation updated

@@
 # load library
-package_list <- c("tidyverse", "kableExtra", "ggplot2", "dplyr", "here", "knitr", "rstanarm", "modelsummary", "readr", "lme4", "tinytex", "reshape2")
+package_list <- c("tidyverse", "kableExtra", "ggplot2", "dplyr", "here", "knitr", "rstanarm", "modelsummary", "readr", "lme4", "tinytex", "reshape2", "arrow")
 
@@
 # load data
-ppu_data <- read_csv(here("data", "02-analysis_data", "ppu_data.csv"), show_col_types = FALSE)
-beef_data <- read_csv(here("data", "02-analysis_data", "beef_data.csv"), show_col_types = FALSE)
+ppu_data <- read_parquet(here("data", "02-analysis_data", "ppu_data.parquet"))
+beef_data <- read_parquet(here("data", "02-analysis_data", "beef_data.parquet"))
 
@@
 Data used in this paper was cleaned, analyzed and modeled with the programming language R [@citeR].
-Also with support of additional packages in R: `readr` [@citeReadr], `ggplot2` [@citeGgplot2], `tidyverse` [@citeTidyverse], `dplyr` [@citeDplyr], `here` [@citeHere], `knitr` [@citeKnitr], `kableExtra` [@citeKableExtra], `rstanarm` [@citeRstanarm], `modelsummary` [@citeModelsummary], `lme4` [@citeLme4], `tinytex` [@citeTinytex], `reshape2` [@citeReshape2].
+Also with support of additional packages in R: `readr` [@citeReadr], `ggplot2` [@citeGgplot2], `tidyverse` [@citeTidyverse], `dplyr` [@citeDplyr], `here` [@citeHere], `knitr` [@citeKnitr], `kableExtra` [@citeKableExtra], `rstanarm` [@citeRstanarm], `modelsummary` [@citeModelsummary], `lme4` [@citeLme4], `tinytex` [@citeTinytex], `reshape2` [@citeReshape2], `arrow` [@citeArrow].
 

### 2024-11-28 scripts/02-download_data.R
script download data added

@@ -4,23 +4,57 @@
 # Date: 11 February 2023 [...UPDATE THIS...]
 # Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites: package must be installed
+
 
 
 #### Workspace setup ####
-library(opendatatoronto)
-library(tidyverse)
-# [...UPDATE THIS...]
+# load library
+library(httr)
+library(here)
+library(utils)
+
+
 
 #### Download data ####
-# [...ADD CODE HERE TO DOWNLOAD...]
+# Specify the URL of the zip file
+url <- "https://jacobfilipp.com/hammerdata/hammer-5-csv.zip"
 
+# Specify the local file path for the zip file using `here`
+zip_file <- here("data", "01-raw_data", "raw.zip")
 
+# Create the directory structure if it doesn't exist
+dir.create(dirname(zip_file), recursive = TRUE, showWarnings = FALSE)
 
-#### Save data ####
-# [...UPDATE THIS...]
-# change the_raw_data to whatever name you assigned when you downloaded it.
-write_csv(the_raw_data, "inputs/data/raw_data.csv") 
+# Download the zip file
+response <- GET(url, write_disk(zip_file, overwrite = TRUE))
+
+# Check the status of the download
+if (status_code(response) == 200) {
+  message("Zip file downloaded successfully and saved as: ", zip_file)
+  
+  # Extract the zip file directly into the `01-raw_data` folder
+  unzip_dir <- here("data", "01-raw_data")
+  unzip(zip_file, exdir = unzip_dir)
+  message("Zip file extracted directly to the directory: ", unzip_dir)
+  
+  # List the extracted files
+  extracted_files <- list.files(unzip_dir, full.names = TRUE)
+  print("Extracted files:")
+  print(extracted_files)
+  
+  # Optional: Load and preview the first CSV file
+  csv_file <- extracted_files[grep("\\.csv$", extracted_files)][1] # Select the first CSV
+  if (!is.na(csv_file)) {
+    dataset <- read.csv(csv_file)
+    print(head(dataset))
+  } else {
+    message("No CSV files found in the extracted content.")
+  }
+  
+} else {
+  message("Failed to download the zip file. Status code: ", status_code(response))
+}
+
 
          

### 2024-11-28 scripts/00-simulate_data.R
scripts preamble edited

@@ -1,12 +1,11 @@
 #### Preamble ####
-# Purpose: Simulates a dataset of Australian electoral divisions, including the 
-  #state and party that won each division.
-# Author: Rohan Alexander
-# Date: 26 September 2024
-# Contact: rohan.alexander@utoronto.ca
+# Purpose: Simulate a dataset of Australian electoral divisions with associated state and party affiliations for data analysis and testing purposes.
+# Author: Ziheng Zhong
+# Date: 20 November 2024
+# Contact: ziheng.zhong@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: The `tidyverse` package must be installed
-# Any other information needed? Make sure you are in the `starter_folder` rproj
+# Pre-requisites: 
+  # - Package must be installed
 
 
 

### 2024-11-28 scripts/01-test_simulated_data.R

@@ -1,12 +1,10 @@
 #### Preamble ####
-# Purpose: Tests the structure and validity of the simulated Australian 
-  #electoral divisions dataset.
-# Author: Rohan Alexander
-# Date: 26 September 2024
-# Contact: rohan.alexander@utoronto.ca
+# Purpose: To validate the integrity and consistency of the simulated beef dataset through a series of unit tests, ensuring accurate data loading, logical constraints, and adherence to expected value ranges.
+# Author: Ziheng Zhong
+# Date: 20 November 2024
+# Contact: ziheng.zhong@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: 
-  # - The `tidyverse` package must be installed and loaded
   # - 00-simulate_data.R must have been run
 
 

### 2024-11-28 scripts/02-download_data.R

@@ -1,10 +1,11 @@
 #### Preamble ####
-# Purpose: Downloads and saves the data from [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: To download a zip file containing raw data from a specified URL, extract its contents into a structured directory, and optionally load and preview the first CSV file to ensure successful extraction and data integrity.
+# Author: Ziheng Zhong
+# Date: 20 November 2024
+# Contact: ziheng.zhong@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: package must be installed
+# Pre-requisites: 
+  # - Package must be installed
 
 
 
@@ -55,6 +56,5 @@ if (status_code(response) == 200) {
 } else {
   message("Failed to download the zip file. Status code: ", status_code(response))
 }
-
 
          

### 2024-11-28 scripts/03-clean_data.R

@@ -1,11 +1,12 @@
 #### Preamble ####
-# Purpose: Cleans the raw plane data recorded by two observers..... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 6 April 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: To load raw product and transaction data, merge and clean the datasets by filtering and transforming columns, and save the cleaned and processed data into parquet files for further analysis.
+# Author: Ziheng Zhong
+# Date: 20 November 2024
+# Contact: ziheng.zhong@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites: 
+  # - Package must be installed
+  # - 02-download_data.R must have been run
 
 
 
@@ -69,3 +70,5 @@ cleaned_data <- merge_data %>%
 #### Save data ####
 write_parquet(ppu_data, here("data", "02-analysis_data", "ppu_data.parquet"))
 write_parquet(cleaned_data, here("data", "02-analysis_data", "beef_data.parquet"))
+
+

### 2024-11-28 scripts/04-test_analysis_data.R

@@ -1,11 +1,12 @@
 #### Preamble ####
-# Purpose: Tests... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 26 September 2024 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: To validate the integrity of the cleaned beef dataset by testing for successful data loading, logical consistency of column values, adherence to specified constraints, and correct calculations for derived metrics.
+# Author: Ziheng Zhong
+# Date: 20 November 2024
+# Contact: ziheng.zhong@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites: 
+  # - Package must be installed
+  # - 03-clean_data.R must have been run
 
 
 

### 2024-11-28 scripts/05-exploratory_data_analysis.R

@@ -1,11 +1,13 @@
 #### Preamble ####
-# Purpose: Models and visualizes beef price trends from a dataset.
+# Purpose: To explore and visualize trends in beef pricing data, this script includes analyses of price distributions, vendor-based differences, temporal trends, and price relationships, while leveraging a Bayesian model to assess data patterns and perform posterior predictive checks.
 # Author: Ziheng Zhong
-# Date: 22 November 2024
-# Contact: ziheng.zhong@example.com
+# Date: 20 November 2024
+# Contact: ziheng.zhong@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: Ensure the 'beef_data.csv' file is available in the appropriate directory.
-# Any other information needed? This script creates three different visualizations for analysis.
+# Pre-requisites: 
+  # - Package must be installed
+  # - 03-clean_data.R must have been run
+  # - 06-model_data.R must have been run
 
 
 

### 2024-11-28 scripts/06-model_data.R

@@ -1,11 +1,12 @@
 #### Preamble ####
-# Purpose: Models... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: To build and save a Bayesian regression model that predicts the current price of beef based on month, old price, and vendor, using the `rstanarm` package with specified priors and a Gaussian family.
+# Author: Ziheng Zhong
+# Date: 20 November 2024
+# Contact: ziheng.zhong@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites: 
+  # - Package must be installed
+  # - 03-clean_data.R must have been run
 
 
 

### 2024-11-28 paper/paper.qmd
charts formatted, appendix formatted

@@
   - Ziheng Zhong
-thanks: "Code and data are available at: [https://github.com/iJustinn/Canadian_Grocery_Analysis.git](https://github.com/iJustinn/Canadian_Grocery_Analysis.git)."
+thanks: "Code and data supporting this analysis is available at: [https://github.com/iJustinn/Canadian_Grocery_Analysis.git](https://github.com/iJustinn/Canadian_Grocery_Analysis.git)."
 date: today
@@
 This paper fills that gap by providing an in-depth analysis of beef pricing patterns across these two distinct retail settings.
+
+The estimand of this study is the current price of beef, determined by factors such as vendor type, historical pricing, etc.
+By estimating how these variables impact the current pricing, the paper aims to provide a comprehensive understanding of the pricing mechanisms that affect retail beef prices in Canada.
 
@@
 #| message: false
+#| fig-width: 5
 
@@
   geom_boxplot(fill = "lightgrey", color = "darkgrey", alpha = 0.7) +
-  labs(title = "Boxplot of the Distribution of Current Prices", y = "Current Price (in $)") +
-  theme_minimal()
+  labs(y = "Current Price (in $)") +
+  theme_minimal() +
+  theme(axis.text.x = element_blank(),
+        axis.ticks.x = element_blank())
 
@@
 
-## Model overview
+## Overview
 
 To model the current price of beef, $y_i$, at time $i$, we use a Bayesian linear regression model implemented with the `stan_glm` function in the R package `rstanarm`.
-The response variable, $y_i$, represents the **current price of beef** in dollars.
-Our predictors include $x_1$, $x_2$, and $x_3$, which represent the **month**, **old price**, and **vendor**, respectively.
+The response variable, $y_i$, represents the current price of beef in dollars.
+Our predictors include $x_1$, $x_2$, and $x_3$, which represent the month, old price, and vendor, respectively.
 Each component of the model is defined and justified below.
 
-## Model set-up
+## Set-up
 
@@
 In this model:
-- $\alpha$ represents the **intercept** term, capturing the baseline price of beef.
-- $\beta_1$ corresponds to the **effect of the month** on the current price, allowing us to capture **seasonal variations** that may impact beef pricing.
+- $\alpha$ represents the intercept term, capturing the baseline price of beef.
+- $\beta_1$ corresponds to the effect of the month on the current price, allowing us to capture seasonal variations that may impact beef pricing.
 The month is treated as a categorical variable, acknowledging that different months can bring different demand or supply effects due to holidays, weather, or other seasonal factors.
-- $\beta_2$ represents the **effect of the old price** on the current price. This allows us to account for price inertia or trends where past pricing influences current pricing.
-- $\beta_3$ represents the **effect of the vendor**, also treated as a categorical variable.
+- $\beta_2$ represents the effect of the old price on the current price. This allows us to account for price inertia or trends where past pricing influences current pricing.
+- $\beta_3$ represents the effect of the vendor, also treated as a categorical variable.
 Different vendors may have distinct pricing strategies, and including vendor as a categorical predictor helps us capture this vendor-specific pricing variation.
@@
 
-These priors are chosen to be **weakly informative**, allowing the data to speak for itself while still providing a reasonable range for the coefficients based on prior expectations.
+These priors are chosen to be weakly informative, allowing the data to speak for itself while still providing a reasonable range for the coefficients based on prior expectations.
 Specifically, a standard deviation of 2.5 reflects our expectation that most reasonable effects should fall within a plausible range without being overly restrictive.
 
-For the residual standard deviation, $\sigma$, we assume an **Exponential(1)** prior:
+For the residual standard deviation, $\sigma$, we assume an Exponential(1) prior:
 
@@
 
-## Model justification
+## Justification
 
 The choice of predictors is well-balanced in terms of complexity and appropriateness for this scenario.
-By including **month** and **vendor** as categorical variables, we acknowledge the significant impact that both seasonal changes and vendor-specific factors can have on the current price of beef.
-The inclusion of the **old price** reflects the idea of price inertia, where historical prices are likely to influence current pricing trends.
+By including month and vendor as categorical variables, we acknowledge the significant impact that both seasonal changes and vendor-specific factors can have on the current price of beef.
+The inclusion of the old price reflects the idea of price inertia, where historical prices are likely to influence current pricing trends.
 
-The model is neither **overly simplistic** nor **unnecessarily complex**: it captures important predictors without adding extraneous complexity that could lead to overfitting.
-**Month** and **vendor** are treated as categorical variables rather than grouping by more arbitrary categories, which maintains the expressiveness of our model while being parsimonious.
+The model is neither overly simplistic nor unnecessarily complex: it captures important predictors without adding extraneous complexity that could lead to overfitting.
+Month and vendor are treated as categorical variables rather than grouping by more arbitrary categories, which maintains the expressiveness of our model while being parsimonious.
 
 The model is implemented using `stan_glm` from the `rstanarm` package, a high-level interface to Stan for Bayesian modeling.
-**Stan** provides powerful sampling algorithms, making it an appropriate tool for fitting our Bayesian model efficiently, ensuring **robust convergence** and accurate posterior estimation.
+Stan provides powerful sampling algorithms, making it an appropriate tool for fitting our Bayesian model efficiently, ensuring robust convergence and accurate posterior estimation.
 
-### Model Assumptions and Limitations
+### Assumptions and Limitations
 - **Linearity**: The model assumes a linear relationship between the predictors and the response variable. This may be a limitation if the true relationship is nonlinear.
@@
 
-### Model Validation and Diagnostics
+### Validation and Diagnostics
 - **Posterior Predictive Checks**: We performed posterior predictive checks to verify that the model adequately captures the variation in the data.
 These checks involve comparing simulated data from the posterior distribution to the observed data to check for any discrepancies.
-- **Convergence**: Model convergence was assessed using **trace plots** and the **R-hat statistic**, with all values being close to 1, indicating successful convergence.
+- **Convergence**: Model convergence was assessed using trace plots and the R-hat statistic, with all values being close to 1, indicating successful convergence.
 - **Alternative Models**: We also considered fitting simpler linear models that did not include the vendor or seasonal components.
@@
 
-## Charts
+## Charting
 
@@
   geom_histogram(binwidth = 1, fill = "lightgrey", color = "darkgrey", alpha = 0.7) +
-  geom_density(aes(y = ..count..), color = "red", size = 1) +
+  geom_density(aes(y = ..count..), color = "red", size = 0.8) +
   labs(x = "Current Price (in $)", y = "Frequency") +
@@
 ggplot(avg_price_per_date, aes(x = date, y = avg_current_price)) +
-  geom_line(color = "darkgrey", size = 1) +
+  geom_point(color = "grey", size = 1.5, alpha = 0.7) +
+  geom_smooth(method = "loess", color = "red", fill = "darkgrey", alpha = 0.3, se = TRUE) +
   labs(x = "Date", y = "Average Current Price (in $)") +
@@
 ggplot(beef_data, aes(x = price_per_unit)) +
-  geom_histogram(binwidth = 0.1, fill = "lightgrey", color = "darkgrey", alpha = 0.7) +
-  geom_density(aes(y = after_stat(count)), color = "red", linewidth = 1) +
-  labs(x = "Unit Price ($ per 100 grams)", y = "Frequency") +
+  geom_density(fill = "lightgrey", color = "red", , linewidth = 0.8, alpha = 0.5, bw = 0.5) +
+  labs(x = "Unit Price ($ per 100 grams)", y = "Density") +
   theme_minimal()
@@
 
-## Model
+## Modeling
 
@@
 
+\newpage
+
 ## Price per unit
 
+In cleaning the `price_per_unit` variable, I began by extracting the numerical value using `str_extract()`, which allowed me to identify the dollar amounts in the original string (`\\$[0-9\\.]+`), as shown in @tbl-price_per_unit-1.
+This step ensured that only valid price values, starting with a dollar sign, were captured, removing any extraneous text or symbols.
+
+Next, I removed the dollar sign itself with `str_remove("\\$")` to convert the extracted value into a format suitable for numerical operations.
+After this, I used `as.numeric()` to convert the string into a numeric type, allowing for further quantitative analysis.
+
+To handle missing values, I applied `ifelse(is.na(price_per_unit), 0, price_per_unit)`, which replaced any `NA` values in the `price_per_unit` column with `0`.
+This step ensured that all rows had a value for `price_per_unit`, preventing issues during subsequent analysis.
+Using zero as the replacement helped maintain continuity in the dataset, although I acknowledge that this choice could imply that missing values are negligible, which might require additional context or consideration during interpretation.
+The price_per_unit column in the dataset after cleaning looks like @tbl-price_per_unit-2.
+
 ```{r}
-#| label: tbl-raw_price_per_unit
+#| label: tbl-price_per_unit
 #| tbl-cap: "Price per Unit Data for TandT and Walmart"
@@
 
-In cleaning the `price_per_unit` variable, I began by extracting the numerical value using `str_extract()`, which allowed me to identify the dollar amounts in the original string (`\\$[0-9\\.]+`), as shown in @tbl-price_per_unit-1.
-This step ensured that only valid price values, starting with a dollar sign, were captured, removing any extraneous text or symbols.
-
-Next, I removed the dollar sign itself with `str_remove("\\$")` to convert the extracted value into a format suitable for numerical operations.
-After this, I used `as.numeric()` to convert the string into a numeric type, allowing for further quantitative analysis.
-
-To handle missing values, I applied `ifelse(is.na(price_per_unit), 0, price_per_unit)`, which replaced any `NA` values in the `price_per_unit` column with `0`.
-This step ensured that all rows had a value for `price_per_unit`, preventing issues during subsequent analysis.
-Using zero as the replacement helped maintain continuity in the dataset, although I acknowledge that this choice could imply that missing values are negligible, which might require additional context or consideration during interpretation.
-The price_per_unit column in the dataset after cleaning looks like @tbl-price_per_unit-2.
+\newpage
 
@@
 ## Posterior predictive check
-In @fig-ppcheckandposteriorvsprior-1, we implement a posterior predictive check.
+In @fig-ppcheck, we implement a posterior predictive check.
 This plot shows the observed data distribution (in dark lines) compared with simulated datasets from the posterior (in light grey lines).
@@
 
-In @fig-ppcheckandposteriorvsprior-2, we compare the posterior with the prior distributions for each parameter.
+```{r}
+#| eval: true
+#| echo: false
+#| message: false
+#| warning: false
+#| label: fig-ppcheck
+#| fig-cap: "Posterior prediction check result"
+
+# Posterior predictive check
+pp_check(beef_model, plotfun = "dens_overlay")
+
+```
+
+\newpage
+
+In @fig-posteriorvsprior, we compare the posterior with the prior distributions for each parameter.
 The posterior distribution (shown on the left) is more concentrated compared to the prior (shown on the right), which demonstrates how the data has influenced and updated our beliefs about the model parameters.
@@
 #| warning: false
-#| label: fig-ppcheckandposteriorvsprior
-#| layout-ncol: 2
-#| fig-cap: "Examining how the model fits, and is affected by, the data"
-#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]
-
-# Posterior predictive check
-pp_check(beef_model, plotfun = "dens_overlay")
+#| label: fig-posteriorvsprior
+#| fig-cap: "Comparing the posterior with the prior"
 
 # Comparing posterior with prior
-posterior_vs_prior(beef_model, color_by = "parameter") + theme_minimal() + scale_fill_grey() + scale_color_grey()
+posterior_vs_prior(beef_model, color_by = "parameter") +
+  theme_minimal() +
+  scale_fill_grey() +
+  scale_color_grey() +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+
 ```
 
+\newpage
+
 ## Diagnostics
-In @fig-stanareyouokay-1, we present a trace plot.
+In @fig-trace, we present a trace plot.
 This plot shows the sampled values for each parameter across the MCMC iterations.
@@
 
-In @fig-stanareyouokay-2, we present an Rhat plot.
+```{r}
+#| echo: false
+#| eval: true
+#| message: false
+#| warning: false
+#| label: fig-trace
+#| fig-cap: "Trace plot"
+
+# Trace plot
+plot(beef_model, plotfun = "trace") + theme_minimal()
+
+```
+
+\newpage
+
+In @fig-rhat, we present an Rhat plot.
 The Rhat statistic (also known as the potential scale reduction factor) measures the ratio of the variance between chains to the variance within chains, and values close to 1 indicate that the chains have converged.
@@
 #| warning: false
-#| label: fig-stanareyouokay
-#| fig-cap: "Checking the convergence of the MCMC algorithm"
-#| fig-subcap: ["Trace plot", "Rhat plot"]
-#| layout-ncol: 2
-
-# Trace plot
-plot(beef_model, plotfun = "trace") + theme_minimal()
+#| label: fig-rhat
+#| fig-cap: "Rhat plot"
 
@@
 plot(beef_model, plotfun = "rhat") + theme_minimal()
+
 ```

### 2024-11-28 paper/paper.qmd
charts caption revised

@@
 #| label: fig-average_price_over_time
-#| fig-cap: "Average Current Price Over Time"
+#| fig-cap: "Average Daily Current Price Over Time"
 #| echo: false
@@
 #| label: fig-current_vs_old_price
-#| fig-cap: "Heatmap of Price Levels between Current Price and Old Price of Beef"
+#| fig-cap: "Relationship Between Old and Current Beef Price Levels"
 #| echo: false
@@
 #| label: tbl-modelresults
-#| tbl-cap: ""
+#| tbl-cap: "Beef model results"
 #| echo: false

### 2024-11-28 paper/paper.qmd
literature referencing added

@@
   - Ziheng Zhong
-thanks: "Code and data supporting this analysis is available at: [https://github.com/iJustinn/Canadian_Grocery_Analysis.git](https://github.com/iJustinn/Canadian_Grocery_Analysis.git)."
+thanks: "Code and data supporting this analysis is available at: [Link to repository](https://github.com/iJustinn/Canadian_Grocery_Analysis.git)."
 date: today
@@
 The pricing of grocery products, particularly beef, plays a significant role in understanding consumer behavior and market dynamics within the Canadian retail sector.
-Beef, as a staple of many households, exhibits complex pricing behaviors influenced by a variety of factors, including vendor strategies, historical pricing, and seasonal trends.
+Beef, as a staple of many households, exhibits complex pricing behaviors influenced by a variety of factors, including vendor strategies [@citePricingStrategies], historical pricing, and seasonal trends [@citeSeasonal].
 The broader context for this study lies in exploring how large-scale retailers and specialty stores adjust their beef prices over time and how these changes reflect broader economic and cultural trends.
@@
 This paper specifically focuses on two prominent vendors, Walmart and T&T, to examine how beef pricing varies across different types of retail environments.
-Walmart, as a major multinational retailer, represents mainstream, standardized pricing strategies, while T&T, catering primarily to Asian communities, represents a more specialized market segment.
+Walmart, as a major multinational retailer, represents mainstream [@citeWalmart], standardized pricing strategies, while T&T, catering primarily to Asian communities, represents a more specialized market segment [@citeTandT].
 By analyzing data from these two vendors, this paper aims to address an important gap in understanding how cultural preferences and retail strategies influence product pricing.
@@
 One critical point of discussion is the apparent distinction in pricing strategies between Walmart and T&T.
-Walmart's pricing, as evidenced by the relatively narrow spread of the boxplot, seems to prioritize consistency and predictability, possibly reflecting its brand strategy of offering everyday low prices.
+Walmart's pricing, as evidenced by the relatively narrow spread of the boxplot, seems to prioritize consistency and predictability, possibly reflecting its brand strategy of offering everyday low prices [@citeWalmartLowPrice].
 This approach may be designed to build consumer trust, ensuring customers that they are receiving competitive pricing without frequent fluctuations.
-On the other hand, T&T's wider range of price differences may reflect a more dynamic pricing strategy, potentially influenced by the specialty nature of its products, diverse supply chain, or its responsiveness to a different customer demographic.
+On the other hand, T&T's wider range of price differences may reflect a more dynamic pricing strategy, potentially influenced by the specialty nature of its products [@citeTandT], diverse supply chain, or its responsiveness to a different customer demographic.
 Such variability in pricing could indicate promotional efforts, targeted price differentiation, or a reflection of its broader range of product quality and types.
@@
 Additionally, the moderate decrease in prices observed during certain months might suggest seasonal promotions or shifts in consumer demand that lead vendors to adjust prices accordingly.
-Such trends could be driven by factors like increased competition, holiday promotions, or seasonal variations in meat consumption.
+Such trends could be driven by factors like increased competition, holiday promotions [@citeHoliday], or seasonal variations in meat consumption.
 
@@
 This could also point to the influence of retailer pricing power, where large players like Walmart are able to leverage economies of scale to keep prices low, thereby shaping overall market dynamics.
-An important next step in understanding these patterns would be to assess how external shocks, such as economic downturns or supply chain disruptions, influence this pricing stability, as well as whether similar patterns hold for other product categories.
+An important next step in understanding these patterns would be to assess how external shocks, such as economic downturns or supply chain disruptions [@citeSupplyChain], influence this pricing stability, as well as whether similar patterns hold for other product categories.
 
@@
 First, the dataset used was limited to two vendors, Walmart and T&T, which may not fully represent the diversity of the Canadian grocery market.
-Including more vendors could provide a broader perspective on pricing strategies across different types of retailers, such as discount chains, premium grocers, or independent stores.
-Additionally, the current analysis did not account for factors like promotions, product quality differences, or customer loyalty programs, which could play significant roles in pricing decisions.
+Including more vendors could provide a broader perspective on pricing strategies across different types of retailers, such as discount chains, premium grocers, or independent stores [@citeBrand].
+Additionally, the current analysis did not account for factors like promotions, product quality differences, or customer loyalty programs [@citeProgram], which could play significant roles in pricing decisions.
 Including such variables in future models could provide a more nuanced understanding of the factors influencing beef prices.
@@
 
+# Idealized Methodology for a Survey on Beef Pricing {#sec-survey}
+
+## Overview
+
+To complement the quantitative analysis presented in this paper, an idealized survey methodology could provide direct insights into consumer behavior and perceptions regarding beef pricing at Walmart and T&T.
+This survey would aim to identify factors influencing purchasing decisions, perceptions of price fairness, and preferences for vendor-specific attributes, offering a holistic understanding of the dynamics underpinning beef pricing in the Canadian grocery market.
+
+## Sampling Approach
+
+A **stratified random sampling** method would be employed to ensure representation across key demographic groups, including age, income levels, and cultural background.
+Stratification variables would include:
+
+1. **Region**: Urban versus rural shoppers across Canada.
+2. **Vendor familiarity**: Regular Walmart shoppers, regular T&T shoppers, and those who shop at both.
+3. **Income brackets**: Low, middle, and high-income households.
+
+A sample size of approximately 1,200 respondents would be targeted, with 400 participants in each stratum (Walmart-focused, T&T-focused, and dual shoppers).
+This stratification would allow for meaningful comparisons between vendor-specific perceptions and strategies.
+
+## Survey Structure
+
+The survey would include a mix of **closed-ended** and **open-ended** questions.
+
+### Question Types
+
+1. **Demographics and Shopping Habits**
+   - Frequency of grocery shopping (weekly, bi-weekly, etc.).
+   - Preferred vendor for beef purchases and reasons for preference.
+   - Average budget allocated for beef purchases.
+
+2. **Price Perception**
+   - Rating the perceived fairness of beef prices on a Likert scale.
+   - Comparison of beef prices at Walmart and T&T (e.g., cheaper, equivalent, more expensive).
+   - Sensitivity to price changes (e.g., a 10% increase in beef prices would lead to…).
+
+3. **Cultural and Quality Preferences**
+   - Importance of cultural alignment in beef products (specific cuts, preparation styles, etc.).
+   - Perceived quality differences between Walmart and T&T beef products.
+   - Willingness to pay a premium for culturally tailored products.
+
+4. **External Factors**
+   - Awareness of seasonal price trends.
+   - Impact of promotions on purchasing decisions.
+   - Responses to supply chain disruptions (e.g., alternative purchasing strategies).
+
+### Question List
+
+1. How often do you shop for groceries?
+   - Weekly
+   - Bi-weekly
+   - Monthly
+   - Less often
+
+2. Which retailer do you prefer for buying beef products?
+   - Walmart
+   - T&T
+   - Both
+
+3. What factors influence your choice of retailer for beef purchases? (Select all that apply)
+   - Price
+   - Quality
+   - Availability of specific cuts
+   - Cultural preferences
+   - Convenience
+   - Promotions
+
+4. On average, how much do you spend on beef per shopping trip?
+   - Less than $20
+   - \$20 - \$50
+   - \$50 - \$100
+   - More than $100
+
+5. How fair do you find the current beef prices at Walmart/T&T?
+   - Very unfair
+   - Somewhat unfair
+   - Neutral
+   - Somewhat fair
+   - Very fair
+
+6. How would you rate the quality of beef at Walmart compared to T&T?
+   - Much lower quality
+   - Slightly lower quality
+   - About the same quality
+   - Slightly higher quality
+   - Much higher quality
+
+7. If beef prices increased by 10%, would you consider switching to an alternative product or retailer?
+   - Definitely would switch
+   - Probably would switch
+   - Might switch
+   - Probably would not switch
+   - Definitely would not switch
+
+8. How important is it for you that beef products align with your cultural preferences (e.g., specific cuts, preparation styles)?
+   - Not important at all
+   - Slightly important
+   - Moderately important
+   - Very important
+   - Extremely important
+
+9. Are you willing to pay more for beef products that are culturally tailored to your preferences?
+   - Not willing at all
+   - Slightly willing
+   - Moderately willing
+   - Very willing
+   - Extremely willing
+
+10. How aware are you of seasonal price changes in beef products?
+    - Not aware at all
+    - Slightly aware
+    - Moderately aware
+    - Very aware
+    - Extremely aware
+
+11. How do promotions impact your decision to purchase beef?
+    - No impact
+    - Slight impact
+    - Moderate impact
+    - Significant impact
+    - Very significant impact
+
+12. How do you typically respond to supply chain disruptions affecting beef availability?
+    - Purchase alternative products
+    - Purchase from a different retailer
+    - Reduce beef consumption
+    - No change in purchasing behavior
+
+## Recruitment Strategy
+
+Participants would be recruited using **multichannel outreach**, including:
+
+1. **Online Panels**: Partnering with Canadian consumer research platforms to recruit participants from diverse regions.
+2. **Retailer Partnerships**: Collaborating with Walmart and T&T to survey in-store shoppers (e.g., QR codes on receipts leading to the survey).
+3. **Community Outreach**: Engaging with cultural organizations to reach underrepresented groups, particularly for T&T shoppers.
+
+## Linkage to Literature
+
+\newpage
+
 # References

### 2024-11-28 paper/paper.qmd
all text revised according to rubric word list

@@
 
-The pricing of grocery products, particularly beef, plays a significant role in understanding consumer behavior and market dynamics within the Canadian retail sector.
-Beef, as a staple of many households, exhibits complex pricing behaviors influenced by a variety of factors, including vendor strategies [@citePricingStrategies], historical pricing, and seasonal trends [@citeSeasonal].
-The broader context for this study lies in exploring how large-scale retailers and specialty stores adjust their beef prices over time and how these changes reflect broader economic and cultural trends.
-Given the increasing concerns around food affordability and accessibility, understanding the dynamics of beef pricing is not only crucial for consumers but also for stakeholders involved in food distribution and policy-making.
+The pricing of grocery items, particularly beef, plays an important role in understanding consumer behavior and market trends within the Canadian retail sector.
+As a staple in many households, beef pricing is shaped by various factors, including vendor strategies [@citePricingStrategies], historical trends, and seasonal variations [@citeSeasonal].
+This study focuses on examining how major retailers adjust beef prices over time and how these adjustments reflect broader economic and cultural patterns.
+With growing concerns around food affordability and access, analyzing beef pricing dynamics is essential for both consumers and stakeholders involved in food distribution and policy-making.
 
-This paper specifically focuses on two prominent vendors, Walmart and T&T, to examine how beef pricing varies across different types of retail environments.
-Walmart, as a major multinational retailer, represents mainstream [@citeWalmart], standardized pricing strategies, while T&T, catering primarily to Asian communities, represents a more specialized market segment [@citeTandT].
-By analyzing data from these two vendors, this paper aims to address an important gap in understanding how cultural preferences and retail strategies influence product pricing.
-While previous studies have highlighted general trends in food pricing, they often overlook the nuanced differences between general retail chains and culturally specific stores.
-This paper fills that gap by providing an in-depth analysis of beef pricing patterns across these two distinct retail settings.
+This paper centers on two significant vendors, Walmart and T&T, to investigate how beef pricing differs across retail environments.
+Walmart, a global retailer, exemplifies standardized pricing strategies [@citeWalmart], while T&T, catering primarily to Asian communities, represents a more targeted market segment [@citeTandT].
+By analyzing data from these vendors, the study addresses an often-overlooked aspect of how cultural preferences and retail strategies influence pricing.
+Previous research has focused on general trends in food pricing but has not fully considered the distinctions between mainstream retail chains and culturally specific stores.
+This analysis provides a detailed comparison of beef pricing patterns in these two diverse settings.
 
@@
 
-Using a Bayesian linear regression approach, this study examines the effect of historical pricing, vendor-specific strategies, and seasonal factors on current beef prices.
-The analysis finds that old prices significantly predict current pricing, indicating a high degree of price consistency.
-Additionally, notable differences in pricing strategies between Walmart and T&T suggest that these vendors approach their markets with distinct pricing objectives, with Walmart favoring price stability and T&T showing more variability in pricing, possibly to cater to niche preferences.
-The findings are significant as they reveal the underlying mechanisms of price determination in a diverse retail environment, highlighting both market-wide factors and vendor-specific influences.
-Such insights could be useful for retailers looking to optimize their pricing strategies and for policymakers interested in addressing food affordability concerns.
+Using a Bayesian linear regression model, the study examines how historical pricing, vendor-specific strategies, and seasonal factors impact current beef prices.
+The analysis indicates that past prices strongly predict current pricing, suggesting a high degree of consistency.
+Differences in pricing strategies between Walmart and T&T further highlight distinct market approaches, with Walmart focusing on stable pricing and T&T displaying greater variability, potentially to address specific consumer preferences.
+These results provide a clearer understanding of pricing strategies in a diverse retail landscape and offer insights into factors that shape affordability and market behavior.
 
-The structure of the paper is organized as follows: @sec-data outlines the data collection and cleaning process, along with a description of the outcome and predictor variables used in the analysis.
-@sec-model introduces the forecasting models and discusses the rationale behind choosing these models for beef pricing prediction.
-@sec-result then presents the main findings, including a breakdown of vendor-level pricing effects and seasonal trends.
-Finally, @sec-discussion interprets the results, highlighting significant trends, comparing vendor strategies, and discussing potential limitations of the analysis.
+The paper is structured as follows: @sec-data describes the data collection and cleaning methods, as well as the outcome and predictor variables used in the analysis.
+@sec-model introduces the forecasting models and explains their selection for predicting beef prices.
+@sec-result presents the key findings, including vendor-specific pricing effects and seasonal trends.
+Lastly, @sec-discussion interprets these results, comparing vendor strategies, identifying significant patterns, and discussing the study's limitations.
 
@@
 
-The dataset used for this research was sourced from Hammer, a publicly available repository designed to provide pricing data from various retail chains.
-The dataset comprises detailed information on product attributes, such as product name, vendor, current price, old price, units available, and additional details like the month of certain price.
-The dataset's broader context is centered on consumer pricing trends across different vendors, helping to understand how factors like vendor type, previous pricing, and seasonal changes influence current prices.
-This information is pivotal for examining market behaviors and assessing pricing strategies in retail environments.
+The dataset for this research was obtained from Hammer, a publicly accessible repository offering pricing information from a variety of retail chains.
+It includes details on product attributes such as name, vendor, current and old prices, available units, and the month associated with certain prices.
+The dataset focuses on consumer pricing trends across different vendors, enabling an analysis of how factors like vendor type, historical pricing, and seasonal changes affect current prices.
+This provides a basis for examining retail market behaviors and assessing vendor pricing strategies.
 
-The dataset variables include several categorical and numerical components.
-Key variables such as ‘vendor’, which identifies the retailer (e.g., Walmart, Galleria), and ‘product_name’, which details the item, are crucial for understanding distribution channels.
-Variables like ‘current_price’ and ‘old_price’ offer insights into pricing dynamics, enabling analyses of price changes over time.
-Additionally, temporal variables such as ‘month’ allow for the identification of seasonal variations in pricing.
-Summary statistics, along with graphs for each of these variables, help illustrate their distributions and relationships.
-Graphs have been included to show the frequency distribution for vendors, trends in price changes, and comparisons between old and current prices.
-Relationships among variables like ‘vendor’ and ‘current_price’ have been highlighted to give a comprehensive view of the dataset.
+The dataset contains both categorical and numerical variables.
+Important variables include ‘vendor,’ which identifies the retailer (e.g., Walmart, Galleria), and ‘product_name,’ which specifies the item being analyzed.
+Variables like ‘current_price’ and ‘old_price’ help track pricing changes, while temporal variables such as ‘month’ allow for the identification of seasonal pricing patterns.
+Summary statistics and visualizations have been generated to illustrate the distributions and relationships of these variables.
+These graphs include frequency distributions for vendors, trends in price changes, and comparisons between old and current prices.
+Relationships such as those between ‘vendor’ and ‘current_price’ are highlighted to present a detailed view of the dataset.
 
-There were other potential datasets available for this analysis, such as proprietary retail data sources or other consumer purchasing databases.
-However, they were not selected due to restrictions on data availability, licensing requirements, or lack of detail in specific product-level attributes.
-The Hammer dataset was chosen as it provides granular pricing data that is crucial for understanding product-level trends across multiple vendors.
-The dataset also allowed for the construction of new variables, such as the calculated ‘price_per_unit’, which enabled more meaningful comparisons between products.
-High-level data cleaning included handling missing values in the ‘old_price’ variable by replacing them with the ‘current_price’, ensuring the analysis was consistent and complete.
+Alternative datasets, such as proprietary retail databases or other consumer purchasing records, were considered for this analysis.
+However, these were not selected due to limitations in accessibility, licensing restrictions, or insufficient detail in product-level attributes.
+The Hammer dataset was chosen because it provides detailed pricing data necessary for analyzing product-level trends across various vendors.
+It also enabled the creation of derived variables, such as ‘price_per_unit,’ which facilitated more meaningful comparisons between products.
+Data cleaning efforts included addressing missing values in the ‘old_price’ variable by replacing them with the ‘current_price,’ ensuring consistency and completeness in the analysis.
 
@@
 
-The process of measurement in this research involves translating real-world phenomena into quantifiable data entries within the dataset.
-For instance, consumer purchasing behavior, influenced by factors such as vendor type, pricing history, and seasonal changes, has been captured through variables like ‘vendor’, ‘current_price’, and ‘month’.
-The ‘vendor’ variable serves as an identifier of the source of the product, which helps to provide insights into how different retail environments may affect pricing strategies.
-The ‘current_price’ and ‘old_price’ variables reflect pricing trends and allow us to measure the temporal changes in consumer costs, thereby linking economic activities to specific data points.
+The data for Project Hammer, focused on historical grocery prices, does not explicitly describe its collection methods.
+However, automated web scraping is a feasible approach for gathering this information.
+This method extracts structured information from retailer websites, capturing details such as product names, prices, brands, and unit measurements.
+The process involves developing scripts or utilizing tools that navigate websites, identify relevant HTML elements, and store the data in structured formats like CSV files or databases.
+To maintain accuracy and consistency, scraping schedules can be automated at regular intervals, such as weekly or monthly, to monitor price changes over time.
+Moreover, including user-agent strings and introducing delays in the scraping process helps simulate human browsing behavior and reduce the risk of triggering website anti-scraping mechanisms.
 
-The temporal dimension, captured through the ‘month’ variable, enables us to measure the impact of seasonal patterns on pricing.
-Constructed variables such as ‘price_per_unit’ provide a standardized measurement that allows comparisons across different products, accounting for varying package sizes or quantities.
-This transformation from abstract consumer behaviors and retail dynamics into structured data ensures that the analysis remains rooted in real-world market phenomena, thereby enabling a more nuanced understanding of pricing behavior and trends across different vendors.
+This research translates real-world phenomena into structured data entries within the dataset to analyze pricing behavior effectively.
+Variables such as ‘vendor,’ ‘current_price,’ and ‘month’ capture consumer purchasing patterns and their relationships with retail environments.
+The ‘vendor’ variable identifies the retailer, allowing an analysis of how retail settings influence pricing strategies.
+The ‘current_price’ and ‘old_price’ variables document pricing trends, enabling the measurement of changes in consumer costs over time and their connections to economic activity.
+
+The ‘month’ variable introduces a temporal aspect, capturing the influence of seasonal patterns on pricing.
+Derived variables, such as ‘price_per_unit,’ standardize product comparisons by accounting for differences in package sizes or quantities.
+This transformation of consumer behavior and retail dynamics into structured data ensures that the analysis remains tied to real-world market activities, enhancing the understanding of pricing trends and behaviors across various retail contexts.
+
+Additional details about the dataset are provided in the datasheet, which can be accessed through the repository associated with this paper.
 
 ## Outcome variables
-The outcome variable for this study, current_price, serves as the key focus of analysis in understanding product pricing dynamics.
-This section provides a structured overview of the current_price variable, including its distribution and any observed trends, supported by both summary statistics and visualizations.
-The aim is to offer a clear understanding of how product prices vary and what factors may influence those variations.
+
+The outcome variable for this study, current_price, is the central focus for analyzing product pricing patterns in the Canadian grocery market.
+It represents the price of beef at the time of data collection, providing a direct measure of how prices vary across different retail settings.
+By examining current_price, the study captures both short-term price changes and longer-term strategies used by vendors.
+
+This section outlines the characteristics of the current_price variable, including its distribution and observed patterns, supported by summary statistics and visualizations.
+Key statistics, such as the mean, median, minimum, and maximum values, describe the central tendency and range of beef prices across vendors.
+Measures of variability, such as the standard deviation, illustrate the extent of price dispersion, showing how consistently prices are set across products and time.
+This descriptive analysis establishes a basis for identifying significant patterns in beef pricing.
 
 ### Current Price
-The outcome variable in this study is current_price.
-This variable represents the price of the product at the time of data collection and is central to understanding the effects of various factors on product pricing.
-By modeling current_price, we aim to explore how factors such as historical pricing, vendor type, and seasonal influences impact current market prices.
-This variable is crucial for determining the pricing dynamics employed by different vendors and how those strategies affect consumer costs.
 
-To provide a comprehensive view of the outcome variable, Here's a table for easier understanding (@tbl-current_price).
-This table provides the minimum, maximum, mean, median, and standard deviation of the current_price variable, offering insights into the overall distribution and variability in product prices.
+This variable represents the price of the product at the time of data collection and serves as the foundation for analyzing how various factors influence pricing.
+By modeling current_price, the study examines the impact of historical pricing, vendor type, and seasonal patterns on current market prices.
+Understanding this variable is essential for identifying pricing strategies used by vendors and their effects on consumer expenses.
+
+To present an overview of the outcome variable, a table is provided (@tbl-current_price).
+This table includes key statistics such as the minimum, maximum, mean, median, and standard deviation of the current_price variable.
+These measures summarize the distribution and variability of product prices, helping to characterize overall pricing patterns in the dataset.
 
@@
     Value = c(
-      min(current_price, na.rm = TRUE),
-      max(current_price, na.rm = TRUE),
-      mean(current_price, na.rm = TRUE),
-      median(current_price, na.rm = TRUE),
-      sd(current_price, na.rm = TRUE)
+      round(min(current_price, na.rm = TRUE), 3),
+      round(max(current_price, na.rm = TRUE), 3),
+      round(mean(current_price, na.rm = TRUE), 3),
+      round(median(current_price, na.rm = TRUE), 3),
+      round(sd(current_price, na.rm = TRUE), 3)
     )
@@
 
-In addition to the summary table, the distribution of current_price is visualized through a histogram (@fig-current_price), highlighting common price ranges and the overall spread.
-This visualization helps to identify any skewness or clustering in the data, which may indicate specific pricing patterns.
-Together, these tables and graphs provide a comprehensive understanding of the outcome variable, offering a detailed view of how product prices vary across different vendors and over time.
+@tbl-current_price highlighting key measures of central tendency and variability.
+The data ranges from a minimum value of 0.77 to a maximum of 23.98, illustrating a broad spread.
+The mean value is 9.11, indicating that most beef prices are concentrated around this range.
+The median is 8.58, suggesting that half of the observations fall below this value.
+The proximity of the mean and median implies a relatively symmetric distribution.
+The standard deviation of 4.33 reflects moderate variability, indicating that most values are distributed within approximately 4.33 units of the mean, providing a clearer understanding of the data's spread.
+
+In addition to the summary table, @fig-current_price visualizes the distribution of current_price, showing common price ranges and the overall spread.
+This graph helps identify any skewness or clustering in the data, which could indicate specific pricing patterns.
+Combined, the table and histogram provide a detailed overview of the outcome variable, illustrating how product prices differ among vendors and change over time.
 
@@
 
-The boxplot (@fig-current_price) presented above provides a detailed summary of the distribution of current_price for the products in the dataset.
-Boxplots are useful visualization tools for understanding the spread, central tendency, and overall variability of continuous data.
-In this case, the boxplot helps in identifying key features such as the median price, the interquartile range (IQR), and any potential outliers that may exist in the pricing data.
+The @fig-current_price provides an overview of the distribution of current_price for the products in the dataset.
+This visualization highlights key features such as the median price, interquartile range (IQR), and any potential outliers in the pricing data.
 
-The median, represented by the horizontal line within the box, provides an indication of the central price of products in the dataset, highlighting where the bulk of prices are situated.
-The box represents the IQR, showing where the middle 50% of data points lie, thus giving insights into price concentration.
-The whiskers, extending from the box, represent the range within which most prices fall, excluding any extreme values or outliers.
-No outliers were visibly identified in this boxplot, which suggests that most product prices are consistently within a particular range.
+The median, represented by the horizontal line within the box, indicates the central tendency of product prices, showing where most values are centered.
+The box itself represents the IQR, which captures the range within which the middle 50% of prices fall, offering a view of price concentration.
+The whiskers extending from the box depict the range of prices outside the IQR, excluding extreme values.
+In this case, no significant outliers were detected, suggesting that product prices are generally within a consistent range.
 
-The flat shape of the boxplot with a wide range of the IQR indicates low variability in current product prices.
-The relatively small difference between the lower and upper bounds of the box implies that the pricing of the products is consistent, with most prices concentrated around the median.
-This consistency may suggest that vendors are using similar pricing strategies or that there is limited differentiation in the types of products sold across vendors.
-Overall, the boxplot is a valuable tool for summarizing and communicating the underlying distribution of current_price, which aids in understanding the dynamics of pricing across different vendors.
+The relatively flat boxplot, combined with a narrow IQR, indicates limited variability in current_price.
+The small difference between the lower and upper bounds of the box suggests that product pricing is consistent, with most prices clustering around the median.
+This consistency could reflect standardized pricing strategies among vendors or a lack of variation in the types of products sold.
+Overall, the boxplot effectively summarizes the distribution of current_price and provides a clear view of pricing patterns across vendors.
 
 ## Predictor variables
-The predictor variables in this study—month, old_price, and vendor—each play an important role in explaining variations in current_price.
-This section outlines how these variables potentially influence pricing, with a focus on capturing temporal patterns, historical effects, and vendor-specific pricing strategies.
-The summary tables and visual analyses help contextualize the impact of each predictor on the outcome, providing a broad understanding of the pricing dynamics within the dataset.
+
+The predictor variables in this study—month, old_price, and vendor—each contribute to explaining variations in current_price and provide a structured view of the factors influencing beef pricing across retail settings.
+These variables reflect temporal trends, the influence of historical pricing, and differences in vendor strategies, collectively aiding in the understanding of product pricing in the Canadian grocery market.
+
+The month variable represents the temporal aspect of pricing, enabling the analysis of seasonal patterns in beef prices.
+Treating month as a categorical variable allows for the identification of months with higher or lower prices, potentially tied to changes in demand due to holidays, promotions, or seasonal supply fluctuations.
+For instance, months associated with holidays like Thanksgiving or cultural festivals may show increased demand, which could lead to price adjustments.
+This temporal perspective helps to assess whether vendors apply seasonal pricing strategies that influence current_price over time.
 
 ### Month
-The month variable is used as a predictor to capture potential seasonal influences on pricing.
-By including month, we can observe if specific times of the year are associated with higher or lower prices.
-This is particularly important for identifying temporal patterns in pricing, which can be influenced by factors such as holidays, sales events, or seasonal demand changes.
-Table (@tbl-month_summary) provides a summary of the observations for each month, showing how frequently each month appears in the dataset.
-This information can help highlight any periods with higher or lower data collection frequency, which could influence the interpretation of seasonal pricing effects.
+
+The month variable serves as a predictor to account for seasonal effects on pricing.
+Including month allows the analysis to identify whether certain times of the year are associated with price fluctuations.
+Factors such as holidays, promotional events, or changes in seasonal demand may contribute to these variations.
+@tbl-month_summary presents a breakdown of the observations for each month, showing the frequency of data collection across the year.
+This summary helps identify any months with disproportionately high or low representation, which could impact the interpretation of seasonal pricing patterns.
 
@@
 
+The @tbl-month_summary presents the number of observations recorded for each month, spanning June to November.
+The highest count is in August (624), while the lowest is in November (342).
+This variation may reflect increased market activity or data collection efforts during the summer, possibly due to higher consumer demand during barbecue season.
+Conversely, the decline in November could be linked to reduced data collection or changes in consumer purchasing patterns.
+Since data collection began in early 2024, the dataset does not represent the entire year, limiting the ability to evaluate pricing trends across all seasons.
+These monthly differences highlight the importance of accounting for seasonal factors when analyzing pricing patterns.
+
 ### Old Price
-old_price serves as another key predictor variable, representing the historical price of a product.
-This variable helps gauge the effect of historical pricing on current pricing strategies, indicating whether a product has undergone discounts or price hikes.
-Table (@tbl-old_price_summary) provides a summary of key statistics for old_price, including minimum, maximum, mean, median, and standard deviation.
-Understanding the distribution of old_price helps us assess whether historical prices significantly differ from current prices, and whether pricing adjustments (e.g., discounts) have been applied uniformly across products.
-The relationship between old_price and current_price is visually explored in Chart 4, which helps understand the influence of past pricing decisions on current prices.
+
+old_price is a key predictor variable that represents a product's historical price.
+This variable is used to examine how past pricing influences current pricing strategies, indicating whether products have experienced discounts or price increases.
+@tbl-old_price_summary summarizes the key statistics for old_price, including its minimum, maximum, mean, median, and standard deviation.
+Analyzing the distribution of old_price allows us to evaluate whether historical prices significantly differ from current prices and whether adjustments, such as discounts, are applied consistently across products.
+The relationship between old_price and current_price is further illustrated in Chart 4, providing a visual representation of how historical pricing affects current price decisions.
 
@@
     Value = c(
-      Min = min(old_price, na.rm = TRUE),
-      Max = max(old_price, na.rm = TRUE),
-      Mean = mean(old_price, na.rm = TRUE),
-      Median = median(old_price, na.rm = TRUE),
-      `Standard Deviation` = sd(old_price, na.rm = TRUE)
+      round(min(current_price, na.rm = TRUE), 3),
+      round(max(current_price, na.rm = TRUE), 3),
+      round(mean(current_price, na.rm = TRUE), 3),
+      round(median(current_price, na.rm = TRUE), 3),
+      round(sd(current_price, na.rm = TRUE), 3)
   )
@@
 
+The @tbl-old_price_summary summarizes the statistics for the predictor variable old_price, representing the historical price of beef.
+The minimum value is \$1.87, and the maximum value is \$25.99, reflecting a wide range of historical prices for beef products.
+The mean value of \$11.48 indicates that, on average, beef products were priced near this level in the past.
+The median of \$10.97 shows that half of the products had historical prices below this value, suggesting a slightly skewed distribution.
+The standard deviation of \$5.32 reflects moderate variability, indicating that while many products were priced close to the average, some were significantly higher or lower.
+These statistics help describe the variation and central tendency in historical beef pricing.
+
 ### Vendor
-The vendor variable is an important categorical predictor that differentiates pricing behavior across different retail chains.
-It allows us to analyze how pricing strategies vary among vendors such as Walmart and Galleria.
-Table (@tbl-vendor_summary) shows the count of observations for each vendor, allowing us to compare how frequently data was collected for each retail chain.
-This helps in understanding the representation of each vendor within the dataset and assessing whether certain vendors may have a stronger influence on the overall analysis.
-Chart 2, which shows the price difference by vendor, provides insights into how different vendors adjust their pricing strategies, offering valuable comparisons among them.
-Additionally, Chart 3 presents the average current_price over time, broken down by vendor, providing a clearer picture of how vendor-based strategies influence pricing trends.
+
+The vendor variable is a categorical predictor that captures differences in pricing behavior across retail chains.
+It enables the analysis of how pricing strategies differ among vendors.
+@tbl-vendor_summary lists the number of observations for each vendor, helping to evaluate the representation of each retail chain in the dataset and determine whether certain vendors have a larger impact on the overall analysis.
 
@@
 
+The @tbl-vendor_summary presents the number of observations for each vendor, with Walmart accounting for 1,696 and T&T for 1,330.
+While there is a difference in the number of observations, the counts are close enough to allow for a meaningful comparison between the two vendors.
+This indicates that while Walmart may offer a slightly wider range or maintain more consistent availability of beef products, T&T’s dataset is still robust enough to analyze pricing patterns effectively.
+This balance ensures the analysis reflects variations between a mainstream retailer and a culturally focused store without being significantly affected by unequal sample sizes.
+
 # Model {#sec-model}
@@
 
-To model the current price of beef, $y_i$, at time $i$, we use a Bayesian linear regression model implemented with the `stan_glm` function in the R package `rstanarm`.
+To model the current price of beef, $y_i$, at time $i$, we apply a Bayesian linear regression model using the `stan_glm` function from the R package `rstanarm`.
 The response variable, $y_i$, represents the current price of beef in dollars.
-Our predictors include $x_1$, $x_2$, and $x_3$, which represent the month, old price, and vendor, respectively.
-Each component of the model is defined and justified below.
+The predictors include $x_1$, $x_2$, and $x_3$, corresponding to the month, old price, and vendor, respectively.
+Each component of the model is defined and explained in the following sections.
 
@@
 
-Define $y_i$ as the current price of beef at time $i$, measured in dollars.
-Let $x_1$, $x_2$, and $x_3$ represent the predictors: $x_1$ is the month (categorical variable), $x_2$ is the old price, and $x_3$ is the vendor (categorical variable).
+Define $y_i$ as the current price of beef at time $i$, expressed in dollars.
+The predictors are represented as follows: $x_1$ is the month (a categorical variable), $x_2$ is the old price, and $x_3$ is the vendor (a categorical variable).
 
@@
 In this model:
-- $\alpha$ represents the intercept term, capturing the baseline price of beef.
-- $\beta_1$ corresponds to the effect of the month on the current price, allowing us to capture seasonal variations that may impact beef pricing.
-The month is treated as a categorical variable, acknowledging that different months can bring different demand or supply effects due to holidays, weather, or other seasonal factors.
-- $\beta_2$ represents the effect of the old price on the current price. This allows us to account for price inertia or trends where past pricing influences current pricing.
-- $\beta_3$ represents the effect of the vendor, also treated as a categorical variable.
-Different vendors may have distinct pricing strategies, and including vendor as a categorical predictor helps us capture this vendor-specific pricing variation.
+- $\alpha$ denotes the intercept term, representing the baseline price of beef when all predictors are at their reference levels.  
+- $\beta_1$ represents the effect of the month on the current price, capturing seasonal patterns that may influence beef prices.
+The month is treated as a categorical variable, recognizing that demand and supply conditions can vary by month due to factors such as holidays, weather, or promotional periods.
+- $\beta_2$ reflects the effect of the old price on the current price, accounting for the influence of historical pricing.
+This term helps identify whether past prices have a consistent impact on current pricing decisions.
+- $\beta_3$ captures the effect of the vendor, also treated as a categorical variable.
+Different vendors may apply unique pricing strategies, and including this variable allows us to model variations specific to each vendor.
 
@@
 
-These priors are chosen to be weakly informative, allowing the data to speak for itself while still providing a reasonable range for the coefficients based on prior expectations.
-Specifically, a standard deviation of 2.5 reflects our expectation that most reasonable effects should fall within a plausible range without being overly restrictive.
+These priors are designed to be weakly informative, offering guidance while allowing the data to determine the outcomes.
+A standard deviation of 2.5 reflects the expectation that most effects are likely to fall within a reasonable range, ensuring the priors are flexible without imposing overly restrictive assumptions.
 
@@
 
-The choice of predictors is well-balanced in terms of complexity and appropriateness for this scenario.
-By including month and vendor as categorical variables, we acknowledge the significant impact that both seasonal changes and vendor-specific factors can have on the current price of beef.
-The inclusion of the old price reflects the idea of price inertia, where historical prices are likely to influence current pricing trends.
+The selected predictors strike a balance between relevance and simplicity for this scenario.
+Including `month` and `vendor` as categorical variables accounts for the influence of seasonal changes and vendor-specific pricing strategies on the current price of beef.
+The use of `old_price` reflects the concept of price inertia, where past prices are likely to affect current pricing patterns.
 
-The model is neither overly simplistic nor unnecessarily complex: it captures important predictors without adding extraneous complexity that could lead to overfitting.
-Month and vendor are treated as categorical variables rather than grouping by more arbitrary categories, which maintains the expressiveness of our model while being parsimonious.
+The model avoids being overly simplistic or unnecessarily complex by incorporating key predictors without introducing excessive variables that could risk overfitting.
+Treating `month` and `vendor` as categorical variables ensures the model captures relevant distinctions without arbitrary groupings, maintaining both clarity and parsimony.
 
-The model is implemented using `stan_glm` from the `rstanarm` package, a high-level interface to Stan for Bayesian modeling.
-Stan provides powerful sampling algorithms, making it an appropriate tool for fitting our Bayesian model efficiently, ensuring robust convergence and accurate posterior estimation.
+The model is implemented using `stan_glm` from the `rstanarm` package, which provides an accessible interface to Stan for Bayesian modeling.
+Stan's efficient sampling algorithms ensure reliable convergence and accurate estimation of posterior distributions, making it a robust choice for this analysis.
 
 ### Assumptions and Limitations
-- **Linearity**: The model assumes a linear relationship between the predictors and the response variable. This may be a limitation if the true relationship is nonlinear.
-- **Normality of Residuals**: The residuals are assumed to follow a normal distribution. If this assumption does not hold, it could bias the model predictions.
-- **Priors**: We used weakly informative priors to allow flexibility. However, if we had prior knowledge of specific pricing patterns, more informative priors could be used for better inference.
+- **Linearity**: The model assumes a linear relationship between the predictors (month, old_price, vendor) and the response variable (current_price).
+This assumption simplifies the analysis and enhances interpretability, but it may limit the model's ability to capture more complex relationships.
+For example, interactions or threshold effects between predictors and the response variable would not be reflected in a purely linear framework.
+If these nonlinear relationships exist but are not modeled, the resulting estimates and predictions may be less accurate, potentially leading to incorrect conclusions about factors affecting beef pricing.
+
+- **Normality of Residuals**: The residuals are assumed to follow a normal distribution, a requirement for making valid statistical inferences about model parameters.
+If this assumption is violated—such as when residuals are skewed or have heavy tails—it could result in biased estimates, unreliable confidence intervals, and less dependable predictions.
+Non-normal residuals might suggest missing variables in the model or that a different modeling approach, such as a generalized linear model, might better fit the data.
+
+- **Priors**: The Bayesian framework employs weakly informative priors to allow the data to primarily determine parameter estimates.
+While this provides flexibility, it also means that stronger prior knowledge of pricing patterns could have been used to improve precision.
+Using more specific priors, informed by domain knowledge, might enhance the model's performance by producing more stable estimates.
+However, weakly informative priors can make the model more sensitive to data quality and outliers, as they do not strongly influence the parameter estimates.
+In cases where the data are limited or noisy, stronger priors could help stabilize the model's outputs.
 
 ### Validation and Diagnostics
-- **Posterior Predictive Checks**: We performed posterior predictive checks to verify that the model adequately captures the variation in the data.
-These checks involve comparing simulated data from the posterior distribution to the observed data to check for any discrepancies.
-- **Convergence**: Model convergence was assessed using trace plots and the R-hat statistic, with all values being close to 1, indicating successful convergence.
-- **Alternative Models**: We also considered fitting simpler linear models that did not include the vendor or seasonal components.
-These models showed significantly lower predictive performance, indicating the importance of including both the vendor and seasonal factors for accuracy.
+- **Posterior Predictive Checks**: Posterior predictive checks were conducted to evaluate how well the model represents the variation in the data.
+These checks compare simulated data generated from the posterior distribution to the observed data to identify any discrepancies.
+Assessing the alignment between these distributions helps determine whether the model captures the underlying patterns in the data.
+A close match indicates that the model is appropriately capturing the structure of the data, while notable differences may suggest issues such as omitted variables or an incorrect model specification.
+These checks are a key step in validating the model's ability to generalize beyond the observed dataset.
 
-In summary, the chosen model provides a comprehensive yet interpretable understanding of how different factors influence the current price of beef.
-The Bayesian approach allows us to incorporate uncertainty and prior beliefs, while our specific choice of priors and predictors ensures that the model is well-suited for the available data and research questions.
+- **Convergence**: Model convergence was assessed using trace plots and the R-hat statistic, with all R-hat values near 1, confirming successful convergence.
+Trace plots illustrate how effectively the Markov Chain Monte Carlo (MCMC) sampling algorithm explores the parameter space.
+Well-mixed chains with stable trajectories suggest the model has reached a stationary distribution.
+The R-hat statistic, also known as the Gelman-Rubin diagnostic, compares between-chain and within-chain variance, with values close to 1 indicating that all chains have converged to the same target distribution.
+Together, these diagnostics ensure the reliability of the parameter estimates and confirm that sampling issues did not influence the results.
+
+- **Alternative Models**: Simpler linear models excluding vendor or seasonal variables were also tested.
+These models showed notably lower predictive performance, highlighting the importance of including both vendor and seasonal factors for accurate modeling.
+Excluding these variables reduced the model's fit, as indicated by metrics like R-squared and predictive accuracy.
+This comparison reinforces the necessity of accounting for vendor-specific strategies and seasonal variations to effectively model beef pricing.
+The evaluation of alternative models confirmed that the selected predictors provide a better representation of the factors influencing price changes in the dataset.
+
+In conclusion, the chosen model effectively captures how various factors influence the current price of beef.
+The Bayesian framework incorporates uncertainty and prior knowledge, while the selection of priors and predictors ensures the model aligns well with the available data and research objectives.
 
@@
 
-Figure 2 (@fig-current_price_distribution) presents the distribution of current beef prices in the dataset, offering insights into the range and frequency of beef pricing.
-The horizontal axis represents the price in dollars, ranging from 0 to approximately 25, while the vertical axis indicates the frequency, or the number of observations, within each price range.
-The grey bars illustrate the frequency distribution, showing how beef prices are spread across different price levels.
+@fig-current_price_distribution illustrates the distribution of current beef prices in the dataset, depicting how prices are spread across different levels.
+The horizontal axis represents beef prices in dollars, ranging from 0 to approximately 25, while the vertical axis shows the frequency, or the number of observations, within each price range.
+The grey bars represent the histogram, visualizing the frequency distribution of beef prices.
 
-The histogram indicates that the most frequent price range is between \$5 and \$10, with a notable peak in the frequency around \$7 to \$8.
-This suggests that a significant proportion of beef products are priced within this range, making it a central pricing cluster in the market.
-Additionally, there are smaller peaks around \$13 and \$15, hinting at secondary groupings where beef prices are relatively common.
-The declining frequency beyond \$15 suggests fewer beef products are priced in the higher range, indicating that higher prices are less typical in this dataset.
+The histogram shows that the most common price range is between \$4 and \$8, with noticeable peaks around \$5 and \$8.
+This indicates that a significant portion of beef products falls within this range, suggesting a central pricing cluster in the market.
+Smaller peaks around \$13 and \$15 point to secondary groupings where beef prices are also relatively frequent.
+Beyond \$15, the frequency declines sharply, indicating that higher-priced beef products are less common and that most items are concentrated within a lower price range.
 
-The overlaid red line represents a smoothed density curve that helps to visualize the general shape and distribution trend of the prices.
-The curve follows the histogram closely, showing a skewed distribution with a concentration towards the lower prices, and tails off as prices increase.
-This pattern suggests that while lower-priced beef products are more common, higher prices are less frequent, potentially indicating a general preference or affordability constraint for lower-cost beef among consumers.
-The presence of multiple peaks could also imply different pricing tiers, perhaps influenced by factors such as beef quality or specific product types.
+The overlaid red line represents a smoothed density curve, highlighting the overall trend in the distribution.
+The curve closely aligns with the histogram, showing a skewed pattern with a higher concentration of prices in the lower range and a gradual decrease as prices rise.
+This distribution suggests that consumers may prefer lower-priced beef products or that pricing aligns with affordability constraints.
+The multiple peaks in the curve indicate distinct pricing tiers, potentially reflecting differences in product quality, cut types, or retailer-specific pricing strategies.
 
@@
 
-Figure 3 (@fig-price_difference_by_vendor) illustrates the distribution of price differences between beef products offered by two vendors, T&T and Walmart, providing an insight into the variability of pricing across these retailers.
-The y-axis represents the price difference in dollars, while the x-axis labels the two vendors under comparison.
-The boxplots visually depict the range, median, and variability in the price differences for each vendor.
+@fig-price_difference_by_vendor shows the distribution of price differences for beef products sold by T&T and Walmart, highlighting variations in pricing between these retailers.
+The y-axis represents the price difference in dollars, while the x-axis lists the two vendors.
+The boxplots display the range, median, and variability of price differences for each retailer.
 
-For T&T, the boxplot shows a wider range of price differences, with the interquartile range (IQR) extending from roughly \$1.5 to \$4, indicating a moderate to substantial variability in price differences among the products they sell.
-The median price difference is around \$2.5, suggesting that many of the products have a price difference near this value.
-However, T&T also shows significant variability with the presence of two outliers beyond \$7.5 and \$10, indicating that a few beef products exhibit much larger price differences compared to the rest of the data.
+For T&T, the boxplot indicates a broader range of price differences, with the interquartile range (IQR) spanning approximately \$1.5 to \$4.
+The median price difference is about \$2.5, suggesting that many products cluster around this value.
+T&T also shows greater variability, with outliers extending beyond \$7.5 and \$10.
+These outliers suggest that certain products have significantly larger price differences, possibly reflecting diverse pricing strategies or variations in product offerings.
 
-In contrast, Walmart exhibits a narrower spread of price differences, as indicated by the more compact boxplot.
-The IQR for Walmart lies approximately between \$1 and \$2.5, and the median price difference is lower compared to T&T, close to \$1.5.
-This suggests that Walmart’s beef products tend to have smaller and less variable price differences, indicating more consistent pricing practices.
-There are also a few outliers above \$4, but these are less extreme compared to T&T's outliers.
+In comparison, Walmart's boxplot shows a narrower distribution of price differences.
+The IQR ranges from roughly \$1 to \$2.5, with a median of around \$1.5.
+This tighter range indicates that Walmart’s beef pricing differences are more consistent, reflecting a standardized pricing strategy.
+While there are a few outliers above \$4, these deviations are less pronounced than those seen for T&T.
 
-Overall, the boxplots suggest that T&T has more variability in their pricing, possibly reflecting differences in product types or marketing strategies, whereas Walmart tends to maintain a narrower price range with fewer significant deviations.
-This disparity may indicate that Walmart focuses on more standardized pricing practices, which could influence consumer choice based on price predictability and consistency.
+The boxplots emphasize a clear difference in pricing approaches: T&T exhibits more variability and larger outliers, potentially due to a broader product assortment or targeted market strategies.
+Walmart, on the other hand, maintains more uniform pricing, offering greater predictability for consumers.
+These patterns reflect how each retailer positions its beef products in the market.
 
@@
 
-Figure 4 (@fig-average_price_over_time) shows the trend of average beef prices over time, providing insight into fluctuations in the market from July to November.
-The y-axis represents the average current price in \$, while the x-axis represents the date, with labels indicating the progression through the months.
+@fig-average_price_over_time displays the trend in average beef prices from July to November, highlighting fluctuations in the market during this period.
+The y-axis represents the average current price in dollars, while the x-axis shows the progression of dates.
 
-The line chart indicates that average beef prices experienced significant variability throughout the observed period, with several sharp peaks and troughs.
-In early July, prices were relatively high, reaching a peak of approximately \$13, but this was followed by a rapid decline, dropping to around \$7 by the end of the month.
-The fluctuations continue through August, where we see notable volatility, including several spikes back to higher average prices of over \$10.
-Such fluctuations suggest sensitivity in pricing, possibly influenced by seasonal changes, supply chain disruptions, or market demand dynamics.
+The chart shows a general downward trend in average prices starting in early July, when prices were around \$11, and decreasing to approximately \$7 by the end of the month.
+This decline could reflect changes in market dynamics, such as seasonal shifts in demand or adjustments in supply.
+In August, prices display moderate fluctuations, with occasional spikes above \$9, possibly due to short-term disruptions or promotional pricing events.
 
-Entering September and October, there is a pronounced drop in prices, with average prices occasionally dipping below \$6, reflecting a period of reduced pricing.
-Following this low, prices exhibit recovery and some stabilization, hovering around \$8 to \$9 from October into early November, albeit with some ongoing small fluctuations.
-The chart’s volatility indicates that beef prices are influenced by various factors throughout these months, with possible implications from supply conditions, retailer pricing strategies, or changing consumer demand.
-This temporal analysis of beef pricing is critical for understanding market stability and identifying periods of significant price changes that may warrant further investigation.
+September and October show a more consistent decline, with average prices dropping below \$6 at times.
+This extended period of lower prices may point to factors like an oversupply or reduced consumer demand.
+Toward the end of October, prices begin to recover, showing a gradual upward trend into early November and stabilizing between \$8 and \$9.
+The shaded region around the line, representing confidence intervals, widens at the edges of the time range, indicating increased uncertainty in the estimates.
+
+This pattern suggests that beef prices during this timeframe were shaped by seasonal trends, retailer pricing strategies, and possible supply chain fluctuations.
+The observed variability and eventual stabilization highlight the dynamic nature of the market and emphasize the importance of tracking temporal pricing trends to better understand market behavior and inform retailer or policy decisions on pricing strategies and affordability.
 
@@
 
-Figure 5 (@fig-current_vs_old_price) has been modified to display a heatmap of the relationship between old and current price levels for beef, providing a more aggregated view of how pricing has evolved over time for various products.
-The x-axis represents the old price level, while the y-axis indicates the current price level, with each cell in the heatmap representing a specific combination of these price levels.
-The color intensity, ranging from light red to dark red, corresponds to the frequency of observations, as indicated by the legend on the right.
+@fig-current_vs_old_price presents a heatmap showing the relationship between old and current beef prices, illustrating how pricing has changed over time across various products.
+The x-axis represents old price levels, and the y-axis represents current price levels, with each cell indicating the frequency of products at specific combinations of these levels.
+The intensity of the cell color, ranging from light red to dark red, reflects the density of observations, as described by the legend on the right.
 
-The heatmap reveals a clear diagonal pattern where higher old price levels correspond to higher current price levels, indicating a strong positive association between historical and current pricing.
-The darker cells along the diagonal reflect the highest frequency of observations, suggesting that most products maintain similar relative pricing over time, reinforcing the consistency in price adjustments observed earlier.
-This alignment along the diagonal suggests that beef pricing adjustments have been proportional and systematic, potentially influenced by inflation, consistent vendor practices, or other market mechanisms.
+The heatmap shows a clear diagonal pattern, indicating that higher old price levels correspond closely to higher current price levels.
+This alignment highlights a strong positive correlation between historical and current prices, suggesting that most products have undergone proportional pricing adjustments.
+The darker cells along the diagonal represent the most frequent combinations of old and current prices, pointing to consistent adjustments influenced by factors such as market stability, inflation, or standardized vendor practices.
 
-The off-diagonal cells that are lightly shaded or empty indicate fewer occurrences where current prices deviate significantly from historical ones.
-The absence of significant variation off the main diagonal suggests that the majority of pricing changes for beef are directly related to past prices, implying stability in pricing trends with limited irregular adjustments.
-This consistency may be indicative of structured pricing strategies across different market conditions, with only minor deviations that could result from specific product promotions or temporary supply issues.
+In contrast, off-diagonal cells, which are lightly shaded or empty, represent rare occurrences of significant deviations from the proportional trend.
+These sparse areas indicate that sharp discounts or large price increases are infrequent, reinforcing the observation that pricing tends to remain stable and closely tied to historical values.
 
-Overall, the heatmap effectively illustrates the overall pattern of pricing consistency, highlighting that both old and current beef prices are closely aligned.
-This representation captures the density of occurrences more clearly than individual scatter points, providing a visual summary of the frequency and typicality of price pairings between historical and current beef prices.
+Overall, the heatmap provides an aggregated view of the consistency in beef pricing across different conditions.
+By grouping pricing data into distinct levels, it visually emphasizes the alignment between old and current prices, highlighting systematic adjustments and the rarity of irregular deviations, which may be influenced by vendor strategies or external market factors.
 
@@
 
-Figure 6 (@fig-unit_price_distribution) illustrates the distribution of unit prices for beef, measured in \$ per 100 grams.
-The x-axis represents the unit price, while the y-axis indicates the frequency of occurrences for each price range.
-The histogram, represented by grey bars, shows how often each unit price appears in the dataset, and the red curve provides a smoothed density estimate to visualize the general trend.
+@fig-unit_price_distribution displays the distribution of unit prices for beef, measured in \$ per 100 grams.
+The x-axis shows unit prices, while the y-axis represents density, indicating the proportion of observations across different price ranges.
+The grey area under the density curve highlights the relative concentration of unit prices, while the red curve provides a smoothed estimate of the overall distribution.
 
-The most striking feature of this figure is the concentration of unit prices near zero, with a significant spike in the frequency at this price point.
-This suggests that a substantial number of observations in the dataset have a very low or nominal unit price, which could indicate promotions, discounted products, or even data entry inconsistencies that require further examination.
-After this initial peak, the frequency drops sharply, followed by a few smaller peaks across higher unit price ranges, particularly around \$1, \$3, and \$6 per 100 grams.
-These smaller peaks indicate the presence of additional groupings of unit prices, potentially corresponding to different product categories, quality grades, or pricing tiers.
+A notable feature of this figure is the high density near \$0, indicating that a large number of observations in the dataset have very low unit prices.
+This concentration could reflect promotions, bulk discounts, or potential data issues that may require further review.
+Beyond this initial spike, the density decreases sharply, with smaller secondary peaks appearing around \$2, \$4, and \$6 per 100 grams.
+These peaks may correspond to different pricing tiers influenced by factors such as product quality, cut type, or retailer-specific strategies.
 
-The overall distribution is heavily right-skewed, suggesting that the majority of beef products have lower unit prices, while higher prices are less common and scattered.
-This right-skewness might be indicative of pricing strategies that cater to a wide consumer base, where most products are priced affordably, with only a few premium items available at significantly higher unit prices.
-The presence of several distinct peaks, rather than a single uniform distribution, suggests price differentiation that could be driven by various factors, such as beef cut type, quality, or retailer-specific pricing strategies.
-This distribution pattern highlights the diverse pricing landscape for beef products in the dataset, suggesting the existence of both low-cost and premium segments in the market.
+The distribution is heavily right-skewed, with the majority of observations clustered at lower price points and fewer cases of higher-priced beef products.
+This pattern suggests that the market is dominated by affordable beef options catering to a broad consumer base, while premium products occupy a smaller share.
+The presence of multiple peaks indicates segmentation within the market, driven by variations in product attributes or retailer pricing approaches.
+This highlights the coexistence of budget-friendly and premium offerings, reflecting diverse consumer preferences and pricing strategies in the Canadian grocery market.
 
@@
 
-Table 5 (@tbl-modelresults) summarizes the results from the Bayesian model applied to analyze beef prices, including key parameter estimates and model diagnostics.
-The model provides insight into the relationship between the current price of beef and various predictors such as the old price, vendor, and the time of observation (represented by the "month" variable).
+@tbl-modelresults presents the outcomes of the Bayesian model used to analyze beef prices, highlighting parameter estimates and diagnostic measures.
+The model evaluates the relationship between the current price of beef and predictors, including the old price, vendor, and month of observation.
 
-The estimated intercept of -0.37 suggests that, after controlling for the predictors in the model, the baseline level of the current price is negative, though this value must be interpreted with caution in terms of its contextual implications.
-The "month" coefficient is -0.01, indicating a slight negative association between the time variable and current price, although the magnitude of the effect is relatively small.
-This suggests that as time progresses, there is a minor decrease in the current price, potentially reflecting seasonal effects or gradual changes in market dynamics.
-The "old_price" coefficient is 0.81, showing a strong positive relationship with the current price, meaning that for each unit increase in the old price, the current price increases by a comparable margin.
-This result implies a consistent upward adjustment in beef pricing over time, reinforcing the trends seen in previous charts.
-Finally, the "vendorWalmart" coefficient is 0.56, indicating that beef sold by Walmart tends to be priced higher, on average, than that sold by the baseline vendor (T&T), which aligns with our earlier findings on vendor price differences.
+The estimated intercept of -0.37, while seemingly unusual, serves as a reference point within the model and should be interpreted in the context of the predictors.
+The "month" coefficient of -0.01 indicates a small downward trend in current prices over time, possibly reflecting seasonal influences or market adjustments.
+The "old_price" coefficient, at 0.81, demonstrates a strong positive relationship, indicating that higher historical prices correspond to proportionally higher current prices, suggesting consistent pricing patterns over time.
+Additionally, the "vendorWalmart" coefficient of 0.56 shows that Walmart's beef prices are, on average, higher than those at T&T, consistent with observed vendor-specific pricing strategies.
 
-In terms of model fit, the R-squared (R²) value of 0.945 and the adjusted R-squared value of 0.944 indicate a high degree of explained variance, suggesting that the model is capable of capturing the main factors influencing beef prices.
-The negative log-likelihood value (-4349.672) and expected log pointwise predictive density (ELPD) of -4354.7 provide additional metrics on model fit, while other diagnostic values, such as the leave-one-out information criterion (LOOIC) and Widely Applicable Information Criterion (WAIC), both around 8709, provide an indication of model performance for out-of-sample predictive accuracy.
-The root mean square error (RMSE) is 1.02, suggesting the model's predictions have a relatively small average deviation from the actual values, indicative of good predictive performance.
+In terms of model fit, the R-squared value of 0.945 and adjusted R-squared of 0.944 indicate that the model accounts for a significant portion of the variability in beef prices.
+Metrics such as the negative log-likelihood (-4349.672) and expected log pointwise predictive density (ELPD) of -4354.7 further support the model’s ability to capture key patterns in the data.
+Predictive measures, including the leave-one-out information criterion (LOOIC) and Widely Applicable Information Criterion (WAIC), both approximately 8709, confirm the model’s reliability in making predictions.
+The root mean square error (RMSE) of 1.02 reflects minimal average deviation between predicted and observed values, underscoring the model’s accuracy.
 
-These results together illustrate the strong relationship between past pricing and current pricing, as well as vendor influence on beef pricing.
-The Bayesian approach has provided robust parameter estimates while accounting for uncertainty, as reflected in the provided standard errors.
-Overall, the model appears to perform well, with high explained variance and reasonable prediction accuracy, making it a valuable tool for understanding and forecasting beef price trends.
+The results demonstrate a strong, consistent relationship between historical and current pricing while highlighting differences in pricing strategies between vendors.
+Using a Bayesian framework allows for uncertainty to be incorporated into the analysis, providing robust parameter estimates.
+The high explained variance and reliable predictive performance affirm the model’s effectiveness in analyzing and forecasting beef price trends, offering a clear understanding of market behaviors for both consumers and stakeholders.
 
@@
 
-The findings of this study provide valuable insights into the dynamics of beef pricing in the Canadian grocery market.
-Through the application of Bayesian linear regression, it is evident that both historical pricing (i.e., old_price) and vendor-specific strategies significantly influence the current pricing of beef.
-The strong positive correlation between old_price and current_price suggests a consistency in pricing practices, potentially driven by factors like inflation and vendor pricing strategies, with beef products generally following historical pricing trends.
-The vendor effect, notably the difference between T&T and Walmart, reflects distinct market positioning strategies, with Walmart generally maintaining a more standardized pricing approach, whereas T&T displays greater variability, possibly influenced by its focus on specialty goods catering to specific cultural preferences.
-Seasonal trends, captured through the month variable, also indicate minor temporal price variations, pointing to possible supply and demand effects linked to particular times of the year.
+The findings of this study provide a detailed look into the factors influencing beef pricing in the Canadian grocery market.
+Using Bayesian linear regression, the results demonstrate that historical pricing (old_price) and vendor-specific strategies are significant predictors of current prices.
+The strong positive relationship between old_price and current_price suggests a high level of consistency in pricing practices, likely influenced by factors such as inflation or vendor strategies.
+Walmart and T&T exhibit distinct pricing approaches, with Walmart showing more consistent pricing while T&T displays greater variability, possibly due to its focus on specialty goods catering to specific cultural preferences.
+Seasonal trends, represented by the month variable, indicate small but noticeable temporal price variations, likely linked to supply and demand fluctuations throughout the year.
 
-The heatmap visualizations, scatter plots, and statistical summaries all pointed towards a consistent trend in pricing, with limited deviations from expected relationships between historical and current prices.
-The consistency observed in the heatmap between old and current price levels suggests that vendors are either directly responding to historical costs or that other external factors, such as cost of production, inflation, or market demand, are similarly affecting historical and current pricing.
-The results of the Bayesian model further support these findings, highlighting the importance of including historical price as a predictor in understanding current pricing trends.
-The high R-squared value of the model also indicates that the chosen predictors adequately capture most of the variability in beef prices, reinforcing the relevance of the variables selected.
+Visualizations in the previous section (@sec-result), along with statistical summaries, reinforce these findings.
+The heatmap in @fig-current_vs_old_price demonstrates a strong alignment between old and current prices, highlighting the significant influence of historical costs on current pricing.
+Factors such as production expenses, inflation, and market demand appear to affect historical and current prices in similar ways.
+The high R-squared value of the Bayesian model (@tbl-modelresults) further confirms that the selected predictors capture the majority of variation in beef prices, emphasizing their importance.
+Additional observations can be drawn from these results, as discussed below.
 
@@
 
-One critical point of discussion is the apparent distinction in pricing strategies between Walmart and T&T.
-Walmart's pricing, as evidenced by the relatively narrow spread of the boxplot, seems to prioritize consistency and predictability, possibly reflecting its brand strategy of offering everyday low prices [@citeWalmartLowPrice].
-This approach may be designed to build consumer trust, ensuring customers that they are receiving competitive pricing without frequent fluctuations.
-On the other hand, T&T's wider range of price differences may reflect a more dynamic pricing strategy, potentially influenced by the specialty nature of its products [@citeTandT], diverse supply chain, or its responsiveness to a different customer demographic.
-Such variability in pricing could indicate promotional efforts, targeted price differentiation, or a reflection of its broader range of product quality and types.
+A key observation is the clear difference in pricing strategies between Walmart and T&T.
+Walmart’s pricing strategy, as reflected by its narrower boxplot range (@fig-price_difference_by_vendor), emphasizes consistency and predictability, aligning with its branding as a retailer offering everyday low prices [@citeWalmartLowPrice].
+This approach likely appeals to price-sensitive consumers seeking stability and affordability.
+In contrast, T&T shows a wider range of price differences, which may reflect a dynamic pricing strategy tailored to its specialty products, diverse supply chain, and specific customer demographic [@citeTandT].
+The variability in T&T’s pricing could stem from promotional efforts, targeted price differentiation, or a wider range of product qualities and types.
 
-This analysis highlights the importance of understanding different vendor strategies when interpreting pricing data.
-Walmart's emphasis on standardized pricing could attract a broader, price-sensitive customer base, while T&T's more variable pricing might appeal to niche customers seeking specific products, regardless of price changes.
-Future research could delve deeper into these strategies by including additional variables that capture product-specific characteristics, customer demographics, or promotion periods to better understand the pricing differences observed.
+Understanding these strategies is essential for interpreting pricing data.
+Walmart’s standardized pricing is likely to attract a broad customer base, while T&T’s variability might appeal to niche markets looking for specialty goods.
+Future research could expand on these findings by incorporating product-level details, customer demographics, or promotional data to better understand the drivers behind these pricing differences.
 
@@
 
-The study also suggests that beef pricing is relatively stable, with most price adjustments being proportional to historical prices.
-This stability might imply a well-functioning market where historical prices act as an anchor, helping vendors and consumers set expectations.
-This could also be indicative of supply-side factors such as contracts with suppliers or production costs that do not fluctuate significantly.
-Additionally, the moderate decrease in prices observed during certain months might suggest seasonal promotions or shifts in consumer demand that lead vendors to adjust prices accordingly.
-Such trends could be driven by factors like increased competition, holiday promotions [@citeHoliday], or seasonal variations in meat consumption.
+The study highlights a general stability in beef pricing, with adjustments largely proportional to historical prices.
+This stability may suggest a well-functioning market where historical prices act as an anchor for setting current prices, potentially influenced by supplier contracts or steady production costs.
+The modest decline in prices observed in certain months could reflect seasonal promotions or changes in demand, driven by factors like holiday sales [@citeHoliday] or shifts in meat consumption habits.
 
-Interestingly, the observed concentration of prices in the lower range might indicate that the majority of beef products are targeted towards value-oriented consumers, potentially reflecting broader consumer preferences or economic constraints.
-This could also point to the influence of retailer pricing power, where large players like Walmart are able to leverage economies of scale to keep prices low, thereby shaping overall market dynamics.
-An important next step in understanding these patterns would be to assess how external shocks, such as economic downturns or supply chain disruptions [@citeSupplyChain], influence this pricing stability, as well as whether similar patterns hold for other product categories.
+The clustering of prices at the lower end of the range suggests that the majority of beef products cater to value-conscious consumers, which may reflect broader economic trends or retailer strategies to maintain affordability.
+Larger players like Walmart might use economies of scale to keep prices competitive, shaping market dynamics as a result.
+To build on these findings, future studies could investigate how external events, such as economic downturns or supply chain disruptions [@citeSupplyChain], affect this pricing stability and whether similar patterns are observed in other product categories.
+
+## Cultural Preferences and Product Differentiation
+
+Cultural preferences play a significant role in beef pricing, as evidenced by the differences between T&T and Walmart.
+T&T’s pricing patterns, which are more varied, may reflect its focus on serving an Asian customer base with specific culinary needs [@citeCulture].
+The diversity in its product cuts and grades could contribute to these patterns, as sourcing requirements and preparation methods may differ based on cultural preferences.
+This may lead to higher variability in prices.
+
+For policymakers and retailers addressing food affordability, these findings underscore the importance of considering cultural factors in consumption patterns.
+Retailers like T&T, which cater to specific demographic groups, might need tailored pricing strategies, while larger chains like Walmart could focus on maintaining uniform pricing to appeal to a broader market.
+
+## Promotional Strategies and Consumer Behavior
+
+Promotional strategies also play a role in shaping beef prices.
+T&T’s greater variability could reflect frequent promotional activities aimed at specific customer segments.
+Promotions may be used to move inventory, introduce new products, or compete with other retailers [@citePromotion].
+This contrasts with Walmart’s approach, which prioritizes stable pricing over frequent discounts.
+
+Consumer preferences likely influence these strategies.
+Price-sensitive shoppers may gravitate toward Walmart for its predictable pricing, while those seeking specialty cuts or products may prefer T&T, even with its price fluctuations.
+Future research could examine the impact of promotions by analyzing temporal pricing data alongside promotional calendars or sales events to better understand how short-term discounts affect pricing patterns and consumer loyalty.
 
@@
 
-While the study offers useful insights into beef pricing dynamics, it is important to acknowledge several limitations.
-First, the dataset used was limited to two vendors, Walmart and T&T, which may not fully represent the diversity of the Canadian grocery market.
-Including more vendors could provide a broader perspective on pricing strategies across different types of retailers, such as discount chains, premium grocers, or independent stores [@citeBrand].
-Additionally, the current analysis did not account for factors like promotions, product quality differences, or customer loyalty programs [@citeProgram], which could play significant roles in pricing decisions.
-Including such variables in future models could provide a more nuanced understanding of the factors influencing beef prices.
+While this study provides a useful analysis of beef pricing dynamics, it has some limitations.
+First, the dataset includes only two vendors, Walmart and T&T, which may not fully represent the diversity of the Canadian grocery market.
+Expanding the study to include additional retailers, such as discount chains, premium grocers, or independent stores [@citeBrand], could offer a broader perspective.
+Additionally, the analysis does not account for factors like promotions, product quality, or customer loyalty programs [@citeProgram], which could significantly influence pricing.
+Future models incorporating these variables could provide a deeper understanding of pricing behavior.
 
-The model assumes a linear relationship between predictors and the outcome variable, which might oversimplify the complexity of real-world pricing behavior.
-Future work could explore more sophisticated modeling techniques, such as non-linear models or machine learning approaches, to capture potential non-linearities in the data.
-Additionally, the study relied on a single dataset, and incorporating external data sources—such as economic indicators, weather data, or detailed sales information—could help validate and enrich the findings.
+The assumption of a linear relationship between predictors and the outcome may oversimplify the complexity of real-world pricing.
+Future studies could explore more flexible modeling techniques, such as non-linear models or machine learning approaches, to better capture these relationships.
+Incorporating external data sources, such as economic indicators, weather patterns, or detailed sales records, could further validate and enrich the findings.
 
-Finally, the assumption of constant variance and normality of residuals may not hold perfectly in the context of pricing data, which can often exhibit heteroscedasticity or other forms of non-normality.
-Implementing more robust regression techniques or exploring heteroscedasticity-consistent estimators could help improve model reliability and accuracy.
-Another interesting avenue for future research would be to examine the impact of consumer behavior changes, such as shifts towards plant-based alternatives, on beef pricing trends, which could provide insights into evolving market dynamics.
+Finally, the assumptions of constant variance and normality of residuals may not hold perfectly in pricing data, which often exhibit heteroscedasticity or other deviations.
+Employing more robust regression techniques or heteroscedasticity-consistent estimators could improve model accuracy and reliability.
+An interesting avenue for future research would be to explore the impact of shifts in consumer preferences, such as increased demand for plant-based alternatives, on beef pricing trends, providing insight into evolving market dynamics.
 

### 2024-11-28 other/datasheet/datasheet.qmd
date fixed

@@
 thanks: "Code and data are available at: [https://jacobfilipp.com/hammer/](https://jacobfilipp.com/hammer/)."
-date: "`r Sys.time()`"
-date-format: "D MMMM YYYY"
-abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
+date: today
+date-format: long
 format: pdf

### 2024-11-28 scripts/03-clean_data.R
updated to save merge data

@@ -68,6 +68,7 @@ cleaned_data <- merge_data %>%
 
 
 #### Save data ####
+write_parquet(merge_data, here("data", "02-analysis_data", "merged_data.parquet"))
 write_parquet(ppu_data, here("data", "02-analysis_data", "ppu_data.parquet"))
 write_parquet(cleaned_data, here("data", "02-analysis_data", "beef_data.parquet"))
 

### 2024-11-28 paper/paper.qmd
paper final draft

@@
 beef_data <- read_parquet(here("data", "02-analysis_data", "beef_data.parquet"))
+merge_data  <- read_parquet(here("data", "02-analysis_data", "merged_data.parquet"))
 
@@
 Also with support of additional packages in R: `readr` [@citeReadr], `ggplot2` [@citeGgplot2], `tidyverse` [@citeTidyverse], `dplyr` [@citeDplyr], `here` [@citeHere], `knitr` [@citeKnitr], `kableExtra` [@citeKableExtra], `rstanarm` [@citeRstanarm], `modelsummary` [@citeModelsummary], `lme4` [@citeLme4], `tinytex` [@citeTinytex], `reshape2` [@citeReshape2], `arrow` [@citeArrow].
+
+Details about data cleaning and variable selection can be found in [Appendix -@sec-data_details].
 
@@
 
-# Additional data details
+# Data details  {#sec-data_details}
+
+## Cleaning methods
+
+In this data cleaning process, the goal was to load raw product and transaction data, merge them, filter and transform the necessary columns, and finally save the cleaned data for further analysis.
+The process began with loading two CSV files: "hammer-4-product.csv" and "hammer-4-raw.csv".
+These files contained information about different products and transactions related to them.
+After loading the data, the next step was to merge the two datasets to create a combined dataset for a more complete product information, as @tbl-merge_data shown.
+
+```{r}
+#| label: tbl-merge_data
+#| tbl-cap: "Raw Merged Data"
+#| echo: false
+#| warning: false
+#| message: false
+
+merge_data %>%
+  head() %>%
+  kable(
+    col.names = c(
+      "Time", 
+      "Vendor", 
+      "Product ID", 
+      "Product Name", 
+      "Brand", 
+      "Current Price", 
+      "Old Price", 
+      "Units", 
+      "Price per Unit"
+    ),
+    booktabs = TRUE
+  )
+
+```
+
+The merging of the two datasets was accomplished using an inner join on the `product_id` and `id` columns.
+This allowed for the selection of relevant columns, such as product details (`product_name`, `brand`), vendor information, and pricing (`current_price`, `old_price`, and `price_per_unit`).
+The goal was to retain only those data attributes that were essential for subsequent analyses.
+
+Once the datasets were merged, the next stage was to focus specifically on price-per-unit data (`price_per_unit`) and create a subset, `ppu_data`, which only included vendors Walmart and TandT.
+Additionally, this subset included only those records where the product names contained "beef", helping to narrow down the dataset to beef-related items.
+Only the `vendor` and `price_per_unit` columns were selected to generate this subset, which would later be saved as a separate parquet file for analysis.
+
+The main cleaning process involved creating a cleaned dataset (@tbl-clean_data), `cleaned_data`, by filtering out only those records from the merged dataset that belonged to Walmart and TandT.
+The data was further refined by selecting important columns, such as transaction time (`nowtime`), vendor, product pricing, and product details.
+Several transformations were then applied to ensure the data was suitable for analysis.
+For example, the `nowtime` column was parsed to extract `year`, `month`, and `day` values, which provided a more granular breakdown of transaction time.
+The `current_price` and `old_price` columns were converted to numeric values using `parse_number()` to handle any non-numeric characters.
+
+```{r}
+#| label: tbl-clean_data
+#| tbl-cap: "Cleaned Data"
+#| echo: false
+#| warning: false
+#| message: false
+
+beef_data %>%
+  select(vendor, current_price, old_price, product_name, price_per_unit, year, month, day) %>%
+  head() %>%
+  kable(
+    col.names = c(
+      "vendor", 
+      "current_price", 
+      "old_price", 
+      "product_name", 
+      "price_per_unit", 
+      "year", 
+      "month", 
+      "day"
+      ),
+    booktabs = TRUE
+  ) %>%
+  kable_styling(full_width = FALSE) %>%
+  column_spec(4, width = "7em")
+
+```
+
+For the `price_per_unit` column, the values were cleaned to ensure consistency.
+The dollar sign was extracted using `str_extract()` and subsequently removed using `str_remove()`.
+The result was then converted to numeric values for accurate comparison and analysis.
+In cases where `price_per_unit` was missing, the `ifelse()` function was used to assign a default value of 0, ensuring no missing data points could interfere with subsequent operations.
+
+To focus specifically on beef products, the `cleaned_data` dataset was filtered to include only those rows where `product_name` contained "beef" while excluding records with terms like "flavour", "vermicelli", "rice", and other similar terms that indicated non-beef products.
+This filtering step was crucial to ensure the final dataset contained only relevant beef-related products, reducing noise from irrelevant items.
+The `nowtime` column, which was no longer needed, was dropped, and rows containing any `NA` values were also removed using `drop_na()`.
+
+Finally, the cleaned data was saved into parquet files: `ppu_data.parquet` for the price-per-unit data subset and `beef_data.parquet` for the full cleaned dataset.
+
+\newpage
 
@@
 
-A **stratified random sampling** method would be employed to ensure representation across key demographic groups, including age, income levels, and cultural background.
-Stratification variables would include:
-
-1. **Region**: Urban versus rural shoppers across Canada.
-2. **Vendor familiarity**: Regular Walmart shoppers, regular T&T shoppers, and those who shop at both.
-3. **Income brackets**: Low, middle, and high-income households.
+A **stratified random sampling** method would be employed to ensure comprehensive representation across key demographic groups, including age, income levels, and cultural background.
+This approach aims to capture the diversity in shopping preferences and behaviors, particularly as they relate to beef pricing in the Canadian grocery market.
+By employing stratified random sampling, the survey would not only provide insights into how different demographic segments make purchasing decisions but also allow for more granular comparisons between consumer groups, which is critical for understanding pricing dynamics in diverse retail environments.
+
+1.
+**Region**: The sample would be divided into urban and rural shoppers across Canada.
+This stratification would ensure representation of different population densities, as consumer preferences and shopping patterns can differ significantly between urban and rural areas.
+Urban shoppers, for example, might have more convenient access to both Walmart and T&T locations, whereas rural shoppers may be limited to fewer options or specific local vendors.
+This distinction is crucial for understanding geographical influences on grocery purchasing behavior, particularly in terms of access, convenience, and price sensitivity.
+
+2.
+**Vendor familiarity**: The survey would categorize respondents into three groups: regular Walmart shoppers, regular T&T shoppers, and those who shop at both retailers.
+Regular Walmart shoppers may prioritize price consistency and availability of products, whereas regular T&T shoppers might be more focused on product variety and specialty items that cater to cultural preferences.
+Including those who shop at both retailers would provide insights into consumer behaviors that shift between general and specialty retail environments, offering a comparative view of how brand familiarity and loyalty shape beef purchasing decisions.
+
+3.
+**Income brackets**: Respondents would be classified into low, middle, and high-income households.
+This stratification is essential for analyzing the impact of financial constraints on beef purchasing patterns.
+Lower-income households might be more sensitive to price fluctuations and promotions, often choosing products based on affordability.
+Middle-income households may exhibit a mix of price sensitivity and preference for product quality, whereas high-income households could prioritize premium product quality over price.
+By segmenting the respondents into income brackets, the survey could identify how economic factors influence both the frequency and type of beef products purchased at Walmart and T&T, revealing the varying levels of price elasticity among different socioeconomic groups.
 
@@
 
-1. **Online Panels**: Partnering with Canadian consumer research platforms to recruit participants from diverse regions.
-2. **Retailer Partnerships**: Collaborating with Walmart and T&T to survey in-store shoppers (e.g., QR codes on receipts leading to the survey).
-3. **Community Outreach**: Engaging with cultural organizations to reach underrepresented groups, particularly for T&T shoppers.
+1.
+**Online Panels**: The recruitment process would involve partnering with established Canadian consumer research platforms to access existing online panels.
+These platforms typically have large databases of participants that are segmented based on various demographics such as age, income, and geographic region, allowing for targeted recruitment.
+By collaborating with these platforms, the survey can include participants from diverse regions—spanning urban centers like Toronto and Vancouver, as well as rural areas across provinces.
+Online panels are advantageous as they provide quick access to a variety of respondents, increasing the efficiency of data collection while allowing for quotas to be set to ensure representativeness across the stratification variables (e.g., income brackets, vendor familiarity).
+
+2.
+**Retailer Partnerships**: Collaborating directly with Walmart and T&T stores would allow for in-store recruitment, making it possible to capture insights from shoppers at the point of purchase.
+A key aspect of this strategy would involve placing QR codes on receipts or at store exits, inviting shoppers to participate in the survey by scanning the code with their mobile devices.
+This in-store recruitment tactic not only encourages immediate participation from shoppers but also provides the opportunity to capture data from consumers while their purchasing experience is still fresh.
+Additionally, leveraging in-store displays, posters, or handouts could help draw more attention to the survey.
+Walmart and T&T staff could also provide gentle prompts to shoppers, further promoting survey participation.
+This approach would help ensure that insights are gathered directly from those actively engaged with these retailers, offering real-time perspectives on beef purchasing behavior.
+
+3.
+**Community Outreach**: To reach underrepresented groups, particularly T&T shoppers from culturally diverse backgrounds, targeted community outreach would be conducted in collaboration with cultural organizations, community centers, and local events.
+Engaging with organizations that have strong connections to the Asian community—such as cultural associations, language schools, or ethnic community centers—would help ensure that the sample includes individuals who may not be as reachable through online panels or in-store promotions alone.
+Additionally, outreach could include attending cultural festivals, placing ads in community newsletters, or utilizing social media channels that serve specific cultural groups.
+This grassroots strategy would be particularly useful in accessing populations that may have different shopping patterns or dietary needs, thereby providing richer insights into how cultural preferences influence beef purchasing decisions.
+Engaging directly with cultural communities would help ensure a higher level of trust and response rate from these groups, who might otherwise be hesitant to participate in market research.
 
@@
 
+The survey design for this study draws on established methodologies in consumer behavior and retail pricing research.
+By employing **stratified random sampling**, this study ensures that the key factors influencing purchasing decisions, such as income, cultural background, and shopping preferences, are effectively captured.
+This approach is informed by existing literature emphasizing the importance of **representative sampling** to understand market segmentation, particularly in heterogeneous consumer markets [@citeSegmentation].
+Stratified sampling is particularly beneficial in capturing the variations in consumer preferences that may be linked to demographic differences, as suggested by **demographic-based segmentation theories** in consumer research.
+
+The use of **multichannel recruitment** has also been shown to be effective in increasing participation rates, especially when attempting to reach a diverse population base [@citeMultichannel].
+Studies have demonstrated that combining online recruitment with community engagement enhances the inclusiveness of the survey sample, thereby reducing selection bias and ensuring that underrepresented groups are adequately represented.
+This is particularly relevant for this study, which aims to understand pricing strategies across different consumer segments, including culturally specific shoppers who might not be well-represented through online-only recruitment.
+
+The inclusion of both **vendor-specific** and **cultural factors** in the survey aligns with research that highlights the role of retail environments and cultural preferences in shaping consumer behavior.
+Literature on cultural consumption patterns suggests that different retail formats, such as mainstream versus culturally specialized stores, influence purchasing decisions based on familiarity, trust, and cultural alignment [@citeCulturalAlignment].
+Therefore, the stratification variables and recruitment strategies are designed to capture these nuanced consumer preferences, providing a comprehensive understanding of how pricing dynamics differ across vendor types and cultural backgrounds.
+
+Moreover, the focus on **vendor familiarity** builds on findings from **brand loyalty** literature, which suggests that consumers' familiarity with a brand or retailer can significantly impact their price sensitivity and shopping behavior [@citeBranding].
+By including vendor familiarity as a key stratification criterion, the survey is structured to better understand how loyalty and brand perception contribute to the pricing dynamics of beef in different retail contexts.
+
+In summary, the survey design leverages established theories and methodologies from consumer behavior, market segmentation, and cultural studies to ensure that the data collected is robust, representative, and capable of answering the core research questions about beef pricing dynamics in the Canadian retail market.
+
 \newpage

### 2024-11-29 paper/paper.qmd
final adjustments

--- 2024-11-28 paper/paper.qmd
@@
 date-format: long
-abstract: "This study analyzes beef pricing dynamics in the Canadian retail market, focusing on two vendors: Walmart and T&T.
-Using Bayesian linear regression, we found that historical pricing and vendor-specific strategies significantly influence current beef prices, with Walmart displaying more price stability and T&T showing greater variability.
-These findings reveal how both market-wide and vendor-specific factors shape beef pricing, offering insights that could help retailers refine their pricing strategies and support policymakers in addressing food affordability.
-This paper provides a clearer understanding of the interplay between cultural preferences and retail strategies in determining food prices."
+abstract: "This study examines beef pricing in the Canadian grocery market, focusing on Walmart and T&T as representative vendors.
+Using Bayesian linear regression, it analyzes how historical prices, vendor-specific strategies, and seasonal effects impact current beef prices.
+The findings indicate that historical pricing strongly predicts current prices, while vendor approaches differ significantly, with Walmart maintaining consistent pricing and T&T showing greater variability.
+This research highlights how cultural preferences and retail strategies influence food pricing, enhancing our understanding of market behavior and supporting both consumer choices and policy development."
 format: pdf
@@
 
-Using a Bayesian linear regression model, the study examines how historical pricing, vendor-specific strategies, and seasonal factors impact current beef prices.
-The analysis indicates that past prices strongly predict current pricing, suggesting a high degree of consistency.
-Differences in pricing strategies between Walmart and T&T further highlight distinct market approaches, with Walmart focusing on stable pricing and T&T displaying greater variability, potentially to address specific consumer preferences.
-These results provide a clearer understanding of pricing strategies in a diverse retail landscape and offer insights into factors that shape affordability and market behavior.
+Using a Bayesian linear regression model, this study analyzes the effects of historical pricing, vendor strategies, and seasonal factors on current beef prices.
+The findings show that past prices are strong predictors of current pricing, indicating a high level of consistency.
+Walmart’s pricing strategy emphasizes stability, while T&T exhibits more variability, likely catering to specific consumer needs.
+These results enhance our understanding of pricing practices across different retail contexts and shed light on factors influencing affordability and purchasing patterns in the market.
 
@@
 
-Finally, the assumptions of constant variance and normality of residuals may not hold perfectly in pricing data, which often exhibit heteroscedasticity or other deviations.
-Employing more robust regression techniques or heteroscedasticity-consistent estimators could improve model accuracy and reliability.
-An interesting avenue for future research would be to explore the impact of shifts in consumer preferences, such as increased demand for plant-based alternatives, on beef pricing trends, providing insight into evolving market dynamics.
+The assumptions of constant variance and normality of residuals may not fully apply to pricing data, which often show heteroscedasticity or other deviations.
+Using more robust regression methods or heteroscedasticity-consistent estimators could enhance the model's accuracy and reliability.
+Future research could investigate how changes in consumer preferences, such as the growing interest in plant-based alternatives, influence beef pricing trends, offering a deeper understanding of shifting market dynamics.
 
@@
 
-# Additional data details
+# Data details  {#sec-data_details}
+
+## Cleaning methods
+
+In the data cleaning process, the objective was to load raw product and transaction data, combine them, filter and transform relevant columns, and save the cleaned dataset for further analysis.
+The process started by loading two CSV files: "hammer-4-product.csv" and "hammer-4-raw.csv." These files contained details about various products and related transactions.
+After loading the data, the datasets were merged to create a unified dataset, providing more complete product information, as shown in @tbl-merge_data.
+
+```{r}
+#| label: tbl-merge_data
+#| tbl-cap: "Raw Merged Data"
+#| echo: false
+#| warning: false
+#| message: false
+
+merge_data %>%
+  head() %>%
+  kable(
+    col.names = c(
+      "Time", 
+      "Vendor", 
+      "Product ID", 
+      "Product Name", 
+      "Brand", 
+      "Current Price", 
+      "Old Price", 
+      "Units", 
+      "Price per Unit"
+    ),
+    booktabs = TRUE
+  )
+
+```
+
+The two datasets were merged using an inner join on the `product_id` and `id` columns.
+This approach ensured that only relevant columns were retained, such as product details (`product_name`, `brand`), vendor information, and pricing (`current_price`, `old_price`, and `price_per_unit`).
+The aim was to keep only the attributes necessary for further analysis.
+
+After merging the datasets, the focus shifted to creating a subset, `ppu_data`, specifically for price-per-unit data (`price_per_unit`).
+This subset included only records from vendors Walmart and T&T, with product names containing "beef," narrowing the dataset to beef-related items.
+Only the `vendor` and `price_per_unit` columns were selected for this subset, which was saved as a separate parquet file for further analysis.
+
+The main cleaning process resulted in the creation of the `cleaned_data` dataset (@tbl-clean_data), which included only records from Walmart and T&T.
+Key columns were retained, such as transaction time (`nowtime`), vendor, product pricing, and product details.
+Additional transformations ensured the dataset was ready for analysis.
+For instance, the `nowtime` column was parsed to extract `year`, `month`, and `day` for a more detailed breakdown of transaction timing.
+The `current_price` and `old_price` columns were converted to numeric values using `parse_number()` to handle any non-numeric characters.
+
+```{r}
+#| label: tbl-clean_data
+#| tbl-cap: "Cleaned Data"
+#| echo: false
+#| warning: false
+#| message: false
+
+beef_data %>%
+  select(vendor, current_price, old_price, product_name, price_per_unit, year, month, day) %>%
+  head() %>%
+  kable(
+    col.names = c(
+      "vendor", 
+      "current_price", 
+      "old_price", 
+      "product_name", 
+      "price_per_unit", 
+      "year", 
+      "month", 
+      "day"
+      ),
+    booktabs = TRUE
+  ) %>%
+  kable_styling(full_width = FALSE) %>%
+  column_spec(4, width = "7em")
+
+```
+
+The `price_per_unit` column was cleaned to ensure consistency.
+The dollar sign was first extracted using `str_extract()` and then removed with `str_remove()`.
+The resulting values were converted to numeric format to allow accurate comparisons and analysis.
+For cases where `price_per_unit` was missing, the `ifelse()` function assigned a default value of 0, ensuring that no missing data points interfered with subsequent operations.
+
+To refine the dataset for beef products, the `cleaned_data` dataset was filtered to retain only rows where `product_name` contained "beef." Records with terms such as "flavour," "vermicelli," and "rice," which indicated non-beef items, were excluded.
+This filtering step ensured the final dataset was focused on beef-related products, reducing noise from irrelevant entries.
+The `nowtime` column, which was no longer required, was dropped, and any rows containing `NA` values were removed using `drop_na()`.
+
+The cleaned data was then saved into parquet files: `ppu_data.parquet` for the price-per-unit subset and `beef_data.parquet` for the complete cleaned dataset.
+
+\newpage
 
 ## Vendor Choice
-The selection of Walmart and T&T as representative vendors for this analysis was driven by their distinctive characteristics and market coverage, which together offer a comprehensive view of beef pricing dynamics.
-Walmart is a prominent multinational retailer with a broad presence across North America, making it a well-suited representative of mainstream, large-scale grocery retailing.
-Its extensive reach and emphasis on standardized pricing provide valuable insights into the general market trends and pricing behaviors that are accessible to a wide segment of consumers.
-By including Walmart in the analysis, we can capture a perspective that is reflective of the typical shopping experience for many consumers, characterized by competitive pricing and a broad product selection.
-
-On the other hand, T&T Supermarket represents a niche segment of the market, catering specifically to Asian communities and consumers seeking specialty goods that align with Asian culinary traditions.
-T&T's focus on products tailored to Asian tastes, along with its unique supply chains and vendor relationships, provides an important contrast to Walmart.
-The inclusion of T&T allows for an exploration of how cultural preferences and niche market positioning influence pricing.
-By analyzing T&T, we gain insights into pricing practices that reflect the demands of a distinct consumer group and how these differ from the broader market.
-
-The combination of Walmart and T&T enables the analysis to explore both mainstream and culturally specific market behaviors.
-This choice of vendors thus enriches the study by accounting for variations in consumer preferences, pricing strategies, and market positioning.
-Walmart's emphasis on scale and cost-efficiency contrasts with T&T's specialization and cultural targeting, allowing us to draw meaningful comparisons in terms of beef pricing strategies between a general and a specialized market context.
-Such a nuanced understanding is essential for capturing the diversity within the beef retail market, providing a more holistic view of the factors driving beef prices in different retail settings.
+
+The selection of Walmart and T&T as vendors for this analysis was based on their distinct characteristics and market coverage, offering a diverse view of beef pricing.
+Walmart, as a leading multinational retailer with a wide presence across North America, represents mainstream, large-scale grocery retailing.
+Its extensive reach and focus on standardized pricing make it a strong representative of general market trends.
+Including Walmart in the analysis helps capture pricing behaviors that reflect the typical shopping experience for a broad consumer base, characterized by competitive pricing and a wide product range.
+
+In contrast, T&T Supermarket caters to a niche market, serving Asian communities and consumers looking for specialty goods that align with Asian culinary traditions.
+T&T's emphasis on culturally specific products, unique supply chains, and vendor relationships provides a valuable comparison to Walmart.
+By analyzing T&T, this study examines how cultural preferences and niche market strategies influence pricing, reflecting the demands of a distinct consumer segment and highlighting differences from broader market trends.
+
+Including both Walmart and T&T allows this analysis to examine mainstream and culturally specific market behaviors.
+Walmart’s focus on scale and cost-efficiency contrasts with T&T’s specialization and cultural targeting, enabling meaningful comparisons in beef pricing strategies across different retail contexts.
+This approach provides a more detailed understanding of the factors shaping beef prices in diverse settings, highlighting variations in consumer preferences, pricing strategies, and market positioning.
 
@@
 
-In cleaning the `price_per_unit` variable, I began by extracting the numerical value using `str_extract()`, which allowed me to identify the dollar amounts in the original string (`\\$[0-9\\.]+`), as shown in @tbl-price_per_unit-1.
-This step ensured that only valid price values, starting with a dollar sign, were captured, removing any extraneous text or symbols.
-
-Next, I removed the dollar sign itself with `str_remove("\\$")` to convert the extracted value into a format suitable for numerical operations.
-After this, I used `as.numeric()` to convert the string into a numeric type, allowing for further quantitative analysis.
-
-To handle missing values, I applied `ifelse(is.na(price_per_unit), 0, price_per_unit)`, which replaced any `NA` values in the `price_per_unit` column with `0`.
-This step ensured that all rows had a value for `price_per_unit`, preventing issues during subsequent analysis.
-Using zero as the replacement helped maintain continuity in the dataset, although I acknowledge that this choice could imply that missing values are negligible, which might require additional context or consideration during interpretation.
-The price_per_unit column in the dataset after cleaning looks like @tbl-price_per_unit-2.
+To clean the `price_per_unit` variable, I first extracted the numerical values using `str_extract()`, capturing valid price amounts formatted with a dollar sign (`\\$[0-9\\.]+`), as shown in @tbl-price_per_unit-1.
+This step ensured that only price-related information was retained, removing any extraneous text or symbols from the original string.
+
+I then removed the dollar sign using `str_remove("\\$")` to prepare the extracted values for numerical operations.
+After this, the cleaned strings were converted into a numeric format using `as.numeric()`, enabling quantitative analysis.
+
+To address missing values, I used `ifelse(is.na(price_per_unit), 0, price_per_unit)`, replacing `NA` values in the `price_per_unit` column with `0`.
+This ensured that every row had a value for `price_per_unit`, preventing disruptions during further analysis.
+While assigning zero helped maintain dataset continuity, this approach assumes that missing values are negligible, which may require additional context during interpretation.
+The cleaned `price_per_unit` column is shown in @tbl-price_per_unit-2.
 
@@
 ## Posterior predictive check
-In @fig-ppcheck, we implement a posterior predictive check.
-This plot shows the observed data distribution (in dark lines) compared with simulated datasets from the posterior (in light grey lines).
-The purpose of this check is to evaluate how well the model can reproduce the observed data.
-The close alignment between the simulated posterior predictive distributions and the observed data suggests that the model is accurately capturing the underlying patterns in the data.
-Specifically, we observe that the peaks and troughs in the observed data are well represented by the posterior samples, indicating that the model is capable of capturing key features such as the general shape and spread of the data distribution.
-The variability across the posterior predictive samples is consistent with the observed variability, suggesting no significant model misfit.
+
+In @fig-ppcheck, a posterior predictive check is performed to assess the model's ability to reproduce the observed data.
+The plot compares the observed data distribution (shown in dark lines) with simulated datasets from the posterior (represented by light grey lines).
+This check evaluates how well the model captures the underlying patterns in the data.
+The close alignment between the observed data and the simulated posterior distributions indicates that the model effectively reflects the data's characteristics.
+Specifically, the observed peaks and troughs are well represented by the posterior samples, demonstrating the model's ability to capture key features such as the overall shape and spread of the distribution.
+Additionally, the variability in the posterior predictive samples aligns with the observed variability, suggesting no significant issues with model fit.
 
@@
 
-In @fig-posteriorvsprior, we compare the posterior with the prior distributions for each parameter.
-The posterior distribution (shown on the left) is more concentrated compared to the prior (shown on the right), which demonstrates how the data has influenced and updated our beliefs about the model parameters.
-For instance, the posterior distribution for the intercept has shrunk considerably compared to its prior, indicating that the data has substantially informed our estimate of the baseline price level.
-Similarly, for the other parameters (month, old_price, sigma, and vendorWalmart), the posteriors are less dispersed than the priors, suggesting a meaningful reduction in uncertainty.
-Notably, the posterior intervals are much narrower, especially for the intercept and old price parameters, indicating that the data has provided strong information about these effects.
-The results demonstrate that the priors were sufficiently weak, allowing the data to dominate the inference and significantly refine the parameter estimates.
+In @fig-posteriorvsprior, the posterior and prior distributions for each parameter are compared.
+The posterior distributions (on the left) are more concentrated than the priors (on the right), showing how the data has updated the initial parameter estimates.
+For example, the posterior distribution for the intercept is much narrower than its prior, indicating that the data has strongly informed the estimate of the baseline price level.
+Similarly, the posterior distributions for the other parameters (month, old_price, sigma, and vendorWalmart) are less dispersed than the priors, reflecting a meaningful reduction in uncertainty.
+
+The posterior intervals are particularly narrow for the intercept and old_price parameters, suggesting that the data has provided substantial information about these effects.
+This comparison demonstrates that the priors were appropriately weak, allowing the data to play a dominant role in refining the parameter estimates.
+The results highlight how the model effectively incorporates data to improve the precision of the estimates.
 
@@
 ## Diagnostics
-In @fig-trace, we present a trace plot.
-This plot shows the sampled values for each parameter across the MCMC iterations.
-The trace plot helps assess whether the chains have converged to a stationary distribution.
-In these plots, we see that each parameter's chains are well mixed, with no discernible trends, indicating that the MCMC algorithm has adequately explored the parameter space.
-The lack of significant upward or downward drift suggests that all chains are consistently sampling from the target distribution, which is a strong indicator of convergence.
-
-Specifically, the chains for each parameter appear to oscillate around a constant mean, and the overlap between the four chains is extensive.
-This well-mixed appearance across all parameters, including the intercept, month, old price, vendorWalmart, and sigma, gives us confidence that our model has reached convergence, with each chain sampling from similar regions of the posterior distribution.
+
+In @fig-trace, a trace plot is presented, showing the sampled values for each parameter across the MCMC iterations.
+The trace plot is used to evaluate whether the chains have converged to a stationary distribution.
+In these plots, the chains for each parameter are well mixed, with no visible trends, indicating that the MCMC algorithm has effectively explored the parameter space.
+The absence of upward or downward drift suggests that the chains are consistently sampling from the target distribution, providing strong evidence of convergence.
+
+Each parameter’s chains oscillate around a stable mean, with significant overlap among the four chains.
+This consistent mixing is observed across all parameters, including the intercept, month, old price, vendorWalmart, and sigma.
+The overlap and stability across chains indicate that the model has achieved convergence, with all chains sampling from similar regions of the posterior distribution.
 
@@
 
-In @fig-rhat, we present an Rhat plot.
-The Rhat statistic (also known as the potential scale reduction factor) measures the ratio of the variance between chains to the variance within chains, and values close to 1 indicate that the chains have converged.
-In this plot, we see that all Rhat values for the parameters are very close to 1, specifically below the threshold of 1.05, suggesting that convergence has been achieved across all parameters.
-
-The fact that all Rhat values are below 1.05 suggests that the different chains are in agreement about the underlying posterior distributions.
-This indicates that our model's parameter estimates are stable, and the MCMC algorithm has effectively converged.
-As such, we can be confident in the reliability of the estimates provided by our Bayesian model.
+In @fig-rhat, an Rhat plot is presented to assess model convergence.
+The Rhat statistic (or potential scale reduction factor) compares the variance between chains to the variance within chains, with values close to 1 indicating convergence.
+In this plot, all Rhat values for the parameters are very close to 1 and below the threshold of 1.05, confirming that the chains have converged.
+
+The low Rhat values indicate that the chains are consistent and in agreement about the posterior distributions.
+This suggests that the model's parameter estimates are stable and that the MCMC algorithm has effectively converged.
+As a result, we can be confident in the reliability of the parameter estimates produced by the Bayesian model.
 
@@
 
-To complement the quantitative analysis presented in this paper, an idealized survey methodology could provide direct insights into consumer behavior and perceptions regarding beef pricing at Walmart and T&T.
-This survey would aim to identify factors influencing purchasing decisions, perceptions of price fairness, and preferences for vendor-specific attributes, offering a holistic understanding of the dynamics underpinning beef pricing in the Canadian grocery market.
+To enhance the quantitative analysis in this paper, an idealized survey methodology could capture consumer behavior and perceptions related to beef pricing at Walmart and T&T.
+The survey would aim to identify factors influencing purchasing decisions, perceptions of price fairness, and preferences for vendor-specific attributes.
+This approach would provide a broader understanding of the factors shaping beef pricing dynamics in the Canadian grocery market.
 
@@
 
-A **stratified random sampling** method would be employed to ensure representation across key demographic groups, including age, income levels, and cultural background.
-Stratification variables would include:
-
-1. **Region**: Urban versus rural shoppers across Canada.
-2. **Vendor familiarity**: Regular Walmart shoppers, regular T&T shoppers, and those who shop at both.
-3. **Income brackets**: Low, middle, and high-income households.
-
-A sample size of approximately 1,200 respondents would be targeted, with 400 participants in each stratum (Walmart-focused, T&T-focused, and dual shoppers).
-This stratification would allow for meaningful comparisons between vendor-specific perceptions and strategies.
+A **stratified random sampling** method would be used to ensure representation across key demographic groups, including age, income levels, and cultural background.
+This approach aims to reflect the diversity in shopping preferences and behaviors, particularly concerning beef pricing in the Canadian grocery market.
+Stratified random sampling would enable the survey to capture differences in purchasing decisions among various demographic segments and facilitate detailed comparisons between consumer groups, providing a clearer understanding of pricing patterns in diverse retail settings.
+
+1.
+**Region**: The sample would include both urban and rural shoppers across Canada to account for differences in population density and shopping patterns.
+Urban consumers are likely to have easier access to Walmart and T&T locations, while rural shoppers may rely on fewer options or local vendors.
+This regional distinction is important for understanding how geography influences purchasing behavior, particularly regarding access, convenience, and sensitivity to price.
+
+2.
+**Vendor Familiarity**: Respondents would be grouped into three categories: regular Walmart shoppers, regular T&T shoppers, and those who shop at both retailers.
+Walmart shoppers might focus on consistent pricing and product availability, while T&T shoppers may prioritize variety and specialty items that align with cultural preferences.
+Including consumers who shop at both stores would provide a comparative perspective on how purchasing behavior shifts between general and specialty retail settings, highlighting the role of brand familiarity and loyalty in shaping beef-buying decisions.
+
+3.
+**Income Brackets**: Respondents would be grouped into low-, middle-, and high-income households to assess how financial resources influence beef purchasing behaviors.
+Lower-income households are likely to emphasize affordability and may be more sensitive to price changes and promotions.
+Middle-income households might balance cost with considerations of quality, while high-income households could prioritize premium quality, showing less sensitivity to price.
+This stratification allows for a detailed analysis of how economic factors impact both the frequency and type of beef products purchased at Walmart and T&T, helping to identify variations in price elasticity across income levels.
+
+A total sample size of approximately 1,200 respondents would be targeted, with 400 participants in each stratum (Walmart-focused, T&T-focused, and dual shoppers).
+This approach ensures meaningful comparisons of consumer perceptions and retailer strategies.
 
@@
 1. **Demographics and Shopping Habits**  
-   - Frequency of grocery shopping (weekly, bi-weekly, etc.).
-   - Preferred vendor for beef purchases and reasons for preference.
-   - Average budget allocated for beef purchases.
+   - Frequency of grocery shopping (e.g., weekly, bi-weekly).  
+   - Preferred vendor for beef purchases, along with reasons for the preference.  
+   - Average budget allocated specifically for beef purchases.
 
 2. **Price Perception**  
-   - Rating the perceived fairness of beef prices on a Likert scale.
-   - Comparison of beef prices at Walmart and T&T (e.g., cheaper, equivalent, more expensive).
-   - Sensitivity to price changes (e.g., a 10% increase in beef prices would lead to…).
+   - Rating of beef price fairness on a Likert scale.  
+   - Comparison of beef prices between Walmart and T&T (e.g., cheaper, equivalent, or more expensive).  
+   - Sensitivity to price changes (e.g., how a 10% increase in beef prices might influence purchasing behavior).
 
 3. **Cultural and Quality Preferences**  
-   - Importance of cultural alignment in beef products (specific cuts, preparation styles, etc.).
-   - Perceived quality differences between Walmart and T&T beef products.
-   - Willingness to pay a premium for culturally tailored products.
+   - Importance of cultural alignment in beef products (e.g., availability of specific cuts or preparation styles).  
+   - Perceptions of quality differences in beef products offered by Walmart and T&T.  
+   - Willingness to pay more for products that are culturally tailored.
 
 4. **External Factors**  
-   - Awareness of seasonal price trends.
-   - Impact of promotions on purchasing decisions.
-   - Responses to supply chain disruptions (e.g., alternative purchasing strategies).
+   - Awareness of seasonal price trends and their impact on buying decisions.  
+   - Influence of promotions on purchasing behavior.  
+   - Adjustments made in response to supply chain disruptions (e.g., choosing alternative products or vendors).
 
@@
 
-5. How fair do you find the current beef prices at Walmart/T&T?
+5. How fair do you find the current beef prices at Walmart or T&T?  
    - Very unfair  
@@
 
-8. How important is it for you that beef products align with your cultural preferences (e.g., specific cuts, preparation styles)?
+8. How important is it that beef products match your cultural preferences (e.g., specific cuts, preparation styles)?  
    - Not important at all  
@@
 
-9. Are you willing to pay more for beef products that are culturally tailored to your preferences?
+9. Are you willing to pay more for beef products tailored to your cultural preferences?  
    - Not willing at all  
@@
 
-11. How do promotions impact your decision to purchase beef?
-    - No impact
-    - Slight impact
-    - Moderate impact
-    - Significant impact
-    - Very significant impact
+11. How much do promotions influence your decision to purchase beef?  
+    - No influence  
+    - Slight influence  
+    - Moderate influence  
+    - Significant influence  
+    - Very significant influence
 
@@
 
-1. **Online Panels**: Partnering with Canadian consumer research platforms to recruit participants from diverse regions.
-2. **Retailer Partnerships**: Collaborating with Walmart and T&T to survey in-store shoppers (e.g., QR codes on receipts leading to the survey).
-3. **Community Outreach**: Engaging with cultural organizations to reach underrepresented groups, particularly for T&T shoppers.
+1.
+**Online Panels**: Recruitment would involve partnering with established Canadian consumer research platforms to access existing online panels.
+These platforms have extensive databases of participants segmented by demographics such as age, income, and geographic location, enabling targeted recruitment.
+Collaborating with these platforms ensures the inclusion of participants from diverse areas, spanning urban centers like Toronto and Vancouver to rural regions across the provinces.
+Online panels offer a quick and efficient way to reach respondents while allowing quotas to be set for representativeness based on stratification variables, such as income levels and familiarity with specific retailers.
+
+2.
+**Retailer Partnerships**: Collaborating with Walmart and T&T stores for in-store recruitment would capture input directly from shoppers at the point of purchase.
+QR codes placed on receipts or near store exits could invite shoppers to participate by scanning with their mobile devices.
+This approach encourages immediate participation while the shopping experience is still fresh.
+Additional promotional efforts, such as in-store displays, posters, or handouts, would increase visibility, while staff could provide gentle reminders to shoppers about the survey.
+This strategy ensures the inclusion of active shoppers at Walmart and T&T, providing real-time feedback on beef purchasing habits and preferences.
+
+3.
+**Community Outreach**: To reach underrepresented groups, particularly T&T shoppers from diverse cultural backgrounds, targeted outreach would be conducted through partnerships with cultural organizations, community centers, and local events.
+Collaborating with groups such as cultural associations, language schools, or ethnic community centers would help include participants who might not be reachable through online panels or in-store recruitment alone.
+Outreach efforts could also involve attending cultural festivals, placing advertisements in community newsletters, and utilizing social media platforms tailored to specific cultural groups.
+This approach ensures the inclusion of shoppers with unique cultural preferences and shopping behaviors.
+Engaging directly with these communities fosters trust and encourages participation, leading to a more complete understanding of how cultural factors shape beef purchasing decisions.
 
@@
 
+The survey design for this study is informed by established methodologies in consumer behavior and retail pricing research.
+By using **stratified random sampling**, the study ensures that key factors influencing purchasing decisions—such as income, cultural background, and shopping preferences—are effectively represented.
+This approach draws on research highlighting the importance of **representative sampling** to analyze market segmentation in diverse consumer markets [@citeSegmentation].
+Stratified sampling is particularly useful for capturing differences in preferences that may arise from demographic variations, as outlined in **demographic-based segmentation theories** in consumer studies.
+
+The adoption of **multichannel recruitment** is also important for increasing participation rates, especially when targeting a diverse population [@citeMultichannel].
+Combining online recruitment with community-based efforts has been shown to improve inclusivity, reduce selection bias, and ensure adequate representation of underrepresented groups.
+This approach is especially relevant for a study aiming to examine pricing strategies across various consumer segments, including culturally specific shoppers who may not engage with online-only recruitment methods.
+
+Including both **vendor-specific** and **cultural factors** in the survey design aligns with research on the influence of retail environments and cultural preferences on consumer behavior.
+Studies on cultural consumption patterns suggest that mainstream and culturally specialized stores influence purchasing decisions differently, based on familiarity, trust, and cultural alignment [@citeCulturalAlignment].
+As a result, the stratification variables and recruitment methods are structured to capture these consumer preferences, offering a detailed understanding of how pricing dynamics vary between vendor types and cultural groups.
+
+The focus on **vendor familiarity** builds on findings from the **brand loyalty** literature, which emphasize the role of familiarity in shaping price sensitivity and shopping behavior [@citeBranding].
+Including vendor familiarity as a stratification criterion allows the survey to investigate how loyalty and perceptions of brand reliability influence beef pricing dynamics in different retail settings.
+
+In conclusion, the survey design integrates established theories and practices from consumer behavior, market segmentation, and cultural studies.
+This approach ensures that the data collected is representative, thorough, and suited to addressing the research objectives regarding beef pricing dynamics in the Canadian retail market.
+
 \newpage

### 2024-11-29 paper/paper.qmd
final draft

@@
 ---
-title: "Beef Pricing Dynamics in the Canadian Grocery Market: A Comparative Analysis of Walmart and T&T"
+title: "Beef Pricing Dynamics in the Canadian Grocery Market, 2024: A Comparative Analysis of Walmart and T&T"
 subtitle: "Evidence of Cultural and Strategic Influences on Beef Pricing"
@@
 #| label: fig-current_price
-#| fig-cap: "Boxplot of the distribution of current prices"
-#| echo: false
-#| warning: false
-#| message: false
-#| fig-width: 5
+#| fig-cap: "Distribution of current prices"
+#| echo: false
+#| warning: false
+#| message: false
+#| fig-width: 3
 
